---
title: "B_NB_stats"
author: "Shayle Matsuda"
date: "08/08/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

Color score for *M. capitata* and *P. compressa* colonies at the Inner Lagoon (PR4) and Outer Lagoon (PR13) after the 2015 coral bleaching event in Kaneohe Bay.  Corals in both the benthic surveys and the tagged individuals were given a color score based on their visual color: 1) white (>80% of colony white with no visible pigmentation); 2) pale (>10% colony affected by pigment loss); or 3) fully pigmented (<10% colony with any pale coloration). 

```{r setup, include=FALSE,results="hide"}
knitr::opts_chunk$set(echo = TRUE)
```

## Bleached/Non Bleached data 
```{r, results="hide", message=FALSE}
#Set up: load libraries

rm(list=ls(all=TRUE)) 
setwd("~/Desktop/B_NB")
library("reshape2") #reshape data
library("plotrix") #functions in tapply
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library(effects)
library("multcomp") 
library("multcompView")
library("coefplot") #coefficient plotting
library("seacarb") #seawater carbonate chemistry
library("vegan") #calculating distance matrices
library(betareg)
library(coin)
library(exactRankTests)
library(mvtnorm)
library(DTK)
library(PMCMR)
library(MuMIn)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(ggsignif)
library(HH)
library(PMCMR)
library(tidyverse)
library(arm)
library(jtools)
library(broom)
library(ggstance)
library(officer)
library(huxtable)
library(flextable)

#We excluded these pairs as part of the pair analysis because the Bleached coral was only a 2: Porites Outer Bay 233_234, Porites Inner Bay 297_298, Porites Outer Bay 227_228, Montipora capitata Inner Bay 251_252.

# Bleaching data
bData<-read.csv("Bleaching_FINAL.csv")
bData$Site<-as.factor(bData$Site)
bData$Pair<-as.factor(bData$Pair)

#organize data
bALL<-bData
  #melt
  bALLData1<-melt(data = bALL, measure.vars = c("Month0","Month1.5","Month3","Month4.5", "Month6","Month10","Month18","Month24"))   
  colnames(bALLData1)[colnames(bALLData1)=="variable"] <- "month"   # change col name from 'variable' to 'month'
  colnames(bALLData1)[colnames(bALLData1)=="value"] <- "score"   # change col name from 'value' to 'score'
  bALLData1$group<-paste(bALLData1$Species, bALLData1$Site, bALLData1$B2015, bALLData1$month) # add column for groups
#write.csv(bALL, "PigmentationBNB_20191122.csv")

#Remove all NON PAIRS and all NAs
  bALLData<-na.omit(bALLData1) # remove the NA colums

#subset 0-3 months
  b03<-subset(bALLData, month=="Month0" |month=="Month3")
  b03<-subset(b03, select = -group )

      b03<-dcast(b03, Species+Site+Pair+B2015~month) # cast()
  b03<-na.omit(b03) # remove the NA colums
  
#subset 0-6 months
b06<-subset(bALLData, month=="Month0" |month=="Month6")
b06<-subset(b06, select = -group )

      b06<-dcast(b06, Species+Site+Pair+B2015~month) # cast()
      b06<-na.omit(b06) # remove the NA colums

#subset 0-1.5-3 months for pairs comparison
b0153<-subset(bALLData, month=="Month0" |month=="Month1.5"|month=="Month3")
   #   b0153<-dcast(b0153, Species+Site+Pair+B2015~month) # cast()
   #   b0153<-na.omit(b0153) # remove the NA colums

#subset 3-4.5-6 months for pairs comparison
b3456<-subset(bALLData, month=="Month3" |month=="Month4.5"|month=="Month6")

#subset 0-1.5-3-4.5-6 months for pairs comparison
b0153456<-subset(bALLData, month=="Month0" |month=="Month1.5"|month=="Month3"|month=="Month4.5"|month=="Month6")
      #b0153456<-dcast(b0153456, Species+Site+Pair+B2015~month) # cast()
      #b0153456<-na.omit(b0153456) # remove the NA colums
    #subset 0-1.5-3-4.5-6 months for pairs comparison
``` 

Mean bleaching scores
```{r, results=TRUE}
#calc average bleaching score per grouping by time point
ballmeans <- ddply(bALLData, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

ballmeans$code<-paste0(ballmeans$Site, ballmeans$B2015, ballmeans$Species)
ballmeans #display table
#write.csv(ballmeans,"ballmeans_20191108.csv")
```

## Bleaching score means 0-24 months by species/site/bleaching history  
*Montipora capitata*  
Figure 5
```{r}
#subset for Mcap only
Mcap024Allplot<-subset(ballmeans, Species=="Montipora capitata")     #subset Mcap
Mcap024Allplot$Month[Mcap024Allplot$month=="Month0"]<-0              # new col with month #s only
Mcap024Allplot$Month[Mcap024Allplot$month=="Month1.5"]<-1.5  
Mcap024Allplot$Month[Mcap024Allplot$month=="Month3"]<-3 
Mcap024Allplot$Month[Mcap024Allplot$month=="Month4.5"]<-4.5
Mcap024Allplot$Month[Mcap024Allplot$month=="Month6"]<-6  
Mcap024Allplot$Month[Mcap024Allplot$month=="Month10"]<-10  
Mcap024Allplot$Month[Mcap024Allplot$month=="Month18"]<-18
Mcap024Allplot$Month[Mcap024Allplot$month=="Month24"]<-24 

#PR13
Mcap024AllplotPR13<-subset(Mcap024Allplot, Site==13)
#PR4
Mcap024AllplotPR4<-subset(Mcap024Allplot, Site==4)
```

*M. capitata* PR13
```{r}
ballplotMcapPR13<-ggplot(data=Mcap024AllplotPR13, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold",hjust = 0.5));ballplotMcapPR13
McapOuter<-ballplotMcapPR13 + theme(legend.position = "none") #remove legend
```

*M. capitata* PR4
```{r}
ballplotMcapPR4<-ggplot(data=Mcap024AllplotPR4, aes(x=Month, y=mean, group = code, color=B2015),show.legend = FALSE) + #color by bleaching hist. 
 geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Montipora capitata\nColor score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold", hjust = 0.5));ballplotMcapPR4
McapInner<-ballplotMcapPR4 + theme(legend.position = "none") #remove legend
McapInner
```

*Porites compressa*  
Figure 5
```{r}
Pcomp024Allplot<-subset(ballmeans, Species=="Porites compressa")

# change Month# to just number
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month0"]<-0
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month1.5"]<-1.5
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month3"]<-3
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month4.5"]<-4.5
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month6"]<-6
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month10"]<-10
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month18"]<-18
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month24"]<-24

Pcomp024AllplotPR13<-subset(Pcomp024Allplot, Site==13)
Pcomp024AllplotPR4<-subset(Pcomp024Allplot, Site==4)
```

*P. compress* 
Outer PR13
```{r}
b024plotPcompPR13<-ggplot(data=Pcomp024AllplotPR13, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Porites compressa Bleaching Scores months 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR13

PcompOuter<-b024plotPcompPR13+ theme(legend.position = "none") #remove legend
```

*P. compressa*   
Inner PR4
```{r}
b024plotPcompPR4<-ggplot(data=Pcomp024AllplotPR4, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Porites compressa\nColor score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Porites compressa Bleaching Scores months 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR4


PcompInner<-b024plotPcompPR4+ theme(legend.position = "none") #remove legend
PcompInner
```

Plot all four together, Figure 5
```{r}
grid.arrange(McapInner, McapOuter, PcompInner,PcompOuter, nrow = 2)
```


Table 2  
PIGMENTATION 0-3 months and 0-6 months
score~month*B2015*lagoon
*M. capitata*
```{r}
Mcappig03All<-subset(b0153, Species=="Montipora capitata")
Mcappig06All<-subset(b0153456, Species=="Montipora capitata")
Mcappig36All<-subset(b3456, Species=="Montipora capitata")

#03
Mcappig03All_mod<-aov(score~month*Site*B2015, Mcappig03All)
summary(Mcappig03All_mod)
TukeyHSD(Mcappig03All_mod)
#06
Mcappig06All_mod<-aov(score~month*Site*B2015, Mcappig06All)
summary(Mcappig06All_mod)
TukeyHSD(Mcappig06All_mod)
```

*P. compressa*  
Table 2
```{r}
Pcomppig03All<-subset(b0153, Species=="Porites compressa")
Pcomppig06All<-subset(b0153456, Species=="Porites compressa")
Pcomppig36All<-subset(b3456, Species=="Porites compressa")

#0-3
Pcomppig03All_mod<-aov(score~month*Site*B2015, Pcomppig03All)
summary(Pcomppig03All_mod)
TukeyHSD(Pcomppig03All_mod)
#0-6
Pcomppig06All_mod<-aov(score~month*Site*B2015, Pcomppig06All)
summary(Pcomppig06All_mod)
TukeyHSD(Pcomppig06All_mod)

#36
Pcomppig36All_mod<-aov(score~month*Site*B2015, Pcomppig36All)
summary(Pcomppig36All_mod)
TukeyHSD(Pcomppig36All_mod)
```

# Bleaching Recovery Models
SI Table 11  
`Bleaching Score ~ Reef Site * Bleaching History   
Model: `score~Site*B2015`
*Montipora capiatata*  
```{r}
Mcap_bALLData<-subset(bALLData, Species=="Montipora capitata")   # pull out Mcap

#Month0
Mcap_b0<-subset(Mcap_bALLData, month=="Month0")
Mcap_b0_mod1<-aov(score~Site*B2015, Mcap_b0)
summary(Mcap_b0_mod1)
TukeyHSD(Mcap_b0_mod1)

#Month1.5
Mcap_b1_5<-subset(Mcap_bALLData, month=="Month1.5")
Mcap_b1_5_mod1<-aov(score~Site*B2015, Mcap_b1_5)
summary(Mcap_b1_5_mod1)
TukeyHSD(Mcap_b1_5_mod1)

#Month3
Mcap_b3<-subset(Mcap_bALLData, month=="Month3")
Mcap_b3_mod1<-aov(score~Site*B2015, Mcap_b3)
summary(Mcap_b3_mod1)
TukeyHSD(Mcap_b3_mod1)

#Month4.5
Mcap_b4_5<-subset(Mcap_bALLData, month=="Month4.5")
Mcap_b4_5_mod1<-aov(score~Site*B2015, Mcap_b4_5)
summary(Mcap_b4_5_mod1)
TukeyHSD(Mcap_b4_5_mod1)

#Month6
Mcap_b6<-subset(Mcap_bALLData, month=="Month6")
Mcap_b6_mod1<-aov(score~Site*B2015, Mcap_b6)
summary(Mcap_b6_mod1)
TukeyHSD(Mcap_b6_mod1)
```

*Porites compressa*  
SI Table 11
```{r}
Pcomp_bALLData<-subset(bALLData, Species=="Porites compressa")   # pull out Pcomp

#Month0
Pcomp_b0<-subset(Pcomp_bALLData, month=="Month0")
Pcomp_b0_mod1<-aov(score~Site*B2015, Pcomp_b0)
summary(Pcomp_b0_mod1)
TukeyHSD(Pcomp_b0_mod1)

#Month1.5
Pcomp_b1_5<-subset(Pcomp_bALLData, month=="Month1.5")
Pcomp_b1_5_mod1<-aov(score~Site*B2015, Pcomp_b1_5)
summary(Pcomp_b1_5_mod1)
TukeyHSD(Pcomp_b1_5_mod1)

#Month3
Pcomp_b3<-subset(Pcomp_bALLData, month=="Month3")
Pcomp_b3_mod1<-aov(score~Site*B2015, Pcomp_b3)
summary(Pcomp_b3_mod1)

#Month4.5
Pcomp_b4_5<-subset(Pcomp_bALLData, month=="Month4.5")
Pcomp_b4_5_mod1<-aov(score~Site*B2015, Pcomp_b4_5)
summary(Pcomp_b4_5_mod1)

#Month6
Pcomp_b6<-subset(Pcomp_bALLData, month=="Month6")
Pcomp_b6_mod1<-aov(score~Site*B2015, Pcomp_b6)
summary(Pcomp_b6_mod1)
```

## Rare of Bleaching Recovery: Percent change from month 0 to month 3  
We calculated the mean bleaching recovery at 0-3 months and 3-6 months.      
0-3 months RATE
```{r}
## 1. 0-3 months
#Calc average rate per colony 0-3 months
# Recovery in %
b03$Rate03<-(((b03$Month3-b03$Month0)))

#Average the means of the rates. 

# calc average  bleaching rate/month in %  per grouping
b03means <- ddply(b03, c("Site", "Species", "B2015"), summarise, 
                   N    = length(Rate03[!is.na(Rate03)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Rate03, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Rate03, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Rate03, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#b03means #display table

b03means$code<-paste0(b03means$Site, b03means$B2015, b03means$Species)
b03means #display table
```

0-6 months RATE
```{r}
#recovery in %
b06$Rate06<-(((b06$Month6-b06$Month0)))

#Average the means of the rates. 

# calc average  bleaching rate/month in %  per grouping
b06means <- ddply(b06, c("Site", "Species", "B2015"), summarise, 
                   N    = length(Rate06[!is.na(Rate06)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Rate06, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Rate06, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Rate06, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
b06means #display table

b06means$code<-paste0(b06means$Site, b06means$B2015, b06means$Species)
```

Rate change  
`Rate ~ Lagoon*B2015`  
*Montipora capitata only*    
 Months 0-3
```{r}
#subset Mcap
Mcap_b03<-subset(b03, Species=="Montipora capitata") #subset Mcap
#0-3
Mcap_b03r<-aov(Rate03~Site* B2015, Mcap_b03)
summary(Mcap_b03r)
TukeyHSD(Mcap_b03r)

#0-6
Mcap_b06<-subset(b06, Species=="Montipora capitata") #subset Mcap
Mcap_b06r<-aov(Rate06~Site * B2015, Mcap_b06)
summary(Mcap_b06r)
TukeyHSD(Mcap_b06r)
```

##*Porites compressa only*
`Bleaching score ~ Reef Site * Bleaching History`  
###Months 0-3, Rate change
Model: `score~Site*B2015`  
```{r}
#subset Pcomp
Pcomp_b03<-subset(b03, Species=="Porites compressa") #subset Pcomp
Pcomp_b03r<-aov(Rate03~Site  * B2015, Pcomp_b03)
summary(Pcomp_b03r)
TukeyHSD(Pcomp_b03r)

#0-6
Pcomp_b06<-subset(b06, Species=="Porites compressa") #subset Pcomp
Pcomp_b06r<-aov(Rate06~Site * B2015, Pcomp_b06)
summary(Pcomp_b06r)
TukeyHSD(Pcomp_b06r)

```


# Pairs 
(only include colonies when both partners of a pair are found)
```{r,results="hide"}
#Figure out which of the pairs we lose 
#subset only pairs
naData<-bALLData1
#dataset to work from (pull unwanted from)
pData<-bALLData1

# Identify which pairs have a partner not found
# Subset Month0 (two colonies missing values, checked datasheet)
M0<-subset(naData, month=="Month0") # subset month 0
M0_NA <- M0 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#keep it all

  # subset month 0
  M0a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M0<-pData[M0a]
  M0<-subset(M0, month=="Month0")
  M0<-na.omit(M0) # remove the NA colums
  #DIFFS
  M0diff<-M0  
  M0diff<-dcast(M0diff, Species+Site+Pair+month~B2015) # cast()
  M0diff$diff<-M0diff$N-M0diff$Y

# Subset Month1.5 
M1.5<-subset(naData, month=="Month1.5") # subset month 1.5 
M1.5_NA <- M1.5 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 113_114 remove
  # subset month 1.5
  M1.5a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M1.5<-pData[M1.5a]
  M1.5<-subset(M1.5, month=="Month1.5")
  M1.5<-na.omit(M1.5) # remove the NA colums
  M1.5<-M1.5[!(M1.5$Pair=="113_114"),]
 #DIFFS
  M1.5diff<-M1.5  
  M1.5diff<-dcast(M1.5diff, Species+Site+Pair+month~B2015) # cast()
  M1.5diff$diff<-M1.5diff$N-M1.5diff$Y
  
## Subset Month3
M3<-subset(naData, month=="Month3") # subset month 3
M3_NA <- M3 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 107_108, 13_14, 213_214, 215_216, 217_218, 285_286, 31_32 <- both not found
#7_8, 113_114,255_256, 259_260 <- Pairs where only one was not found :/

  # subset month 3
  M3a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M3<-pData[M3a]
  M3<-subset(M3, month=="Month3")
  M3<-na.omit(M3) # remove the NA colums
  # remove 7_8, 113_114,255_256, 259_260 <- Pairs where only one was not found :/
  M3<-M3[!(M3$Pair=="7_8"),]
  M3<-M3[!(M3$Pair=="113_114"),]
  M3<-M3[!(M3$Pair=="255_256"),]
  M3<-M3[!(M3$Pair=="259_260"),]
  
  #DIFFS
  M3diff<-M3  
  M3diff<-dcast(M3diff, Species+Site+Pair+month~B2015) # cast()
  M3diff$diff<-M3diff$N-M3diff$Y

## Subset Month 4.5
M4.5<-subset(naData, month=="Month4.5") # subset month 3
M4.5_NA <- M4.5 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 285_286,9_10, 209_210, 37)38, 255_256 <- both not found
# 7_8, 113_114,255_256, 259_260, 263_264, 267_268, 82_83 <- Pairs where only one was not found not 

  # subset month 4.5
  M4.5a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M4.5<-pData[M4.5a]
  M4.5<-subset(M4.5, month=="Month4.5")
  M4.5<-na.omit(M4.5) # remove the NA colums
  # remove # 7_8, 113_114,255_256, 259_260, 263_264, 267_268, 82_83
  M4.5<-M4.5[!(M4.5$Pair=="7_8"),]
  M4.5<-M4.5[!(M4.5$Pair=="113_114"),]
  M4.5<-M4.5[!(M4.5$Pair=="255_256"),]
  M4.5<-M4.5[!(M4.5$Pair=="259_260"),]
  M4.5<-M4.5[!(M4.5$Pair=="263_264"),]
  M4.5<-M4.5[!(M4.5$Pair=="267_268"),]
  M4.5<-M4.5[!(M4.5$Pair=="82_83"),]
 
  #DIFFS
  M4.5diff<-M4.5  
  M4.5diff<-dcast(M4.5diff, Species+Site+Pair+month~B2015) # cast()
  M4.5diff$diff<-M4.5diff$N-M4.5diff$Y
  

# Subset month 6
M6<-subset(naData, month=="Month6")
M6_NA <- M6 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#  113_114,13_14, 17_18, 213_214, 215_216, 217_218, 223_224,82_83
#<- 111_112, 115_116, 119_120, 207_208, 255_256, 259_260, 263_264,267_269 <-Pairs where only one was not found

  # Subset month 6
  M6a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M6<-pData[M6a]
  M6<-subset(M6, month=="Month6")
  M6<-na.omit(M6) # remove the NA colums
  # remove 111_112, 115_116, 119_120, 207_208, 255_256, 259_260, 263_264, 267_269 <-Pairs where only one was not found :/
  M6<-M6[!(M6$Pair=="111_112"),] # remove the other partner
  M6<-M6[!(M6$Pair=="115_116"),]
  M6<-M6[!(M6$Pair=="119_120"),]
  M6<-M6[!(M6$Pair=="207_208"),]
  M6<-M6[!(M6$Pair=="255_256"),]
  M6<-M6[!(M6$Pair=="259_260"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  M6<-M6[!(M6$Pair=="267_268"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  
  #DIFFS
  M6diff<-M6  
  M6diff<-dcast(M6diff, Species+Site+Pair+month~B2015) # cast()
  M6diff$diff<-M6diff$N-M6diff$Y

#make dfs  
pigPairs<-rbind(M0,M1.5,M3,M4.5,M6)
PairDat<-pigPairs
pigPairs06<-rbind(M0,M1.5,M3,M4.5,M6)
pigPairs03<-rbind(M0,M1.5,M3)
pigPairs36<-rbind(M3,M4.5,M6)

#Mean bleaching scores across the entire project

#calc average bleaching score per grouping by time point
pairmeans <- ddply(PairDat, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

pairmeans$code<-paste0(pairmeans$Site, pairmeans$B2015, pairmeans$Species)
pairmeans #display table
```

*Montipora capitata* 
```{r}
#subset for Mcap only
McapPairplot<-subset(pairmeans, Species=="Montipora capitata")


PairplotMcapB<-ggplot(data=McapPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.2) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Montipora capitata Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));PairplotMcapB
```

*Porites compressa* 
```{r}
#subset for Pcomp only
PcompPairplot<-subset(pairmeans, Species=="Porites compressa")

PairplotPcompB<-ggplot(data=PcompPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Porites compressa Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic"));PairplotPcompB
```

#Mcap only full model Pairs 0-6 months
SI Table 11
Model: `score~Site*B2015*Species`  
```{r}
#set up Pairs data
Mcappig<-subset(PairDat, Species =="Montipora capitata")            

#Month0
MpigPair0<-subset(Mcappig, month =="Month0")           
MpigPair0_mod1<-aov(score~Site*B2015, MpigPair0)
summary(MpigPair0_mod1)
TukeyHSD(MpigPair0_mod1) 

#Month1.5
MpigPair1.5<-subset(Mcappig, month =="Month1.5")
MpigPair1.5_mod1<-aov(score~Site*B2015, MpigPair1.5)
summary(MpigPair1.5_mod1)
TukeyHSD(MpigPair1.5_mod1)

#Month3
MpigPair3<-subset(Mcappig, month =="Month3")
MpigPair3_mod1<-aov(score~Site*B2015, MpigPair3)
summary(MpigPair3_mod1)
TukeyHSD(MpigPair3_mod1)

#Month4.5
MpigPair4.5<-subset(Mcappig, month =="Month4.5")
MpigPair4.5_mod1<-aov(score~Site*B2015, MpigPair4.5)
summary(MpigPair4.5_mod1)
TukeyHSD(MpigPair4.5_mod1)

#Month6
MpigPair6<-subset(Mcappig, month =="Month6")
# anova
MpigPair6_mod1<-aov(score~Site*B2015, MpigPair6)
summary(MpigPair6_mod1)
TukeyHSD(MpigPair6_mod1)
```

###Pcomp only full model Pairs 0-6 months
Model: `score~Site*B2015`  
SI Table 11
```{r}
Pcomppig<- subset(PairDat, Species == "Porites compressa") 

#Month 0
PpigPair0<-subset(Pcomppig, month =="Month0")            
PpigPair0_mod1<-aov(score~Site*B2015, PpigPair0)
summary(PpigPair0_mod1)
TukeyHSD(PpigPair0_mod1) #post hoc

#Month1.5
PpigPair1.5<-subset(Pcomppig, month =="Month1.5")
PpigPair1.5_mod1<-aov(score~Site*B2015, PpigPair1.5)
summary(PpigPair1.5_mod1)
TukeyHSD(PpigPair1.5_mod1)

#Month3
PpigPair3<-subset(Pcomppig, month =="Month3")
PpigPair3_mod1<-aov(score~Site*B2015, PpigPair3)
summary(PpigPair3_mod1)

#Month4.5
PpigPair4.5<-subset(Pcomppig, month =="Month4.5")
PpigPair4.5_mod1<-aov(score~Site*B2015, PpigPair4.5)
summary(PpigPair4.5_mod1)
TukeyHSD(PpigPair4.5_mod1)

#Month6
PpigPair6<-subset(Pcomppig, month =="Month6")
PpigPair6_mod1<-aov(score~Site*B2015, PpigPair6)
summary(PpigPair6_mod1)
```

0-3 and 0-6 month differences
Pairs data
score~month*B2015*lagoon
*M. capitata* Pairs
```{r}
Mcappig03Pairs<-subset(pigPairs03, Species=="Montipora capitata")
Mcappig06Pairs<-subset(pigPairs06, Species=="Montipora capitata")

#0-3
Mcappig03Pairs_mod<-aov(score~month*Site*B2015, Mcappig03Pairs)
summary(Mcappig03Pairs_mod)

#0-6
Mcappig06Pairs_mod<-aov(score~month*Site*B2015, Mcappig06Pairs)
summary(Mcappig06Pairs_mod)
```

Pcomp Pairs
```{r}
Pcomppig03Pairs<-subset(pigPairs03, Species=="Porites compressa")
Pcomppig06Pairs<-subset(pigPairs06, Species=="Porites compressa")
Pcomppig36Pairs<-subset(pigPairs36, Species=="Porites compressa")

#0-3
Pcomppig03Pairs_mod<-aov(score~month*Site*B2015, Pcomppig03Pairs)
summary(Pcomppig03Pairs_mod)
#06
Pcomppig06Pairs_mod<-aov(score~month*Site*B2015, Pcomppig06Pairs)
summary(Pcomppig06Pairs_mod)
```


# DIFF DATA
Differences (nonbleached-bleached) for pairs
```{r}
# data = combine all the M#diffs into a single dataframe 
DiffData06<-rbind(M0diff, M1.5diff, M3diff, M4.5diff, M6diff)
DiffData03<-rbind(M0diff, M1.5diff, M3diff)
DiffData36<-rbind(M3diff, M4.5diff, M6diff)
```

Mean bleaching scores across the entire project
```{r, results=TRUE}
#calc average bleaching score per grouping by time point
diffmeans <- ddply(DiffData06, c("Site", "Species", "month"), summarise, #
                   N    = length(diff[!is.na(diff)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(diff, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(diff, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(diff, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

diffmeans$code<-paste0(diffmeans$Site, diffmeans$B2015, diffmeans$Species)
diffmeans #display table
```

SI Figure 2
*Montipora capitata* diff 
```{r}
#subset for Mcap only
Mcapdiffplot<-subset(diffmeans, Species=="Montipora capitata")

Mcapdiffplota<-ggplot(data=Mcapdiffplot, aes(x=month, y=mean, group = code, color=Site)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=Site, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Difference score") +
    xlab("") + #Label the X Axis
    ylim(-.2, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Montipora capitata")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));Mcapdiffplota
Mcapdiffplota<-Mcapdiffplota + theme(legend.position = "none") #remove legend
Mcapdiffplota
```

###Porites compressa diff 
```{r}
#subset for Mcap only
Pcomdiffplot<-subset(diffmeans, Species=="Porites compressa")

Pcompdiffplota<-ggplot(data=Pcomdiffplot, aes(x=month, y=mean, group = code, color=Site)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=Site, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Difference score") +
    xlab("") + #Label the X Axis
    ylim(-.2, 2.5) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Porites compressa")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));Pcompdiffplota
Pcompdiffplota<-Pcompdiffplota + theme(legend.position = "none") #remove legend

```

SI Figure 2
```{r}
grid.arrange(Mcapdiffplota, Pcompdiffplota, nrow = 1) 
```

SI Table 1
Mcap diff~Lagoon
```{r}
DiffData03_Mcap<-subset(DiffData03, Species=="Montipora capitata")
DiffData06_Mcap<-subset(DiffData06, Species=="Montipora capitata")
DiffData36_Mcap<-subset(DiffData36, Species=="Montipora capitata")

#Month0
dMpigPair03<-subset(DiffData06_Mcap, month =="Month0")          
dMpigPair03_mod1<-aov(diff~Site, dMpigPair03)
summary(dMpigPair03_mod1)

#Month1.5
dMpigPair1.5<-subset(DiffData06_Mcap, month =="Month1.5")
dMpigPair1.5_mod1<-aov(diff~Site, dMpigPair1.5)
summary(dMpigPair1.5_mod1)

#Month3
dMpigPair3<-subset(DiffData06_Mcap, month =="Month3")
dMpigPair3_mod1<-aov(diff~Site, dMpigPair3)
summary(dMpigPair3_mod1)
TukeyHSD(dMpigPair3_mod1)

#Month4.5
dMpigPair4.5<-subset(DiffData06_Mcap, month =="Month4.5")
dMpigPair4.5_mod1<-aov(diff~Site, dMpigPair4.5)
summary(dMpigPair4.5_mod1)

#Month6
dMpigPair6<-subset(DiffData06_Mcap, month =="Month6")
dMpigPair6_mod1<-aov(diff~Site, dMpigPair6)
summary(dMpigPair6_mod1)
TukeyHSD(dMpigPair6_mod1)
```

diffs 0-3 months and 0-6 months
*M. capitata*
```{r}
# 0-3
d_Mcap03_mod1<-aov(diff~Site*month, DiffData03_Mcap)
summary(d_Mcap03_mod1)
TukeyHSD(d_Mcap03_mod1)

# 0-6
d_Mcap06_mod1<-aov(diff~Site*month, DiffData06_Mcap)
summary(d_Mcap06_mod1)
TukeyHSD(d_Mcap06_mod1)
```

*P. compressa*
```{r}
DiffData03_Pcomp<-subset(DiffData03, Species=="Porites compressa")
DiffData06_Pcomp<-subset(DiffData06, Species=="Porites compressa")

#Month0
dPpigPair03<-subset(DiffData03_Pcomp, month =="Month0")          
dPpigPair03_mod1<-aov(diff~Site, dPpigPair03)
summary(dPpigPair03_mod1)
TukeyHSD(dPpigPair03_mod1)

#Month1.5
dPpigPair1.5<-subset(DiffData06_Pcomp, month =="Month1.5")
dPpigPair1.5_mod1<-aov(diff~Site, dPpigPair1.5)
summary(dPpigPair1.5_mod1)

#Month3
dPpigPair3<-subset(DiffData06_Pcomp, month =="Month3")
dPpigPair3_mod1<-aov(diff~Site, dPpigPair3)
summary(dPpigPair3_mod1)

#Month4.5
dPpigPair4.5<-subset(DiffData06_Pcomp, month =="Month4.5")
dPpigPair4.5_mod1<-aov(diff~Site, dPpigPair4.5)
summary(dPpigPair4.5_mod1)

#Month6
dPpigPair6<-subset(DiffData06_Pcomp, month =="Month6")
dPpigPair6_mod1<-aov(diff~Site, dPpigPair6)
summary(dPpigPair6_mod1)
TukeyHSD(dPpigPair6_mod1)
```

diffs 0-3 months and 0-6 months
*P. compressa*
```{r}
# 0-3
d_Pcomp03_mod1<-aov(diff~Site*month, DiffData03_Pcomp)
summary(d_Pcomp03_mod1)
TukeyHSD(d_Pcomp03_mod1)

# 0-6
d_Pcomp06_mod1<-aov(diff~Site*month, DiffData06_Pcomp)
summary(d_Pcomp06_mod1)
TukeyHSD(d_Pcomp06_mod1)

# 3-6
d_Pcomp36_mod1<-aov(diff~Site*month, DiffData36_Pcomp)
summary(d_Pcomp36_mod1)
TukeyHSD(d_Pcomp36_mod1)
```
