---
title: "B_NB_stats"
author: "Shayle Matsuda"
date: "08/08/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
update:"03/31/3030"

Color score for *M. capitata* and *P. compressa* colonies at the Inner Lagoon (PR4) and Outer Lagoon (PR13) after the 2015 coral bleaching event in Kaneohe Bay.  Tagged individuals were given a color score based on their visual color: 1) white (>80% of colony white with no visible pigmentation); 2) pale (>10% colony affected by pigment loss); or 3) fully pigmented (<10% colony with any pale coloration). 

```{r setup, include=FALSE,results="hide"}
knitr::opts_chunk$set(echo = TRUE)
```

# Bleached/Non Bleached data 
```{r, results="hide", message=FALSE}
#Set up: load libraries

#rm(list=ls(all=TRUE)) 
library(phia)
library(emmeans)
library(ARTool)
library("reshape2") #reshape data
library("plotrix") #functions in tapply
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library(effects)
library("multcomp") 
library("multcompView")
library("coefplot") #coefficient plotting
library("seacarb") #seawater carbonate chemistry
library("vegan") #calculating distance matrices
library(betareg)
library(coin)
library(exactRankTests)
library(mvtnorm)
library(DTK)
library(PMCMR)
library(MuMIn)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(ggsignif)
#library(HH)
library(PMCMR)
library(tidyverse)
library(arm)
library(jtools)
library(broom)
library(ggstance)
library(officer)
library(huxtable)
library(flextable)
```

Load in data and organize 
```{r}
#We excluded these pairs as part of the pair analysis because the Bleached coral was only a 2: Porites Outer Bay 233_234, Porites Inner Bay 297_298, Porites Outer Bay 227_228, Montipora capitata Inner Bay 251_252.

# Bleaching data
bData<-read.csv("Bleaching_FINAL.csv") #read in bleaching df
bData$Site<-as.factor(bData$Site) #site as factor
bData$Pair<-as.factor(bData$Pair) #pair ID as factor

#organize data
bALL<-bData
  #melt
  bALLData1<-melt(data = bALL, measure.vars = c("Month0","Month1.5","Month3","Month4.5", "Month6","Month10","Month18","Month24"))   
  colnames(bALLData1)[colnames(bALLData1)=="variable"] <- "month"   # change col name from 'variable' to 'month'
  colnames(bALLData1)[colnames(bALLData1)=="value"] <- "score"   # change col name from 'value' to 'score'
  bALLData1$group<-paste(bALLData1$Species, bALLData1$Site, bALLData1$B2015, bALLData1$month) # add column for groups for figures only
#write.csv(bALL, "PigmentationBNB_20191122.csv")

#Remove all NON PAIRS and all NAs
  bALLData<-na.omit(bALLData1) # remove the NA rows (this removes individuals and NAs from pairs with no score at particular timepoints (due to not finding the coral))
  
  
  #Change month into TIME numeric 
  bALLData$time[bALLData$month=="Month0"]<-0
  bALLData$time[bALLData$month=="Month1.5"]<-1.5
  bALLData$time[bALLData$month=="Month3"]<-3
  bALLData$time[bALLData$month=="Month4.5"]<-4.5
  bALLData$time[bALLData$month=="Month6"]<-6
  bALLData$time[bALLData$month=="Month10"]<-10
  bALLData$time[bALLData$month=="Month18"]<-18
  bALLData$time[bALLData$month=="Month19"]<-19
  bALLData$time[bALLData$month=="Month24"]<-24

  
```

Subsets of df for Mcap and Pcomp
```{r} 
Mcap_df<-subset(bALLData, Species=="Montipora capitata")   #Mcap only
Pcomp_df<-subset(bALLData, Species=="Porites compressa")   #Pcomp only
``` 

Mean bleaching scores full df
```{r, results=TRUE}
#calc average bleaching  score per grouping by time point
ballmeans <- ddply(bALLData, c("Site", "Species", "B2015", "time"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
ballmeans #display table

ballmeans$code<-paste0(ballmeans$Site, ballmeans$B2015, ballmeans$Species)
ballmeans #display table
#write.csv(ballmeans,"ballmeans.csv"

```

#Figure 5
Bleaching score means 0-24 months by species. 
*Montipora capitata*  
Figure 5 
```{r}
#subset for Mcap only 
Mcap024Allplot<-subset(ballmeans, Species=="Montipora capitata")     #subset Mcap
 
#subset by reef
#PR13 
Mcap024AllplotPR13<-subset(Mcap024Allplot, Site==13)
#PR4
Mcap024AllplotPR4<-subset(Mcap024Allplot, Site==4)
```

*M. capitata* PR13

```{r}

ballplotMcapPR13<-ggplot(data=Mcap024AllplotPR13, aes(x=time, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold",hjust = 0.5));ballplotMcapPR13
McapOuter<-ballplotMcapPR13 + theme(legend.position = "none") #remove legend
```

*M. capitata* PR4
```{r}
ballplotMcapPR4<-ggplot(data=Mcap024AllplotPR4, aes(x=time, y=mean, group = code, color=B2015),show.legend = FALSE) + #color by bleaching hist. 
 geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Montipora capitata\nColor score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold", hjust = 0.5));ballplotMcapPR4
McapInner<-ballplotMcapPR4 + theme(legend.position = "none") #remove legend
McapInner
```

*Porites compressa*  
Figure 5
```{r}
Pcomp024Allplot<-subset(ballmeans, Species=="Porites compressa")

#subset by reef
Pcomp024AllplotPR13<-subset(Pcomp024Allplot, Site==13)
Pcomp024AllplotPR4<-subset(Pcomp024Allplot, Site==4)
```

*P. compress* 
Outer PR13
```{r}
b024plotPcompPR13<-ggplot(data=Pcomp024AllplotPR13, aes(x=time, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Porites compressa Bleaching Scores times 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR13

PcompOuter<-b024plotPcompPR13+ theme(legend.position = "none") #remove legend
```

*P. compressa*   
Inner PR4
```{r}
b024plotPcompPR4<-ggplot(data=Pcomp024AllplotPR4, aes(x=time, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Porites compressa\nColor score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Porites compressa Bleaching Scores times 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR4


PcompInner<-b024plotPcompPR4+ theme(legend.position = "none") #remove legend
PcompInner
```

Plot all four together, Figure 5
```{r}
grid.arrange(McapInner, McapOuter, PcompInner,PcompOuter, nrow = 2)
```




# Bleaching Recovery Models

#Table 2  
PIGMENTATION
lmer(score~month*B2015*lagoon+(1|TagID)
Mcap_df  full model, lmer
```{r}
#MONTH AS FACTOR
Mcap_df_mod<-lmer(score~B2015*Site*month+(1|TagID), data=Mcap_df)
summary(Mcap_df_mod)
anova(Mcap_df_mod) #pvals

#####qqPlot(residuals(Mcap_df_mod))    # DO WE INCLUDE RESIDUALS QQPLOTS, THEY SHOW WE ARE NOT NORMAL

#post hoc
test<-emmeans(Mcap_df_mod, "Site","B2015") #this right for showing bleached corals differeed btwn sites over entire experiment (aka across timepoints)?
cld(test, Letter="abcde")

 emm<-emmeans(Mcap_df_mod,  ~ B2015*Site*month, adjust="Tukey")
cld(emm, Letter="abcdefg")

test1a<-emmeans(Mcap_df_mod, "time")
cld(test1a, Letter="abcdefg")

## trying GLMER
#Mcap_df_mod_glmer<-glmer(score~B2015*Site*month+(1|TagID), family=poisson(link="log"), data=Mcap_df)
#Anova(Mcap_df_mod_glmer) #pvals use Anova()
  #conduct Tukey post hoc test 
 # emm = emmeans(Mcap_df_mod_glmer, ~ Site*B2015*month, adjust="tukey")
  #letter display
  #cld(emm, Letter="abcdefghijklmnopq")
```

Pcomp_df - full model lmer()
```{r}
#pcomp full model
Pcomp_df_mod2<-lmer(score~B2015*Site*month+(1|TagID), data=Pcomp_df)
summary(Pcomp_df_mod2)
anova(Pcomp_df_mod2) #pvals

#post hoc <- if you can use full do that!!!!!!!
emmeans(Pcomp_df_mod, "Site","B2015") #this right for showing bleached corals differeed btwn sites over entire experiment (aka across timepoints)?
cld(Letter="abcde")

emm<-emmeans(Pcomp_df_mod,  ~ B2015*Site*month) #this right for showing bleached corals differeed btwn sites over entire experiment (aka across timepoints)?
cld(emm)


## trying GLMER
#Pcomp_df_mod_glmer<-glmer(score~B2015*Site*month+(1|TagID), family=gaussian, data=Pcomp_df)
#Anova(Pcomp_df_mod_glmer) #pvals use Anova()

  #conduct Tukey post hoc test 
#  emm = emmeans(Pcomp_df_mod_glmer, ~ Site*B2015*month, adjust="tukey")
  #letter display
#  cld(emm)
```

Models at each timepoint -NEW
Mcap lm
```{r} 
Mcap_0<-subset(Mcap_df, month=="Month0")
Mcap_1.5<-subset(Mcap_df, month=="Month1.5")
Mcap_3<-subset(Mcap_df, month=="Month3")
Mcap_4.5<-subset(Mcap_df, month=="Month4.5")
Mcap_6<-subset(Mcap_df, month=="Month6")
Mcap_10<-subset(Mcap_df, month=="Month10")
Mcap_18<-subset(Mcap_df, month=="Month18")
Mcap_24<-subset(Mcap_df, month=="Month24")


#Month0
Month0_mcap0<-lm(score~Site*B2015, Mcap_0)
summary(Month0_mcap0)
anova(Month0_mcap0)

#Month1.5
Month1.5_mcap1.5<-lm(score~Site*B2015, Mcap_1.5)
summary(Month1.5_mcap1.5)
anova(Month1.5_mcap1.5)

#Month3
Month3_mcap3<-lm(score~Site*B2015, Mcap_3)
summary(Month3_mcap3)
anova(Month3_mcap3)

#Month4.5
Month4.5_mcap4.5<-lm(score~Site*B2015, Mcap_4.5)
summary(Month4.5_mcap4.5)

#Month6
Month6_mcap6<-lm(score~Site*B2015, Mcap_6)
summary(Month6_mcap6)
anova(Month6_mcap6)

#Month10
Month10_mcap10<-lm(score~Site*B2015, Mcap_10)
summary(Month10_mcap10)

#Month18
Month18_mcap18<-lm(score~Site*B2015, Mcap_18)
summary(Month18_mcap18)
anova(Month18_mcap18)

#Month24
Month24_mcap24t<-lm(score~Site*B2015, Mcap_24)
summary(Month24_mcap24t)
anova(Month24_mcap24t)
```

Mcap glm (by month) DELETE
```{r} 
Mcap_0<-subset(Mcap_df, month=="Month0")
Mcap_1.5<-subset(Mcap_df, month=="Month1.5")
Mcap_3<-subset(Mcap_df, month=="Month3")
Mcap_4.5<-subset(Mcap_df, month=="Month4.5")
Mcap_6<-subset(Mcap_df, month=="Month6")
Mcap_10<-subset(Mcap_df, month=="Month10")
Mcap_18<-subset(Mcap_df, month=="Month18")
Mcap_24<-subset(Mcap_df, month=="Month24")


#Month0
gMonth0_mcap0<-glm(score~Site*B2015, family=gaussian, Mcap_0)
Anova(gMonth0_mcap0)

#Month1.5
gMonth1.5_mcap1.5<-glm(score~Site*B2015, family=gaussian,Mcap_1.5)
Anova(gMonth1.5_mcap1.5)

#Month3
gMonth3_mcap3<-glm(score~Site*B2015,family=gaussian,  Mcap_3)
Anova(gMonth3_mcap3)

#Month4.5
gMonth4.5_mcap4.5<-glm(score~Site*B2015, family=gaussian, Mcap_4.5)
Anova(gMonth4.5_mcap4.5)

#Month6
gMonth6_mcap6<-glm(score~Site*B2015, family=gaussian, Mcap_6)
Anova(gMonth6_mcap6)

#Month10
gMonth10_mcap10<-glm(score~Site*B2015,family=gaussian,  Mcap_10)
Anova(gMonth10_mcap10)

#Month18
gMonth18_mcap18<-glm(score~Site*B2015, family=gaussian, Mcap_18)
Anova(gMonth18_mcap18)

#Month24
gMonth24_mcap24<-glm(score~Site*B2015, family=gaussian, Mcap_24)
Anova(gMonth24_mcap24)
```

Pcomp lm (by month)
```{r} 
Pcomp_0<-subset(Pcomp_df, month=="Month0")
Pcomp_1.5<-subset(Pcomp_df, month=="Month1.5")
Pcomp_3<-subset(Pcomp_df, month=="Month3")
Pcomp_4.5<-subset(Pcomp_df, month=="Month4.5")
Pcomp_6<-subset(Pcomp_df, month=="Month6")
Pcomp_10<-subset(Pcomp_df, month=="Month10")
Pcomp_18<-subset(Pcomp_df, month=="Month18")
Pcomp_24<-subset(Pcomp_df, month=="Month24")


#Month0
Month0_Pcomp0<-lm(score~Site*B2015, Pcomp_0)
summary(Month0_Pcomp0)
anova(Month0_Pcomp0)

#Month1.5
Month1.5_Pcomp1.5<-lm(score~Site*B2015, Pcomp_1.5)
summary(Month1.5_Pcomp1.5)
anova(Month1.5_Pcomp1.5)

#Month3
Month3_Pcomp3<-lm(score~Site*B2015, Pcomp_3)
summary(Month3_Pcomp3)
anova(Month3_Pcomp3)

#Month4.5
Month4.5_Pcomp4.5<-lm(score~Site*B2015, Pcomp_4.5)
summary(Month4.5_Pcomp4.5)
anova(Month4.5_Pcomp4.5)

#Month6
Month6_Pcomp6<-lm(score~Site*B2015,Pcomp_6)
summary(Month6_Pcomp6)
anova(Month6_Pcomp6)

#Month10
Month10_Pcomp10<-lm(score~Site*B2015, Pcomp_10)
summary(Month10_Pcomp10)
anova(Month10_Pcomp10)

#Month18
Month18_Pcomp18<-lm(score~Site*B2015,Pcomp_18)
summary(Month18_Pcomp18)
anova(Month18_Pcomp18)

#Month24
Month24_Pcomp24<-lm(score~Site*B2015,Pcomp_24)
summary(Month24_Pcomp24)
anova(Month24_Pcomp24)

```

# Pairs only bleaching recovery models
(subset of full dataset whi colonies only when both partners of a pair are found at each time point)  
Setup df.
```{r,results="hide"}
#Figure out which of the pairs we lose 
#subset only pairs
naData<-bALLData
#dataset to work from (pull unwanted from)
pData<-bALLData

# Identify which pairs have a partner not found
# Subset Month0 (two colonies missing values, checked datasheet)
M0<-subset(naData, month=="Month0") # subset month 0
#M0_NA <- M0 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#keep it all

  # subset month 0
  M0a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M0<-pData[M0a]
  M0<-subset(M0, month=="Month0")
  #M0<-na.omit(M0) # remove the NA colums
  #DIFFS
  M0diff<-M0  
  M0diff<-dcast(M0diff, Species+Site+Pair+month~B2015) # cast()
  M0diff$diff<-M0diff$N-M0diff$Y

# Subset Month1.5 
M1.5<-subset(naData, month=="Month1.5") # subset month 1.5 
#M1.5_NA <- M1.5 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 113_114 remove
  # subset month 1.5
  M1.5a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M1.5<-pData[M1.5a]
  M1.5<-subset(M1.5, month=="Month1.5")
  #M1.5<-na.omit(M1.5) # remove the NA colums
  M1.5<-M1.5[!(M1.5$Pair=="113_114"),]
 #DIFFS 
  M1.5diff<-M1.5  
  M1.5diff<-dcast(M1.5diff, Species+Site+Pair+month~B2015) # cast()
  M1.5diff$diff<-M1.5diff$N-M1.5diff$Y
  
## Subset Month3
M3<-subset(naData, month=="Month3") # subset month 3
#M3_NA <- M3 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 107_108, 13_14, 213_214, 215_216, 217_218, 285_286, 31_32 <- both not found
#7_8, 113_114,255_256, 259_260 <- Pairs where only one was not found :/

  # subset month 3
  M3a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M3<-pData[M3a]
  M3<-subset(M3, month=="Month3")
  #M3<-na.omit(M3) # remove the NA colums
  # remove 7_8, 113_114,255_256, 259_260 <- Pairs where only one was not found :/
  M3<-M3[!(M3$Pair=="7_8"),]
  M3<-M3[!(M3$Pair=="113_114"),]
  M3<-M3[!(M3$Pair=="255_256"),]
  M3<-M3[!(M3$Pair=="259_260"),]
  
  #DIFFS
  M3diff<-M3  
  M3diff<-dcast(M3diff, Species+Site+Pair+month~B2015) # cast()
  M3diff$diff<-M3diff$N-M3diff$Y

## Subset Month 4.5
M4.5<-subset(naData, month=="Month4.5") # subset month 3
#M4.5_NA <- M4.5 %>% filter_all(any_vars(is.na(.))) 
# 7_8, 113_114,255_256, 259_260, 263_264, 267_268, 82_83 <- Pairs where only one was not found not 

  # subset month 4.5
  M4.5a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M4.5<-pData[M4.5a]
  M4.5<-subset(M4.5, month=="Month4.5")
  M4.5<-na.omit(M4.5) # remove the NA colums
  # remove # 7_8, 113_114,255_256, 259_260, 263_264, 267_268, 82_83
  M4.5<-M4.5[!(M4.5$Pair=="7_8"),]
  M4.5<-M4.5[!(M4.5$Pair=="113_114"),]
  M4.5<-M4.5[!(M4.5$Pair=="255_256"),]
  M4.5<-M4.5[!(M4.5$Pair=="259_260"),]
  M4.5<-M4.5[!(M4.5$Pair=="263_264"),]
  M4.5<-M4.5[!(M4.5$Pair=="267_268"),]
  M4.5<-M4.5[!(M4.5$Pair=="82_83"),]
 
  #DIFFS
  M4.5diff<-M4.5  
  M4.5diff<-dcast(M4.5diff, Species+Site+Pair+month~B2015) # cast()
  M4.5diff$diff<-M4.5diff$N-M4.5diff$Y
  

# Subset month 6
M6<-subset(naData, month=="Month6")
#M6_NA <- M6 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#  113_114,13_14, 17_18, 213_214, 215_216, 217_218, 223_224,82_83
#<- 111_112, 115_116, 119_120, 207_208, 255_256, 259_260, 263_264,267_269 <-Pairs where only one was not found

  # Subset month 6
  M6a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M6<-pData[M6a]
  M6<-subset(M6, month=="Month6")
  #M6<-na.omit(M6) # remove the NA colums
  # remove 111_112, 115_116, 119_120, 207_208, 255_256, 259_260, 263_264, 267_269 <-Pairs where only one was not found :/
  M6<-M6[!(M6$Pair=="111_112"),] # remove the other partner
  M6<-M6[!(M6$Pair=="115_116"),]
  M6<-M6[!(M6$Pair=="119_120"),]
  M6<-M6[!(M6$Pair=="207_208"),]
  M6<-M6[!(M6$Pair=="255_256"),]
  M6<-M6[!(M6$Pair=="259_260"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  M6<-M6[!(M6$Pair=="267_268"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  
  #DIFFS
  M6diff<-M6  
  M6diff<-dcast(M6diff, Species+Site+Pair+month~B2015) # cast()
  M6diff$diff<-M6diff$N-M6diff$Y
  
  
# Subset month 10
M10<-subset(naData, month=="Month10")
#M10_NA <- M10 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#  113_114,13_14, 17_18, 213_214, 215_2110, 217_218, 223_224,82_83
#<- 111_112, 115_1110, 119_120, 207_208, 255_2510, 259_2100, 2103_2104,2107_2109 <-Pairs where only one was not found

  # Subset month 10
  M10a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M10<-pData[M10a]
  M10<-subset(M10, month=="Month10")
  #M10<-na.omit(M10) # remove the NA colums
  # remove 253_254, 259_260, 161_262, 263_264, 267_268, 271_272, 273_274, 275_276,277_278 , 31_2
  M10<-M10[!(M10$Pair=="253_254"),] # remove the other partner
  M10<-M10[!(M10$Pair=="259_260"),]
  M10<-M10[!(M10$Pair=="113_114"),]
  M10<-M10[!(M10$Pair=="161_262"),]
  M10<-M10[!(M10$Pair=="263_264"),]
  M10<-M10[!(M10$Pair=="267_268"),]
  M10<-M10[!(M10$Pair=="271_272"),]
  M10<-M10[!(M10$Pair=="273_274"),]
  M10<-M10[!(M10$Pair=="275_276"),]
  M10<-M10[!(M10$Pair=="277_278"),]
  M10<-M10[!(M10$Pair=="31_32"),]
    M10<-M10[!(M10$Pair=="261_262"),]


  
  #DIFFS
  M10diff<-M10  
  M10diff<-dcast(M10diff, Species+Site+Pair+month~B2015) # cast()
  M10diff$diff<-M10diff$N-M10diff$Y
  
  
  # Subset month 18
M18<-subset(naData, month=="Month18")
#M18_NA <- M18 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#<- 101_102,1-3_104, 115_116, 117_118, 119_120, 19_20, 205-206, 223_224, 261_262, 265_266, 269_270, 271_272, 275_276, 39_40

  # Subset month 18
  M18a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M18<-pData[M18a]
  M18<-subset(M18, month=="Month18")
  #M18<-na.omit(M18) # remove the NA colums
  M18<-M18[!(M18$Pair=="253_254"),] # remove the other partner
  M18<-M18[!(M18$Pair=="259_260"),]
  M18<-M18[!(M18$Pair=="161_262"),]
  M18<-M18[!(M18$Pair=="263_264"),]
  M18<-M18[!(M18$Pair=="267_268"),]
  M18<-M18[!(M18$Pair=="271_272"),]
  M18<-M18[!(M18$Pair=="273_274"),]
  M18<-M18[!(M18$Pair=="275_276"),]
  M18<-M18[!(M18$Pair=="277_278"),]
  M18<-M18[!(M18$Pair=="31_2"),]
    M18<-M18[!(M18$Pair=="115_116"),]
  M18<-M18[!(M18$Pair=="119_120"),]
    M18<-M18[!(M18$Pair=="261_262"),]
  M18<-M18[!(M18$Pair=="269_270"),]
  M18<-M18[!(M18$Pair=="101_102"),]
  M18<-M18[!(M18$Pair=="117_118"),]
    M18<-M18[!(M18$Pair=="265_266"),]
  M18<-M18[!(M18$Pair=="19_20"),]
  M18<-M18[!(M18$Pair=="205_206"),]
  M18<-M18[!(M18$Pair=="223_224"),]
    M18<-M18[!(M18$Pair=="39_40"),]
    
    
  #DIFFS
  M18diff<-M18  
  M18diff<-dcast(M18diff, Species+Site+Pair+month~B2015) # cast()
  M18diff$diff<-M18diff$N-M18diff$Y
  
   
  # Subset month 24
M24<-subset(naData, month=="Month24")


# Subset month 24
  M24a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M24<-pData[M24a]
  M24<-subset(M24, month=="Month24")
  M24<-M24[!(M24$Pair=="107_108"),] # remove the other partner
  M24<-M24[!(M24$Pair=="13_14"),]
  M24<-M24[!(M24$Pair=="223_224"),]
  M24<-M24[!(M24$Pair=="231_232"),]
  M24<-M24[!(M24$Pair=="261_262"),]
  M24<-M24[!(M24$Pair=="263_264"),]
  M24<-M24[!(M24$Pair=="265_266"),]
  M24<-M24[!(M24$Pair=="283_284"),]
  M24<-M24[!(M24$Pair=="289_290"),]
  M24<-M24[!(M24$Pair=="31_32"),]
  M24<-M24[!(M24$Pair=="33_34"),]
  M24<-M24[!(M24$Pair=="39_40"),]
  M24<-M24[!(M24$Pair=="76_77"),]
  M24<-M24[!(M24$Pair=="78_79"),]

  #DIFFS
  M24diff<-M24  
  M24diff<-dcast(M24diff, Species+Site+Pair+month~B2015) # cast()
  M24diff$diff<-M24diff$N-M24diff$Y
  
#make dfs   
PairDat<-rbind(M0,M1.5,M3,M4.5,M6,M10,M18,M24)
 
#Mean bleaching scores across the entire project

#calc average bleaching score per grouping by time point
pairmeans <- ddply(PairDat, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

pairmeans$code<-paste0(pairmeans$Site, pairmeans$B2015, pairmeans$Species)
pairmeans #display table
```

*Montipora capitata*  plot pairs
```{r}
#subset for Mcap only
McapPairplot<-subset(pairmeans, Species=="Montipora capitata")


PairplotMcapB<-ggplot(data=McapPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.2) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Montipora capitata Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));PairplotMcapB
```

*Porites compressa* plot pairs
```{r}
#subset for Pcomp only 
PcompPairplot<-subset(pairmeans, Species=="Porites compressa")

PairplotPcompB<-ggplot(data=PcompPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Porites compressa Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic"));PairplotPcompB
```
#Mcap: Pairs 
SI Table 1 
Model: glmer(score~Site*B2015*Species+(1|Pair)+(1|TagID))
LMER/lm
```{r}
#set up Pairs data
Mcappig<-subset(PairDat, Species =="Montipora capitata")            

#All lmer
MpigPair_all_mod1<-lmer(score~Site*B2015*month+(1|TagID)+(1|Pair), Mcappig)
summary(MpigPair_all_mod1)
anova(MpigPair_all_mod1)
 
#Month0
MpigPair0<-subset(Mcappig, month =="Month0")           
MpigPair0_mod1<-lm(score~Site*B2015, MpigPair0)
summary(MpigPair0_mod1)
anova(MpigPair0_mod1)

#Month1.5
MpigPair1.5<-subset(Mcappig, month =="Month1.5")
MpigPair1.5_mod1<-lm(score~Site*B2015, MpigPair1.5)
summary(MpigPair1.5_mod1)
anova(MpigPair1.5_mod1)

#Month3
MpigPair3<-subset(Mcappig, month =="Month3")
MpigPair3_mod1<-lm(score~Site*B2015, MpigPair3)
summary(MpigPair3_mod1)
anova(MpigPair3_mod1)

#Month4.5
MpigPair4.5<-subset(Mcappig, month =="Month4.5")
MpigPair4.5_mod1<-lm(score~Site*B2015, MpigPair4.5)
summary(MpigPair4.5_mod1)
anova(MpigPair4.5_mod1)

#Month6
MpigPair6<-subset(Mcappig, month =="Month6")
MpigPair6_mod1<-lm(score~Site*B2015, MpigPair6)
summary(MpigPair6_mod1)
anova(MpigPair6_mod1)

#Month10
MpigPair10<-subset(Mcappig, month =="Month10")
MpigPair10_mod1<-lm(score~Site*B2015, MpigPair10)
summary(MpigPair10_mod1)
anova(MpigPair10_mod1)

#Month18
MpigPair18<-subset(Mcappig, month =="Month18")
MpigPair18_mod1<-lm(score~Site*B2015, MpigPair18)
summary(MpigPair18_mod1)
anova(MpigPair18_mod1)

#Month24
MpigPair24<-subset(Mcappig, month =="Month24")
MpigPair24_mod1<-lm(score~Site*B2015, MpigPair24)
summary(MpigPair24_mod1)
anova(MpigPair24_mod1)
```
GLMER/glm DELETE
```{r}
           
#All glmer
gMpigPair_all_mod1<-glmer(score~Site*B2015*month+(1|TagID)+(1|Pair), family=gaussian, Mcappig)
summary(gMpigPair_all_mod1)
Anova(gMpigPair_all_mod1)
 
#Month0
gMpigPair0<-subset(Mcappig, month =="Month0")           
gMpigPair0_mod1<-glm(score~Site*B2015, family=gaussian,gMpigPair0)
summary(gMpigPair0_mod1)
Anova(gMpigPair0_mod1)

#Month1.5
gMpigPair1.5<-subset(Mcappig, month =="Month1.5")
gMpigPair1.5_mod1<-glm(score~Site*B2015, family=gaussian,gMpigPair1.5)
summary(gMpigPair1.5_mod1)
Anova(gMpigPair1.5_mod1)

#Month3
gMpigPair3<-subset(Mcappig, month =="Month3")
gMpigPair3_mod1<-glm(score~Site*B2015, family=gaussian,gMpigPair3)
summary(gMpigPair3_mod1)
Anova(gMpigPair3_mod1)

#Month4.5
gMpigPair4.5<-subset(Mcappig, month =="Month4.5")
gMpigPair4.5_mod1<-glm(score~Site*B2015,family=gaussian, gMpigPair4.5)
summary(gMpigPair4.5_mod1)
Anova(gMpigPair4.5_mod1)

#Month6
gMpigPair6<-subset(Mcappig, month =="Month6")
gMpigPair6_mod1<-glm(score~Site*B2015,family=gaussian, gMpigPair6)
summary(gMpigPair6_mod1)
Anova(gMpigPair6_mod1)

#Month10
gMpigPair10<-subset(Mcappig, month =="Month10")
gMpigPair10_mod1<-glm(score~Site*B2015,family=gaussian, gMpigPair10)
summary(gMpigPair10_mod1)
Anova(gMpigPair10_mod1)

#Month18
gMpigPair18<-subset(Mcappig, month =="Month18")
gMpigPair18_mod1<-glm(score~Site*B2015, family=gaussian,gMpigPair18)
summary(gMpigPair18_mod1)
Anova(gMpigPair18_mod1)

#Month24
gMpigPair24<-subset(Mcappig, month =="Month24")
gMpigPair24_mod1<-glm(score~Site*B2015,family=gaussian, gMpigPair24)
summary(gMpigPair24_mod1)
Anova(gMpigPair24_mod1)
```

#Pcomp Pairs
LMER
```{r}
Pcomppig<- subset(PairDat, Species == "Porites compressa") 

#All timepoints lmer
PpigPair_all_mod1<-lmer(score~Site*B2015*month+(1|TagID)+(1|Pair), Pcomppig)
summary(PpigPair_all_mod1)
anova(PpigPair_all_mod1)


#Month 0
PpigPair0<-subset(Pcomppig, month =="Month0")            
PpigPair0_mod1<-lm(score~Site*B2015, PpigPair0)
summary(PpigPair0_mod1)
anova(PpigPair0_mod1) 

#Month1.5
PpigPair1.5<-subset(Pcomppig, month =="Month1.5")
PpigPair1.5_mod1<-lm(score~Site*B2015, PpigPair1.5)
summary(PpigPair1.5_mod1)
anova(PpigPair1.5_mod1)

#Month3
PpigPair3<-subset(Pcomppig, month =="Month3")
PpigPair3_mod1<-lm(score~Site*B2015, PpigPair3)
anova(PpigPair3_mod1)

#Month4.5
PpigPair4.5<-subset(Pcomppig, month =="Month4.5")
PpigPair4.5_mod1<-lm(score~Site*B2015, PpigPair4.5)
summary(PpigPair4.5_mod1)
anova(PpigPair4.5_mod1)

#Month6
PpigPair6<-subset(Pcomppig, month =="Month6")
PpigPair6_mod1<-lm(score~Site*B2015, PpigPair6)
anova(PpigPair6_mod1)

#Month10
PpigPair10<-subset(Pcomppig, month =="Month10")
PpigPair10_mod1<-lm(score~Site*B2015, PpigPair10)
anova(PpigPair10_mod1)

#Month18
PpigPair18<-subset(Pcomppig, month =="Month18")
PpigPair18_mod1<-lm(score~Site*B2015, PpigPair18)
anova(PpigPair18_mod1)

#Month24
PpigPair24<-subset(Pcomppig, month =="Month24")
PpigPair24_mod1<-lm(score~Site*B2015, PpigPair24)
anova(PpigPair24_mod1)
```
GLMER Delete
```{r}
           
#All glmer
gPpigPair_all_mod1<-glmer(score~Site*B2015*month+(1|TagID)+(1|Pair), family=gaussian, Pcomppig)
summary(gPpigPair_all_mod1)
Anova(gPpigPair_all_mod1)
 
#Month0
gPpigPair0<-subset(Pcomppig, month =="Month0")           
gPpigPair0_mod1<-glm(score~Site*B2015, family=gaussian,gPpigPair0)
summary(gPpigPair0_mod1)
Anova(gPpigPair0_mod1)

#Month1.5
gPpigPair1.5<-subset(Pcomppig, month =="Month1.5")
gPpigPair1.5_mod1<-glm(score~Site*B2015, family=gaussian,gPpigPair1.5)
summary(gPpigPair1.5_mod1)
Anova(gPpigPair1.5_mod1)

#Month3
gPpigPair3<-subset(Pcomppig, month =="Month3")
gPpigPair3_mod1<-glm(score~Site*B2015, family=gaussian,gPpigPair3)
summary(gPpigPair3_mod1)
Anova(gPpigPair3_mod1)

#Month4.5
gPpigPair4.5<-subset(Pcomppig, month =="Month4.5")
gPpigPair4.5_mod1<-glm(score~Site*B2015,family=gaussian, gPpigPair4.5)
summary(gPpigPair4.5_mod1)
Anova(gPpigPair4.5_mod1)

#Month6
gPpigPair6<-subset(Pcomppig, month =="Month6")
gPpigPair6_mod1<-glm(score~Site*B2015,family=gaussian, gPpigPair6)
summary(gPpigPair6_mod1)
Anova(gPpigPair6_mod1)

#Month10
gPpigPair10<-subset(Pcomppig, month =="Month10")
gPpigPair10_mod1<-glm(score~Site*B2015,family=gaussian, gPpigPair10)
summary(gPpigPair10_mod1)
Anova(gPpigPair10_mod1)

#Month18
gPpigPair18<-subset(Pcomppig, month =="Month18")
gPpigPair18_mod1<-glm(score~Site*B2015, family=gaussian,gPpigPair18)
summary(gPpigPair18_mod1)
Anova(gPpigPair18_mod1)

#Month24
gPpigPair24<-subset(Pcomppig, month =="Month24")
gPpigPair24_mod1<-glm(score~Site*B2015,family=gaussian, gPpigPair24)
summary(gPpigPair24_mod1)
Anova(gPpigPair24_mod1)
```


# DIFF DATA
Differences (nonbleached-bleached) for pairs over time

```{r}
# data = combine all the M#diffs into a single dataframe
All_diff<-rbind(M0diff, M1.5diff, M3diff, M4.5diff, M6diff,M10diff, M18diff, M24diff)
```
 
Mean bleaching scores across the entire project
```{r, results=TRUE}
#calc average bleaching score per grouping by time point
diffmeans <- ddply(All_diff, c("Site", "Species", "month"), summarise, #
                   N    = length(diff[!is.na(diff)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(diff, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(diff, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(diff, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

diffmeans$code<-paste0(diffmeans$Site, diffmeans$B2015, diffmeans$Species)
diffmeans #display table
```

SI Figure 2
*Montipora capitata* diff plot
```{r}
#subset for Mcap only 
Mcapdiffplot<-subset(diffmeans, Species=="Montipora capitata")

Mcapdiffplota<-ggplot(data=Mcapdiffplot, aes(x=month, y=mean, group = code, color=Site)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=Site, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Difference score") +
    xlab("") + #Label the X Axis
    ylim(-.2, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Montipora capitata")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));Mcapdiffplota
Mcapdiffplota<-Mcapdiffplota + theme(legend.position = "none") #remove legend
Mcapdiffplota
```

*Porites compressa* diff plot
```{r}
#subset for Mcap only
Pcomdiffplot<-subset(diffmeans, Species=="Porites compressa")

Pcompdiffplota<-ggplot(data=Pcomdiffplot, aes(x=month, y=mean, group = code, color=Site)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=Site, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Difference score") +
    xlab("") + #Label the X Axis
    ylim(-.2, 2.5) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Porites compressa")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));Pcompdiffplota
Pcompdiffplota<-Pcompdiffplota + theme(legend.position = "none") #remove legend

```

SI Figure 2 combined
```{r}
grid.arrange(Mcapdiffplota, Pcompdiffplota, nrow = 1) 
```

SI Table 1-full mod, difference data only
Mcap diff
lmer
```{r} 
DiffData_all_Mcap<-subset(All_diff, Species=="Montipora capitata")

#Month ALL
dMpigPairall_mod1<-lmer(diff~Site*month+(1|Pair), DiffData_all_Mcap)
summary(dMpigPairall_mod1)
anova(dMpigPairall_mod1)

#Month0 
dMpigPair03<-subset(DiffData_all_Mcap, month =="Month0")          
dMpigPair03_mod1<-lm(diff~Site, dMpigPair03)
summary(dMpigPair03_mod1)
anova(dMpigPair03_mod1)

#Month1.5
dMpigPair1.5<-subset(DiffData_all_Mcap, month =="Month1.5")
dMpigPair1.5_mod1<-lm(diff~Site, dMpigPair1.5)
summary(dMpigPair1.5_mod1)
anova(dMpigPair1.5_mod1)

#Month3
dMpigPair3<-subset(DiffData_all_Mcap, month =="Month3")
dMpigPair3_mod1<-lm(diff~Site, dMpigPair3)
summary(dMpigPair3_mod1)
anova(dMpigPair3_mod1)

#Month4.5
dMpigPair4.5<-subset(DiffData_all_Mcap, month =="Month4.5")
dMpigPair4.5_mod1<-lm(diff~Site, dMpigPair4.5)
anova(dMpigPair4.5_mod1)

#Month6
dMpigPair6<-subset(DiffData_all_Mcap, month =="Month6")
dMpigPair6_mod1<-lm(diff~Site, dMpigPair6)
summary(dMpigPair6_mod1)
anova(dMpigPair6_mod1)

#Month10
dMpigPair10<-subset(DiffData_all_Mcap, month =="Month10")
dMpigPair10_mod1<-lm(diff~Site, dMpigPair10)
summary(dMpigPair10_mod1)
anova(dMpigPair10_mod1)

#Month18
dMpigPair18<-subset(DiffData_all_Mcap, month =="Month18")
dMpigPair18_mod1<-lm(diff~Site, dMpigPair18)
summary(dMpigPair18_mod1)
anova(dMpigPair18_mod1)

#Month24
dMpigPair24<-subset(DiffData_all_Mcap, month =="Month24")
dMpigPair24_mod1<-lm(diff~Site, dMpigPair24)
summary(dMpigPair24_mod1)
anova(dMpigPair24_mod1)
```
Mcap glmer DELETE
```{r} 
DiffData_all_Mcap<-subset(All_diff, Species=="Montipora capitata")

#Month ALL
gdMpigPairall_mod1<-glmer(diff~Site*month+(1|Pair), family=gaussian, DiffData_all_Mcap)
summary(gdMpigPairall_mod1)
Anova(gdMpigPairall_mod1)

#Month0 
gdMpigPair03<-subset(DiffData_all_Mcap, month =="Month0")          
gdMpigPair03_mod1<-glm(diff~Site, family=gaussian,gdMpigPair03)
summary(gdMpigPair03_mod1)
Anova(gdMpigPair03_mod1)

#Month1.5
gdMpigPair1.5<-subset(DiffData_all_Mcap, month =="Month1.5")
gdMpigPair1.5_mod1<-glm(diff~Site, family=gaussian,gdMpigPair1.5)
summary(gdMpigPair1.5_mod1)
Anova(gdMpigPair1.5_mod1)

#Month3
gdMpigPair3<-subset(DiffData_all_Mcap, month =="Month3")
gdMpigPair3_mod1<-glm(diff~Site, family=gaussian,gdMpigPair3)
summary(gdMpigPair3_mod1)
Anova(gdMpigPair3_mod1)

#Month4.5
gdMpigPair4.5<-subset(DiffData_all_Mcap, month =="Month4.5")
gdMpigPair4.5_mod1<-glm(diff~Site, family=gaussian,gdMpigPair4.5)
summary(gdMpigPair4.5_mod1)
Anova(gdMpigPair4.5_mod1)

#Month6
gdMpigPair6<-subset(DiffData_all_Mcap, month =="Month6")
gdMpigPair6_mod1<-glm(diff~Site, family=gaussian,gdMpigPair6)
summary(gdMpigPair6_mod1)
Anova(gdMpigPair6_mod1)

#Month10
gdMpigPair10<-subset(DiffData_all_Mcap, month =="Month10")
gdMpigPair10_mod1<-glm(diff~Site,family=gaussian, gdMpigPair10)
summary(gdMpigPair10_mod1)
Anova(gdMpigPair10_mod1)

#Month18
gdMpigPair18<-subset(DiffData_all_Mcap, month =="Month18")
gdMpigPair18_mod1<-glm(diff~Site, family=gaussian,gdMpigPair18)
summary(gdMpigPair18_mod1)
Anova(gdMpigPair18_mod1)

#Month24
gdMpigPair24<-subset(DiffData_all_Mcap, month =="Month24")
gdMpigPair24_mod1<-glm(diff~Site, family=gaussian,gdMpigPair24)
summary(gdMpigPair24_mod1)
Anova(gdMpigPair24_mod1)
```

Pcomp diff
lmer
```{r} 
DiffData_all_Pcomp<-subset(All_diff, Species=="Porites capitata")

#Month ALL
dPpigPairall_mod1<-lmer(diff~Site*month+(1|Pair), DiffData_all_Pcomp)
summary(dPpigPairall_mod1)
anova(dPpigPairall_mod1)

#Month0 
dPpigPair03<-subset(DiffData_all_Pcomp, month =="Month0")          
dPpigPair03_mod1<-lm(diff~Site, dPpigPair03)
summary(dPpigPair03_mod1)
anova(dPpigPair03_mod1)

#Month1.5
dPpigPair1.5<-subset(DiffData_all_Pcomp, month =="Month1.5")
dPpigPair1.5_mod1<-lm(diff~Site, dPpigPair1.5)
summary(dPpigPair1.5_mod1)
anova(dPpigPair1.5_mod1)

#Month3
dPpigPair3<-subset(DiffData_all_Pcomp, month =="Month3")
dPpigPair3_mod1<-lm(diff~Site, dPpigPair3)
summary(dPpigPair3_mod1)
anova(dPpigPair3_mod1)

#Month4.5
dPpigPair4.5<-subset(DiffData_all_Pcomp, month =="Month4.5")
dPpigPair4.5_mod1<-lm(diff~Site, dPpigPair4.5)
summary(dPpigPair4.5_mod1)
anova(dPpigPair4.5_mod1)

#Month6
dPpigPair6<-subset(DiffData_all_Pcomp, month =="Month6")
dPpigPair6_mod1<-lm(diff~Site, dPpigPair6)
summary(dPpigPair6_mod1)
anova(dPpigPair6_mod1)

#Month10
dPpigPair10<-subset(DiffData_all_Pcomp, month =="Month10")
dPpigPair10_mod1<-lm(diff~Site, dPpigPair10)
summary(dPpigPair10_mod1)
anova(dPpigPair10_mod1)

#Month18
dPpigPair18<-subset(DiffData_all_Pcomp, month =="Month18")
dPpigPair18_mod1<-lm(diff~Site, dPpigPair18)
summary(dPpigPair18_mod1)
anova(dPpigPair18_mod1)

#Month24
dPpigPair24<-subset(DiffData_all_Pcomp, month =="Month24")
dPpigPair24_mod1<-lm(diff~Site, dPpigPair24)
summary(dPpigPair24_mod1)
anova(dPpigPair24_mod1)
```
Pcomp glmer dELTE
```{r} 
DiffData_all_Pcomp<-subset(All_diff, Species=="Montipora capitata")

#Month ALL
gdPpigPairall_mod1<-glmer(diff~Site*month+(1|Pair), family=gaussian, DiffData_all_Pcomp)
summary(gdPpigPairall_mod1)
Anova(gdPpigPairall_mod1)

#Month0 
gdPpigPair03<-subset(DiffData_all_Pcomp, month =="Month0")          
gdPpigPair03_mod1<-glm(diff~Site, family=gaussian,gdPpigPair03)
summary(gdPpigPair03_mod1)
Anova(gdPpigPair03_mod1)

#Month1.5
gdPpigPair1.5<-subset(DiffData_all_Pcomp, month =="Month1.5")
gdPpigPair1.5_mod1<-glm(diff~Site, family=gaussian,gdPpigPair1.5)
summary(gdPpigPair1.5_mod1)
Anova(gdPpigPair1.5_mod1)

#Month3
gdPpigPair3<-subset(DiffData_all_Pcomp, month =="Month3")
gdPpigPair3_mod1<-glm(diff~Site, family=gaussian,gdPpigPair3)
summary(gdPpigPair3_mod1)
Anova(gdPpigPair3_mod1)

#Month4.5
gdPpigPair4.5<-subset(DiffData_all_Pcomp, month =="Month4.5")
gdPpigPair4.5_mod1<-glm(diff~Site, family=gaussian,gdPpigPair4.5)
summary(gdPpigPair4.5_mod1)
Anova(gdPpigPair4.5_mod1)

#Month6
gdPpigPair6<-subset(DiffData_all_Pcomp, month =="Month6")
gdPpigPair6_mod1<-glm(diff~Site, family=gaussian,gdPpigPair6)
summary(gdPpigPair6_mod1)
Anova(gdPpigPair6_mod1)

#Month10
gdPpigPair10<-subset(DiffData_all_Pcomp, month =="Month10")
gdPpigPair10_mod1<-glm(diff~Site,family=gaussian, gdPpigPair10)
summary(gdPpigPair10_mod1)
Anova(gdPpigPair10_mod1)

#Month18
gdPpigPair18<-subset(DiffData_all_Pcomp, month =="Month18")
gdPpigPair18_mod1<-glm(diff~Site, family=gaussian,gdPpigPair18)
summary(gdPpigPair18_mod1)
Anova(gdPpigPair18_mod1)

#Month24
gdPpigPair24<-subset(DiffData_all_Pcomp, month =="Month24")
gdPpigPair24_mod1<-glm(diff~Site, family=gaussian,gdPpigPair24)
summary(gdPpigPair24_mod1)
Anova(gdPpigPair24_mod1)
```







DELETE below. 





ARTools diffs DELETE
Mcap
```{r}
DiffData_all_Mcap<-subset(All_diff, Species=="Montipora capitata")
DiffData_all_Mcap<-na.omit(DiffData_all_Mcap)
#Month ALL
DiffdMpigPairall_mod1<-art(diff~Site*month+(1|Pair), DiffData_all_Mcap)
summary(DiffdMpigPairall_mod1) #Max is 4.8, violation
anova(DiffdMpigPairall_mod1)

#Month0
DiffdMpigPair03<-subset(DiffData_all_Mcap, month =="Month0")          
DiffdMpigPair03_mod1<-art(diff~Site, DiffdMpigPair03)
summary(DiffdMpigPair03_mod1)
anova(DiffdMpigPair03_mod1)


#Month1.5
DiffdMpigPair1.5<-subset(DiffData_all_Mcap, month =="Month1.5")
DiffdMpigPair1.5_mod1<-art(diff~Site, DiffdMpigPair1.5)
summary(DiffdMpigPair1.5_mod1)
anova(DiffdMpigPair1.5_mod1)

#Month3
DiffdMpigPair3<-subset(DiffData_all_Mcap, month =="Month3")
DiffdMpigPair3_mod1<-art(diff~Site, DiffdMpigPair3)
summary(DiffdMpigPair3_mod1)
anova(DiffdMpigPair3_mod1)

#Month4.5
DiffdMpigPair4.5<-subset(DiffData_all_Mcap, month =="Month4.5")
DiffdMpigPair4.5_mod1<-art(diff~Site, DiffdMpigPair4.5)
summary(DiffdMpigPair4.5_mod1)
anova(DiffdMpigPair4.5_mod1)

#Month6
DiffdMpigPair6<-subset(DiffData_all_Mcap, month =="Month6")
DiffdMpigPair6_mod1<-art(diff~Site, DiffdMpigPair6)
summary(DiffdMpigPair6_mod1)
anova(DiffdMpigPair6_mod1)

#Month10
DiffdMpigPair10<-subset(DiffData_all_Mcap, month =="Month10")
DiffdMpigPair10_mod1<-art(diff~Site, DiffdMpigPair10)
summary(DiffdMpigPair10_mod1)
anova(DiffdMpigPair10_mod1)

#Month18
DiffdMpigPair18<-subset(DiffData_all_Mcap, month =="Month18")
DiffdMpigPair18_mod1<-art(diff~Site, DiffdMpigPair18)
summary(DiffdMpigPair18_mod1)
anova(DiffdMpigPair18_mod1)

#Month24
DiffdMpigPair24<-subset(DiffData_all_Mcap, month =="Month24")
DiffdMpigPair24_mod1<-art(diff~Site, DiffdMpigPair24)
summary(DiffdMpigPair24_mod1)
anova(DiffdMpigPair24_mod1)
```

Pcomp diff ARTools (CODE not edited yet) Delete
```{r}
DiffData_all_Pcomp<-subset(All_diff, Species=="Porites compressa")
DiffData_all_Pcomp<-na.omit(DiffData_all_Pcomp)
#Month ALL
DiffdPpigPairall_mod1<-art(diff~Site*month+(1|Pair), DiffData_all_Pcomp)
summary(DiffdPpigPairall_mod1) #violation
anova(DiffdPpigPairall_mod1)

#Month0
DiffdPpigPair03<-subset(DiffData_all_Pcomp, month =="Month0")          
DiffdPpigPair03_mod1<-aov(diff~Site, DiffdPpigPair03)
summary(DiffdPpigPair03_mod1)

#Month1.5
DiffdPpigPair1.5<-subset(DiffData_all_Pcomp, month =="Month1.5")
DiffdPpigPair1.5_mod1<-aov(diff~Site, DiffdPpigPair1.5)
summary(DiffdPpigPair1.5_mod1)

#Month3
DiffdPpigPair3<-subset(DiffData_all_Pcomp, month =="Month3")
DiffdPpigPair3_mod1<-aov(diff~Site, DiffdPpigPair3)
summary(DiffdPpigPair3_mod1)

#Month4.5
DiffdPpigPair4.5<-subset(DiffData_all_Pcomp, month =="Month4.5")
DiffdPpigPair4.5_mod1<-aov(diff~Site, DiffdPpigPair4.5)
summary(DiffdPpigPair4.5_mod1)

#Month6
DiffdPpigPair6<-subset(DiffData_all_Pcomp, month =="Month6")
DiffdPpigPair6_mod1<-aov(diff~Site, DiffdPpigPair6)
summary(DiffdPpigPair6_mod1)

#Month10
DiffdPpigPair10<-subset(DiffData_all_Pcomp, month =="Month10")
DiffdPpigPair10_mod1<-aov(diff~Site, DiffdPpigPair10)
summary(DiffdPpigPair10_mod1)

#Month18
DiffdPpigPair18<-subset(DiffData_all_Pcomp, month =="Month18")
DiffdPpigPair18_mod1<-aov(diff~Site, DiffdPpigPair18)
summary(DiffdPpigPair18_mod1)

#Month24
DiffdPpigPair24<-subset(DiffData_all_Pcomp, month =="Month24")
DiffdPpigPair24_mod1<-aov(diff~Site, DiffdPpigPair24)
summary(DiffdPpigPair24_mod1)
```



diffs 0-3 months and 0-6 months
*P. compressa*
```{r}
# 0-3
d_Pcomp03_mod1<-aov(diff~Site*month, DiffData03_Pcomp)
summary(d_Pcomp03_mod1)
TukeyHSD(d_Pcomp03_mod1)

# 0-6
d_Pcomp06_mod1<-aov(diff~Site*month, DiffData06_Pcomp)
summary(d_Pcomp06_mod1)
TukeyHSD(d_Pcomp06_mod1)

# 3-6
d_Pcomp36_mod1<-aov(diff~Site*month, DiffData36_Pcomp)
summary(d_Pcomp36_mod1)
TukeyHSD(d_Pcomp36_mod1)
```







ARTools, keep?
# ARTools stats for non-parametric non-normality  
Full model
```{r} 
# replace Site with PR4 and PR13


#Full timeframe Mcap (log, arcsinsqrt don't help normailty)

hist(Mcap_df$sqrt) #not normal,use ARTools for non-parametric
Mcap_df$sqrt<- arcsin(sqrt(Mcap_df$score))
  
Mcap_df<-Mcap_df[,c(1,2,4:7)] #subset and order cols for ARTool

Full_mod_mcap<-art(score~month*Site*B2015+(1|TagID), Mcap_df)  #using TagID as only random factor not PAIR
summary(Full_mod_mcap) #all should be ~0 and are
anova(Full_mod_mcap) # Everything is significant

#Analysis of Variance of Aligned Rank Transformed Data

    #Table Type: Analysis of Deviance Table (Type III Wald F tests with Kenward-Roger df) 
    #Model: Mixed Effects (lmer)
    #Response: art(score)

    #                          F Df  Df.res     Pr(>F)    
    #1 month             40.8916  7 417.205 < 2.22e-16 ***
    #2 Site               7.0226  1  90.170  0.0095052  **
    #3 B2015            371.2417  1  90.061 < 2.22e-16 ***
    #4 month:Site         9.2513  7 417.356 1.1689e-10 ***
    #5 month:B2015       44.4792  7 417.394 < 2.22e-16 ***
    #6 Site:B2015         4.8256  1  90.062  0.0306073   *
    #7 month:Site:B2015   9.1426  7 417.452 1.5846e-10 ***


### tried with error not sure how to interpret
#Full_mod_mcaperr<-art(score~month*Site*B2015+Error(TagID), Mcap_df)  #using TagID as only random factor not PAIR

#post-hoc for 2-way model (From what I read, I think we can do this for things with 2 levels (ie     
		  #Bleached or not bleached, Reef 4 and Reef13, but not more, so month doesn't work)
 	#Reef by Bleaching history (not by month! moths are averaged)
		model.int_1=artlm(Full_mod_mcap, "Site:B2015")
		marginal1=emmeans(model.int_1, ~Site:B2015)
      contrast(marginal1, method="pairwise", adjust="none")	 #<Thoughts on "adjust"? none vs Tukey?
		  contrast(marginal1, method="pairwise", adjust="tukey")	#only 13N-13Y show sig diff across months.

		  	  
		  
#### My argument is that we use this model to show that all interactions are significant, and THEN we run models at specific time points to show what is going on at that time. I feel like that should make reviewers happy. 
		  
  ##** I dont think we want this
 	   #posthoc lib phia, testInteractions, Site:B2015  
    testInteractions(artlm(Month1.5_mcap, "Site:B2015"), pairwise=c("Site", "B2015"), adjustment="holm")	
      # 4-13 : N-Y 0.0111  
      # This is testing diffs between 4-13 NB corals vs diffs between 4-13 B corals, can't switch it...Any ideas?
    
        testInteractions(artlm(Month1.5_mcap, "B2015:Site"), across=c("B2015", "Site"), adjustment="holm")	#Works         across not pairwise. "The factors contained in across will appear as a full set of contrasts. By default          they will be orthogonal contrasts, unless overriden by the contrasts of the model data frame or by the            arguments passed down to testFactors. See the documentation of that function for further details.
        
        #F Test: 
            #P-value adjustment method: holm
            #           Value Df Sum of Sq      F Pr(>F)  
            #Mean      23.017  1      2461 6.8013 0.0111 *
            #Residuals        71     25691   		  


```

Mcap models at each time point, main effects Reef and B2015
```{r}
Mcap_0<-subset(Mcap_df, month=="Month0")
Mcap_0<-Mcap_0[,c(1,3,4,6)] 
Mcap_1.5<-subset(Mcap_df, month=="Month1.5")
Mcap_1.5<-Mcap_1.5[,c(1,3,4,6)] 
Mcap_3<-subset(Mcap_df, month=="Month3")
Mcap_3<-Mcap_1.5[,c(1,3,4,6)] 
Mcap_4.5<-subset(Mcap_df, month=="Month4.5")
Mcap_4.5<-Mcap_4.5[,c(1,3,4,6)] 
Mcap_6<-subset(Mcap_df, month=="Month6")
Mcap_6<-Mcap_6[,c(1,3,4,6)] 
Mcap_10<-subset(Mcap_df, month=="Month10")
Mcap_10<-Mcap_10[,c(1,3,4,6)] 
Mcap_18<-subset(Mcap_df, month=="Month18")
Mcap_18<-Mcap_18[,c(1,3,4,6)] 
Mcap_24<-subset(Mcap_df, month=="Month24")
Mcap_24<-Mcap_24[,c(1,3,4,6)] 


#Month0
Month0_mcap<-art(score~Site*B2015, Mcap_0)  #TagID removed bc not repeated measures
summary(Month0_mcap) #all should be ~0
anova(Month0_mcap) 
  #posthoc for single factors:
 	   contrast(emmeans(artlm(Month0_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

 	
 	   
#Month1.5
Month1.5_mcap<-art(score~B2015+Site+B2015:Site, Mcap_1.5)  #using TagID as only random factor not PAIR
summary(Month1.5_mcap) #all should be ~0,F-vals not all ~0, other options? 
anova(Month1.5_mcap) 

  #posthoc for single factors:
 	   contrast(emmeans(artlm(Month1.5_mcap, "Site"), ~ Site), method="pairwise") 
        #4-13 0.0051
 	   contrast(emmeans(artlm(Month1.5_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

            
# WE WANT THIS FOR INTERACTIONS THAT ARE ALREADY SIG. 
        #Mann-Whitney U tests tests on the original data.
        # cross-factor pairwise comparisons using Mann-Whitney U tests.
 	   ### AH can you look at the WARNINGS thrown and give your opinion?

		N4.Y4 <- wilcox.test(Mcap_1.5[Mcap_1.5$Site == "4" & Mcap_1.5$B2015 == "N",]$score, Mcap_1.5[Mcap_1.5$Site == "4" & Mcap_1.5$B2015 == "Y",]$score)$p.value  
		N13.Y13 <- wilcox.test(Mcap_1.5[Mcap_1.5$Site == "13" & Mcap_1.5$B2015 == "N",]$score, Mcap_1.5[Mcap_1.5$Site == "13" & Mcap_1.5$B2015 == "Y",]$score)$p.value 
	Y4.Y13 <- wilcox.test(Mcap_1.5[Mcap_1.5$Site == "4" & Mcap_1.5$B2015 == "Y",]$score, Mcap_1.5[Mcap_1.5$Site == "13" & Mcap_1.5$B2015 == "Y",]$score)$p.value 
        # correct for multiple comparisons using Holm's sequential Bonferroni procedure (Holm 1979)
		p.adjust(c(N4.Y4, N13.Y13, Y4.Y13), method="holm")
		# This returns all significantly different. 
    
#Month3
Month3_mcap<-art(score~Site+B2015+Site:B2015, Mcap_3)  #using TagID as only random factor not PAIR
summary(Month3_mcap) #all should be ~0, are for cols, aren't for F-value
anova(Month3_mcap)

  contrast(emmeans(artlm(Month3_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

# WE WANT THIS FOR INTERACTIONS THAT ARE ALREADY SIG. 
        #Mann-Whitney U tests tests on the original data.
        # cross-factor pairwise comparisons using Mann-Whitney U tests.

		N4.Y4 <- wilcox.test(Mcap_3[Mcap_3$Site == "4" & Mcap_3$B2015 == "N",]$score, Mcap_3[Mcap_3$Site == "4" & Mcap_3$B2015 == "Y",]$score)$p.value  
		N13.Y13 <- wilcox.test(Mcap_3[Mcap_3$Site == "13" & Mcap_3$B2015 == "N",]$score, Mcap_3[Mcap_3$Site == "13" & Mcap_3$B2015 == "Y",]$score)$p.value 
	Y4.Y13 <- wilcox.test(Mcap_3[Mcap_3$Site == "4" & Mcap_3$B2015 == "Y",]$score, Mcap_3[Mcap_3$Site == "13" & Mcap_3$B2015 == "Y",]$score)$p.value 
        # correct for multiple comparisons using Holm's sequential Bonferroni procedure (Holm 1979)
		p.adjust(c(N4.Y4, N13.Y13, Y4.Y13), method="holm")
		# This returns all significantly different.  4.443438e-04 1.227426e-07 3.355914e-02

#Month4.5
Month4.5_mcap<-art(score~Site*B2015, Mcap_4.5)  #TagID removed bc not repeated measures
summary(Month4.5_mcap) #all should be ~0 and are
anova(Month4.5_mcap) 

 contrast(emmeans(artlm(Month4.5_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

# WE WANT THIS FOR INTERACTIONS THAT ARE ALREADY SIG. 
        #Mann-Whitney U tests tests on the original data.
        # cross-factor pairwise comparisons using Mann-Whitney U tests.

		N4.Y4 <- wilcox.test(Mcap_4.5[Mcap_4.5$Site == "4" & Mcap_4.5$B2015 == "N",]$score, Mcap_4.5[Mcap_4.5$Site == "4" & Mcap_4.5$B2015 == "Y",]$score)$p.value  
		N13.Y13 <- wilcox.test(Mcap_4.5[Mcap_4.5$Site == "13" & Mcap_4.5$B2015 == "N",]$score, Mcap_4.5[Mcap_4.5$Site == "13" & Mcap_4.5$B2015 == "Y",]$score)$p.value 
	Y4.Y13 <- wilcox.test(Mcap_4.5[Mcap_4.5$Site == "4" & Mcap_4.5$B2015 == "Y",]$score, Mcap_4.5[Mcap_4.5$Site == "13" & Mcap_4.5$B2015 == "Y",]$score)$p.value 
        # correct for multiple comparisons using Holm's sequential Bonferroni procedure (Holm 1979)
		p.adjust(c(N4.Y4, N13.Y13, Y4.Y13), method="holm")
		# This returns all significantly different.  3.333033e-02 4.355652e-07 3.333033e-02

#Month6
Month6_mcap<-art(score~Site*B2015, Mcap_6)  #TagID removed bc not repeated measures
summary(Month6_mcap) #all should be ~0 and are
anova(Month6_mcap) 

contrast(emmeans(artlm(Month6_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

# WE WANT THIS FOR INTERACTIONS THAT ARE ALREADY SIG. 
        #Mann-Whitney U tests tests on the original data.
        # cross-factor pairwise comparisons using Mann-Whitney U tests.

		N4.Y4 <- wilcox.test(Mcap_6[Mcap_6$Site == "4" & Mcap_6$B2015 == "N",]$score, Mcap_6[Mcap_6$Site == "4" & Mcap_6$B2015 == "Y",]$score)$p.value  
		N13.Y13 <- wilcox.test(Mcap_6[Mcap_6$Site == "13" & Mcap_6$B2015 == "N",]$score, Mcap_6[Mcap_6$Site == "13" & Mcap_6$B2015 == "Y",]$score)$p.value 
	Y4.Y13 <- wilcox.test(Mcap_6[Mcap_6$Site == "4" & Mcap_6$B2015 == "Y",]$score, Mcap_6[Mcap_6$Site == "13" & Mcap_6$B2015 == "Y",]$score)$p.value 
        # correct for multiple comparisons using Holm's sequential Bonferroni procedure (Holm 1979)
		p.adjust(c(N4.Y4, N13.Y13, Y4.Y13), method="holm")
		# This returns all significantly different.  7.186167e-02 5.878972e-05 8.629725e-03

#Month10
Month10_mcap<-art(score~Site+B2015+Site:B2015, Mcap_10)  #using TagID as only random factor not PAIR
summary(Month10_mcap) #all should be ~0 and are
anova(Month10_mcap)


contrast(emmeans(artlm(Month10_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

#Month18
Month18_mcap<-art(score~Site+B2015+Site:B2015, Mcap_18)  #using TagID as only random factor not PAIR
summary(Month18_mcap) #all should be ~0 and are
anova(Month18_mcap)

contrast(emmeans(artlm(Month18_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001

#Month24
Month24_mcap<-art(score~Site+B2015+Site:B2015, Mcap_24)  #using TagID as only random factor not PAIR
summary(Month24_mcap) #all should be ~0 and are
anova(Month24_mcap)

contrast(emmeans(artlm(Month24_mcap, "B2015"), ~ B2015), method="pairwise") 
 	        #N - Y <.0001
```

*P.comp*  
# ARTools stats for non-parametric non-normality  

Full model ARTool

```{r} 

#Full timeframe Pcomp (log, arcsinsqrt don't help normailty)
 
hist(Pcomp_df$score) #not normal,use ARTools for non-parametric

Pcomp_df<-Pcomp_df[,c(1,2,4:7)] #subset and order cols for ARTool

Full_mod_Pcomp<-art(score~month*Site*B2015+(1|TagID), Pcomp_df)  #using TagID as only random factor not PAIR
summary(Full_mod_Pcomp) #all should be ~0 and are
anova(Full_mod_Pcomp) # Everything is significant

#Analysis of Variance of Aligned Rank Transformed Data

    #Table Type: Analysis of Deviance Table (Type III Wald F tests with Kenward-Roger df) 
    #Model: Mixed Effects (lmer)
    #Response: art(score) 
    #                             F Df  Df.res     Pr(>F)    
    #1 month            109.281  7 485.147 < 2.22e-16 ***
    #2 Site              93.571  1  80.893 3.7726e-15 ***
    #3 B2015            409.758  1  80.797 < 2.22e-16 ***
    #4 month:Site        22.472  7 483.989 < 2.22e-16 ***
    #5 month:B2015      132.779  7 485.104 < 2.22e-16 ***
    #6 Site:B2015        33.393  1  80.878 1.3546e-07 ***
    #7 month:Site:B2015  19.542  7 484.067 < 2.22e-16 ***


### tried with error not sure how to interpret
#Full_mod_Pcomperr<-art(score~month*Site*B2015+Error(TagID), Pcomp_df)  #using TagID as only random factor not PAIR

#post-hoc for 2-way model (From what I read, I think we can do this for things with 2 levels (ie     
		  #Bleached or not bleached, Reef 4 and Reef13, but not more, so month doesn't work)
 	#Reef by Bleaching history (not by month! moths are averaged)
		model.int_1=artlm(Full_mod_Pcomp, "Site:B2015")
		marginal1=emmeans(model.int_1, ~Site:B2015)
      contrast(marginal1, method="pairwise", adjust="none")	 #<Thoughts on "adjust"? none vs Tukey?
		  contrast(marginal1, method="pairwise", adjust="tukey")	#PR4 diff btw B/NB, PR13 diff btwn B/NB

```

Pcomp models at each time point, main effects Reef and B2015 ARTool
```{r}
Pcomp_0<-subset(Pcomp_df, month=="Month0") 
Pcomp_0<-Pcomp_0[,c(1,3,4,6)] 
Pcomp_1.5<-subset(Pcomp_df, month=="Month1.5")
Pcomp_1.5<-Pcomp_1.5[,c(1,3,4,6)] 
Pcomp_3<-subset(Pcomp_df, month=="Month3")
Pcomp_3<-Pcomp_3[,c(1,3,4,6)] 
Pcomp_4.5<-subset(Pcomp_df, month=="Month4.5")
Pcomp_4.5<-Pcomp_4.5[,c(1,3,4,6)] 
Pcomp_6<-subset(Pcomp_df, month=="Month6")
Pcomp_6<-Pcomp_6[,c(1,3,4,6)] 
Pcomp_10<-subset(Pcomp_df, month=="Month10")
Pcomp_10<-Pcomp_10[,c(1,3,4,6)] 
Pcomp_18<-subset(Pcomp_df, month=="Month18")
Pcomp_18<-Pcomp_18[,c(1,3,4,6)] 
Pcomp_24<-subset(Pcomp_df, month=="Month24")
Pcomp_24<-Pcomp_24[,c(1,3,4,6)] 


#Month0
Month0_Pcomp<-art(score~Site*B2015, Pcomp_0)  #TagID removed bc not repeated measures
summary(Month0_Pcomp) #all should be ~0 and are
anova(Month0_Pcomp) 

#Month1.5
Month1.5_Pcomp<-art(score~Site*B2015, Pcomp_1.5)  #using TagID as only random factor not PAIR
summary(Month1.5_Pcomp) #all should be ~0 and are
anova(Month1.5_Pcomp) 

#Month3
Month3_Pcomp<-art(score~Site*B2015, Pcomp_3)  #using TagID as only random factor not PAIR
summary(Month3_Pcomp) #all should be ~0 and most are, max 0.0012
anova(Month3_Pcomp)

#Month4.5
Month4.5_Pcomp<-art(score~Site*B2015, Pcomp_4.5)  #TagID removed bc not repeated measures
summary(Month4.5_Pcomp) #all should be ~0 and are close
anova(Month4.5_Pcomp) 


#Month6
Month6_Pcomp<-art(score~Site*B2015, Pcomp_6)  #TagID removed bc not repeated measures
summary(Month6_Pcomp) # all scores are 3, no variation
anova(Month6_Pcomp) ### Error in Anova.lm(m, type = "III", test = test, ...) : 
                          #residual sum of squares is 0 (within rounding error)

#Month10
Month10_Pcomp<-art(score~Site*B2015, Pcomp_10)  #using TagID as only random factor not PAIR
summary(Month10_Pcomp) #all should be ~0 and are
anova(Month10_Pcomp)

#Month18
Month18_Pcomp<-art(score~Site*B2015, Pcomp_18)  #using TagID as only random factor not PAIR
summary(Month18_Pcomp) #all should be ~0 and are
anova(Month18_Pcomp)

#Month24
Month24_Pcomp<-art(score~Site*B2015, Pcomp_24)  #using TagID as only random factor not PAIR
summary(Month24_Pcomp) #all should be ~0 and are
anova(Month24_Pcomp)
```







Do we want to even keep?
subsets for mcap stats
```{r} 
Mcappig06All<-subset(b0153456, Species=="Montipora capitata")
Mcappig36All<-subset(b3456, Species=="Montipora capitata")

#03
Mcappig03All_mod<-aov(score~month*Site*B2015, Mcappig03All)
summary(Mcappig03All_mod)
TukeyHSD(Mcappig03All_mod)
#06
Mcappig06All_mod<-aov(score~month*Site*B2015, Mcappig06All)
summary(Mcappig06All_mod)
TukeyHSD(Mcappig06All_mod)

#36
Mcappig06All_mod<-aov(score~month*Site*B2015, Mcappig06All)
summary(Mcappig06All_mod)
TukeyHSD(Mcappig06All_mod)
```


# ARTools TEST

ARTools: full model  AH START HERE
https://cran.r-project.org/web/packages/ARTool/vignettes/art-contrasts.html
```{r}

#03 
Mcappig03All<-Mcappig03All[,c(1,2,4:7)] #get data in order and only data needed for model (unique ID on left col, variable on far right)
hist(Mcappig03All$score) #not normal,use ARTools for non-parametric 
 
model_mcap03_1<-art(score~month*Site*B2015+(1|Pair)+(1|TagID), Mcappig03All)
anova(model_mcap03_1)
  #results
  #Table Type: Analysis of Deviance Table (Type III Wald F tests with Kenward-Roger df) 
  #Model: Mixed Effects (lmer)
  #Response: art(score)

      #                           F Df  Df.res     Pr(>F)    
      #1 month            120.46674  2 149.303 < 2.22e-16 ***
      #2 Site               0.93441  1  39.611    0.33958    
      #3 B2015            918.05363  1  38.942 < 2.22e-16 ***
      #4 month:Site        19.76773  2 149.737 2.4072e-08 ***
      #5 month:B2015      115.88767  2 149.945 < 2.22e-16 ***
      #6 Site:B2015         1.16330  1  39.366    0.28735    
      #7 month:Site:B2015  23.32563  2 150.006 1.5120e-09 ***


#posthoc - so this keeps giving me warnings due to involvement with interactions, ok to use? just warning. These are remodeling! not post hoc on above model
  
  #month only		 - over time (regardless of reef or Bleaching history, months differ)
    model.lm_1<-artlm(model_mcap03_1, "month")  #post hoc, get model ready to be passed into emmeans
		marginal<-emmeans(model.lm_1, ~month) # only modeling by month
		pairs(marginal, adjust="tukey") #tukey, all significant
		cld(marginal, alpha=0.05, Letters=letters, adjust="tukey") # this shows they are all diff. 
		#OR (this might be the same)
		contrast(emmeans(model.lm_1, ~month), method = "pairwise") # only modeling by month

  #Bleaching History (B2015)	- (regarless of time, Bleached corals and NB are diff still)	
		model.lm_2<-artlm(model_mcap03_1, "B2015")  #post hoc, get model ready to be passed into emmeans
		marginal2<-emmeans(model.lm_2, ~B2015) # only modeling by month
		pairs(marginal2, adjust="tukey") #tukey, significant
		cld(marginal2, alpha=0.05, Letters=letters, adjust="tukey") 
		
  #post-hoc for 2-way model (From what I read, I think we can do this for things with 2 levels (ie     
		  #Bleached or not bleached, Reef 4 and Reef13, but not more, so month doesn't work)
 	#Reef by Bleaching history (not by month! moths are averaged)
		model.int_1=artlm(model_mcap03_1, "Site:B2015")
		marginal1=emmeans(model.int_1, ~Site:B2015)
      contrast(marginal1, method="pairwise", adjust="none")	 #<Thoughts on "adjust"? none vs Tukey?
		  contrast(marginal, method="pairwise", adjust="tukey")		

#I think we use this for things with 3+ levels)
		  #Bleaching history by Reef
testInteractions(artlm(model_mcap03_1, "month:B2015"), pairwise=c("month","B2015"), adjustment="holm")		  #Chisq Test: 
    #P-value adjustment method: holm
      #                      Value Df   Chisq Pr(>Chisq)    
    #Month0-Month1.5 : N-Y 148.760  1 103.645  < 2.2e-16 ***
    #  Month0-Month3 : N-Y 217.877  1 217.685  < 2.2e-16 ***
    #Month1.5-Month3 : N-Y  69.117  1  20.488  6.002e-06 ***
# Interpret above as diff between Month0-Month1.5 of NB, and the diff between Month0-Month1.5 of B. FOr some reason, i can't change the order of "month:B2015" to "B2015:month".

#contrast tests...not sure if ok
testInteractions(artlm(model_mcap03_1, "month:Site"), pairwise=c("month","Site"), adjustment="holm")		   #month by Reef
    #Chisq Test: 
   #		                            Value Df   Chisq Pr(>Chisq)    
    #Month0-Month1.5 : 4-13   28.507  1  2.9161     0.0877 .  
    #  Month0-Month3 : 4-13  -77.427  1 21.0712  8.850e-06 ***
    #Month1.5-Month3 : 4-13 -105.933  1 36.9487  3.638e-09 ***
		  

#what about nonparametric Mann-Whitney U tests (bewteen subjects) or wilcoxon (within subjects)
#http://depts.washington.edu/acelab/proj/art/index.html
# cross-factor pairwise comparisons using Mann-Whitney U tests. Look ok? Not by Month, across them. 
		N4_vs_Y4 = wilcox.test(Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "N",]$score, Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "Y",]$score)$p.value
	 	N13_vs_Y13 = wilcox.test(Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "N",]$score, Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "Y",]$score)$p.value
	  Y4_vs_Y13 = wilcox.test(Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "Y",]$score, Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "Y",]$score)$p.value
	  N4_vs_N13 = wilcox.test(Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "N",]$score, Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "N",]$score)$p.value

# did below bc doc suggested, not sure how I feel. Now all are significant. http://depts.washington.edu/acelab/proj/art/index.html
# correct for multiple comparisons using Holm's sequential Bonferroni procedure (Holm 1979)
		p.adjust(c(N4_vs_Y4, N13_vs_Y13, Y4_vs_Y13,N4_vs_N13), method="holm")

		
		
#new idea #### AH do you think we can present the 'full model' and then for post hoc run specific models (or do we run into same issues again in review)
  

```


## Rare of Bleaching Recovery: Percent change from month 0 to month 3  
We calculated the mean bleaching recovery at 0-3 months and 3-6 months.      
0-3 months RATE
```{r}
## 1. 0-3 months
#Calc average rate per colony 0-3 months
# Recovery in %
b03$Rate03<-(((b03$Month3-b03$Month0)))

#Average the means of the rates. 

# calc average  bleaching rate/month in %  per grouping
b03means <- ddply(b03, c("Site", "Species", "B2015"), summarise, 
                   N    = length(Rate03[!is.na(Rate03)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Rate03, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Rate03, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Rate03, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#b03means #display table

b03means$code<-paste0(b03means$Site, b03means$B2015, b03means$Species)
b03means #display table
```

0-6 months RATE
```{r}
#recovery in %
b06$Rate06<-(((b06$Month6-b06$Month0)))

#Average the means of the rates. 

# calc average  bleaching rate/month in %  per grouping
b06means <- ddply(b06, c("Site", "Species", "B2015"), summarise, 
                   N    = length(Rate06[!is.na(Rate06)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Rate06, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Rate06, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Rate06, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
b06means #display table

b06means$code<-paste0(b06means$Site, b06means$B2015, b06means$Species)
```

Rate change       <- we didn't use this for Color score, just Tissue loss

`Rate ~ Lagoon*B2015`  
*Montipora capitata only*    
 Months 0-3
```{r}
#subset Mcap
Mcap_b03<-subset(b03, Species=="Montipora capitata") #subset Mcap
#0-3
Mcap_b03r<-aov(Rate03~Site* B2015, Mcap_b03)
summary(Mcap_b03r)
TukeyHSD(Mcap_b03r)

#0-6
Mcap_b06<-subset(b06, Species=="Montipora capitata") #subset Mcap
Mcap_b06r<-aov(Rate06~Site * B2015, Mcap_b06)
summary(Mcap_b06r)
TukeyHSD(Mcap_b06r)
```

##*Porites compressa only*
`Bleaching score ~ Reef Site * Bleaching History`  
###Months 0-3, Rate change
Model: `score~Site*B2015`  
```{r}
#subset Pcomp
Pcomp_b03<-subset(b03, Species=="Porites compressa") #subset Pcomp
Pcomp_b03r<-aov(Rate03~Site  * B2015, Pcomp_b03)
summary(Pcomp_b03r)
TukeyHSD(Pcomp_b03r)

#0-6
Pcomp_b06<-subset(b06, Species=="Porites compressa") #subset Pcomp
Pcomp_b06r<-aov(Rate06~Site * B2015, Pcomp_b06)
summary(Pcomp_b06r)
TukeyHSD(Pcomp_b06r)

```
ARTools mcap:pairs Delete
```{r}
#set up Pairs data
Mcappig<-subset(PairDat, Species =="Montipora capitata")            
McappigArt<-Mcappig



#All
MpigPair_all_mod1ART<-art(score~Site*B2015*month+(1|TagID)+(1|Pair), Mcappig)
summary(MpigPair_all_mod1ART)
anova(MpigPair_all_mod1ART)

#Month0
MpigPair0ART<-subset(Mcappig, month =="Month0")           
MpigPair0_mod1ART<-art(score~Site*B2015, MpigPair0ART)
summary(MpigPair0_mod1ART)
anova(MpigPair0_mod1ART)

#Month1.5
MpigPair1.5ART<-subset(Mcappig, month =="Month1.5")           
MpigPair1.5_mod1ART<-art(score~Site*B2015, MpigPair1.5ART)
summary(MpigPair1.5_mod1ART) #not good
anova(MpigPair1.5_mod1ART)

#Month3
MpigPair3ART<-subset(Mcappig, month =="Month3")           
MpigPair3_mod1ART<-art(score~Site*B2015, MpigPair3ART)
summary(MpigPair3_mod1ART)
anova(MpigPair3_mod1ART)

#Month4.5
MpigPair4.5ART<-subset(Mcappig, month =="Month4.5")           
MpigPair4.5_mod1ART<-art(score~Site*B2015, MpigPair4.5ART)
summary(MpigPair4.5_mod1ART) #not good
anova(MpigPair4.5_mod1ART)

#Month6
MpigPair6ART<-subset(Mcappig, month =="Month6")           
MpigPair6_mod1ART<-art(score~Site*B2015, MpigPair6ART)
summary(MpigPair6_mod1ART)#not good
anova(MpigPair6_mod1ART)

#Month10
MpigPair10ART<-subset(Mcappig, month =="Month10")           
MpigPair10_mod1ART<-art(score~Site*B2015, MpigPair10ART)
summary(MpigPair10_mod1ART)#not good
anova(MpigPair10_mod1ART)

#Month18
MpigPair18ART<-subset(Mcappig, month =="Month18")           
MpigPair18_mod1ART<-art(score~Site*B2015, MpigPair18ART)
summary(MpigPair18_mod1ART)#notgood
anova(MpigPair18_mod1ART)

#Month24
MpigPair24ART<-subset(Mcappig, month =="Month24")           
MpigPair24_mod1ART<-art(score~Site*B2015, MpigPair24ART)
summary(MpigPair24_mod1ART)#really not good
anova(MpigPair24_mod1ART)
```



ARTools pcomp:pairs Delete
```{r}
#set up Pairs data
Pcomppig<-subset(PairDat, Species =="Porites compressa")            

#All
PpigPair_all_mod1ART<-art(score~Site*B2015*month+(1|TagID)+(1|Pair), Pcomppig)
summary(PpigPair_all_mod1ART)# not all 0
anova(PpigPair_all_mod1ART)

#Month0
PpigPair0ART<-subset(Pcomppig, month =="Month0")           
PpigPair0_mod1ART<-art(score~Site*B2015, PpigPair0ART)
summary(PpigPair0_mod1ART)#not0
anova(PpigPair0_mod1ART)

#Month1.5
PpigPair1.5ART<-subset(Pcomppig, month =="Month1.5")           
PpigPair1.5_mod1ART<-art(score~Site*B2015, PpigPair1.5ART)
summary(PpigPair1.5_mod1ART) #0
anova(PpigPair1.5_mod1ART)

#Month3
PpigPair3ART<-subset(Pcomppig, month =="Month3")           
PpigPair3_mod1ART<-art(score~Site*B2015, PpigPair3ART)
summary(PpigPair3_mod1ART) #close to 0
anova(PpigPair3_mod1ART)

#Month4.5
PpigPair4.5ART<-subset(Pcomppig, month =="Month4.5")           
PpigPair4.5_mod1ART<-art(score~Site*B2015, PpigPair4.5ART)
summary(PpigPair4.5_mod1ART) #not all 0
anova(PpigPair4.5_mod1ART)

#Month6
PpigPair6ART<-subset(Pcomppig, month =="Month6")           
PpigPair6_mod1ART<-art(score~Site*B2015, PpigPair6ART)
summary(PpigPair6_mod1ART)#Error in Anova.lm(m, type = "III", test = test, ...) :residual sum of squares is 0 (within rounding error)
anova(PpigPair6_mod1ART)

#Month10
PpigPair10ART<-subset(Pcomppig, month =="Month10")           
PpigPair10_mod1ART<-art(score~Site*B2015, PpigPair10ART)
summary(PpigPair10_mod1ART)#not all 0
anova(PpigPair10_mod1ART)

#Month18
PpigPair18ART<-subset(Pcomppig, month =="Month18")           
PpigPair18_mod1ART<-art(score~Site*B2015, PpigPair18ART)
summary(PpigPair18_mod1ART)#not all 0
anova(PpigPair18_mod1ART)

#Month24
PpigPair24ART<-subset(Pcomppig, month =="Month24")           
PpigPair24_mod1ART<-art(score~Site*B2015, PpigPair24ART)
summary(PpigPair24_mod1ART)#not all 0
anova(PpigPair24_mod1ART)
```

0-3 and 0-6 month differences <- cut?
Pairs data
score~month*B2015*lagoon
*M. capitata* Pairs
```{r}
Pcomppig03Pairs<-subset(pigPairs03, Species=="Montipora capitata")
Pcomppig06Pairs<-subset(pigPairs06, Species=="Montipora capitata")

#0-3
Pcomppig03Pairs_mod<-aov(score~month*Site*B2015, Pcomppig03Pairs)
summary(Pcomppig03Pairs_mod)

#0-6
Pcomppig06Pairs_mod<-aov(score~month*Site*B2015, Pcomppig06Pairs)
summary(Pcomppig06Pairs_mod)
```

Pcomp Pairs
```{r}
Pcomppig03Pairs<-subset(pigPairs03, Species=="Porites compressa")
Pcomppig06Pairs<-subset(pigPairs06, Species=="Porites compressa")
Pcomppig36Pairs<-subset(pigPairs36, Species=="Porites compressa")

#0-3
Pcomppig03Pairs_mod<-aov(score~month*Site*B2015, Pcomppig03Pairs)
summary(Pcomppig03Pairs_mod)
#06
Pcomppig06Pairs_mod<-aov(score~month*Site*B2015, Pcomppig06Pairs)
summary(Pcomppig06Pairs_mod)
```
