---
title: "B_NB_stats"
author: "Shayle Matsuda"
date: "08/08/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
update:"03/31/3030"

Color score for *M. capitata* and *P. compressa* colonies at the Inner Lagoon (PR4) and Outer Lagoon (PR13) after the 2015 coral bleaching event in Kaneohe Bay.  Tagged individuals were given a color score based on their visual color: 1) white (>80% of colony white with no visible pigmentation); 2) pale (>10% colony affected by pigment loss); or 3) fully pigmented (<10% colony with any pale coloration). 

```{r setup, include=FALSE,results="hide"}
knitr::opts_chunk$set(echo = TRUE)
```

# Bleached/Non Bleached data 
```{r, results="hide", message=FALSE}
#Set up: load libraries

rm(list=ls(all=TRUE)) 
library(phia)
library(emmeans)
library(ARTool)
library("reshape2") #reshape data
library("plotrix") #functions in tapply
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library(effects)
library("multcomp") 
library("multcompView")
library("coefplot") #coefficient plotting
library("seacarb") #seawater carbonate chemistry
library("vegan") #calculating distance matrices
library(betareg)
library(coin)
library(exactRankTests)
library(mvtnorm)
library(DTK)
library(PMCMR)
library(MuMIn)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(ggsignif)
#library(HH)
library(PMCMR)
library(tidyverse)
library(arm)
library(jtools)
library(broom)
library(ggstance)
library(officer)
library(huxtable)
library(flextable)
```

Load in data and organize 
```{r}
#We excluded these pairs as part of the pair analysis because the Bleached coral was only a 2: Porites Outer Bay 233_234, Porites Inner Bay 297_298, Porites Outer Bay 227_228, Montipora capitata Inner Bay 251_252.

# Bleaching data
bData<-read.csv("Bleaching_FINAL.csv") #read in bleaching df
bData$Site<-as.factor(bData$Site) #site as factor
bData$Pair<-as.factor(bData$Pair) #pair ID as factor

#organize data
bALL<-bData
  #melt
  bALLData1<-melt(data = bALL, measure.vars = c("Month0","Month1.5","Month3","Month4.5", "Month6","Month10","Month18","Month24"))   
  colnames(bALLData1)[colnames(bALLData1)=="variable"] <- "month"   # change col name from 'variable' to 'month'
  colnames(bALLData1)[colnames(bALLData1)=="value"] <- "score"   # change col name from 'value' to 'score'
  bALLData1$group<-paste(bALLData1$Species, bALLData1$Site, bALLData1$B2015, bALLData1$month) # add column for groups for figures only
#write.csv(bALL, "PigmentationBNB_20191122.csv")

#Remove all NON PAIRS and all NAs
  bALLData<-na.omit(bALLData1) # remove the NA rows (this removes individuals and NAs from pairs with no score at particular timepoints (due to not finding the coral))
```

Subsets of df: (edit for what we end up using!)
```{r}

Mcap_df<-subset(bALLData, Species=="Montipora capitata")   #Mcap only
Pcomp_df<-subset(bALLData, Species=="Porites compressa")   #Pcomp only


#subset 0-3 months
  b03<-subset(bALLData, month=="Month0" |month=="Month3")
  b03<-subset(b03, select = -group )

      b03<-dcast(b03, Species+Site+Pair+B2015~month) # cast()
  b03<-na.omit(b03) # remove the NA colums
  
#subset 0-6 months
b06<-subset(bALLData, month=="Month0" |month=="Month6")
b06<-subset(b06, select = -group )

      b06<-dcast(b06, Species+Site+Pair+B2015~month) # cast()
      b06<-na.omit(b06) # remove the NA colums

#subset 0-1.5-3 months for pairs comparison
b0153<-subset(bALLData, month=="Month0" |month=="Month1.5"|month=="Month3")
   #   b0153<-dcast(b0153, Species+Site+Pair+B2015~month) # cast()
   #   b0153<-na.omit(b0153) # remove the NA colums

#subset 3-4.5-6 months for pairs comparison
b3456<-subset(bALLData, month=="Month3" |month=="Month4.5"|month=="Month6")

#subset 0-1.5-3-4.5-6 months for pairs comparison
b0153456<-subset(bALLData, month=="Month0" |month=="Month1.5"|month=="Month3"|month=="Month4.5"|month=="Month6")
      #b0153456<-dcast(b0153456, Species+Site+Pair+B2015~month) # cast()
      #b0153456<-na.omit(b0153456) # remove the NA colums
    #subset 0-1.5-3-4.5-6 months for pairs comparison
``` 

Mean bleaching scores full df
```{r, results=TRUE}
#calc average bleaching score per grouping by time point
ballmeans <- ddply(bALLData, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
ballmeans #display table

ballmeans$code<-paste0(ballmeans$Site, ballmeans$B2015, ballmeans$Species)
ballmeans #display table
#write.csv(ballmeans,"ballmeans.csv")

#replace "Month#" with actual numbers
ballmeans$Month[ballmeans$month=="Month0"]<-0              # new col replace month #s only
ballmeans$Month[ballmeans$month=="Month1.5"]<-1.5  
ballmeans$Month[ballmeans$month=="Month3"]<-3 
ballmeans$Month[ballmeans$month=="Month4.5"]<-4.5
ballmeans$Month[ballmeans$month=="Month6"]<-6  
ballmeans$Month[ballmeans$month=="Month10"]<-10  
ballmeans$Month[ballmeans$month=="Month18"]<-18
ballmeans$Month[ballmeans$month=="Month24"]<-24

```

#Figure 5
Bleaching score means 0-24 months by species. 
*Montipora capitata*  
Figure 5 
```{r}
#subset for Mcap only
Mcap024Allplot<-subset(ballmeans, Species=="Montipora capitata")     #subset Mcap
 
#subset by reef
#PR13
Mcap024AllplotPR13<-subset(Mcap024Allplot, Site==13)
#PR4
Mcap024AllplotPR4<-subset(Mcap024Allplot, Site==4)
```

*M. capitata* PR13
```{r}
ballplotMcapPR13<-ggplot(data=Mcap024AllplotPR13, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold",hjust = 0.5));ballplotMcapPR13
McapOuter<-ballplotMcapPR13 + theme(legend.position = "none") #remove legend
```

*M. capitata* PR4
```{r}
ballplotMcapPR4<-ggplot(data=Mcap024AllplotPR4, aes(x=Month, y=mean, group = code, color=B2015),show.legend = FALSE) + #color by bleaching hist. 
 geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Montipora capitata\nColor score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold", hjust = 0.5));ballplotMcapPR4
McapInner<-ballplotMcapPR4 + theme(legend.position = "none") #remove legend
McapInner
```

*Porites compressa*  
Figure 5
```{r}
Pcomp024Allplot<-subset(ballmeans, Species=="Porites compressa")

#subset by reef
Pcomp024AllplotPR13<-subset(Pcomp024Allplot, Site==13)
Pcomp024AllplotPR4<-subset(Pcomp024Allplot, Site==4)
```

*P. compress* 
Outer PR13
```{r}
b024plotPcompPR13<-ggplot(data=Pcomp024AllplotPR13, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Porites compressa Bleaching Scores months 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR13

PcompOuter<-b024plotPcompPR13+ theme(legend.position = "none") #remove legend
```

*P. compressa*   
Inner PR4
```{r}
b024plotPcompPR4<-ggplot(data=Pcomp024AllplotPR4, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=3)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Porites compressa\nColor score") +
    xlab("Recovery Time (months)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Porites compressa Bleaching Scores months 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR4


PcompInner<-b024plotPcompPR4+ theme(legend.position = "none") #remove legend
PcompInner
```

Plot all four together, Figure 5
```{r}
grid.arrange(McapInner, McapOuter, PcompInner,PcompOuter, nrow = 2)
```


#Table 2  
PIGMENTATION 0-3 months and 0-6 months
score~month*B2015*lagoon
*M. capitata*  
# ARTools stats for non-parametric non-normality  
Full model
```{r} 

#Full timeframe Mcap (log, arcsinsqrt don't help normailty)

hist(Mcap_df$log) #not normal,use ARTools for non-parametric

Mcap_df<-Mcap_df[,c(1,2,4:7)] #subset and order cols for ARTool

Full_mod_mcap<-art(score~month*Site*B2015+(1|TagID), Mcap_df)  #using TagID as only random factor not PAIR

summary(Full_mod_mcap) #all should be ~0 and are
anova(Full_mod_mcap) # Everything is significant

#Analysis of Variance of Aligned Rank Transformed Data

    #Table Type: Analysis of Deviance Table (Type III Wald F tests with Kenward-Roger df) 
    #Model: Mixed Effects (lmer)
    #Response: art(score)

    #                          F Df  Df.res     Pr(>F)    
    #1 month             40.8916  7 417.205 < 2.22e-16 ***
    #2 Site               7.0226  1  90.170  0.0095052  **
    #3 B2015            371.2417  1  90.061 < 2.22e-16 ***
    #4 month:Site         9.2513  7 417.356 1.1689e-10 ***
    #5 month:B2015       44.4792  7 417.394 < 2.22e-16 ***
    #6 Site:B2015         4.8256  1  90.062  0.0306073   *
    #7 month:Site:B2015   9.1426  7 417.452 1.5846e-10 ***


#post-hoc for 2-way model (From what I read, I think we can do this for things with 2 levels (ie     
		  #Bleached or not bleached, Reef 4 and Reef13, but not more, so month doesn't work)
 	#Reef by Bleaching history (not by month! moths are averaged)
		model.int_1=artlm(Full_mod_mcap, "Site:B2015")
		marginal1=emmeans(model.int_1, ~Site:B2015)
      contrast(marginal1, method="pairwise", adjust="none")	 #<Thoughts on "adjust"? none vs Tukey?
		  contrast(marginal1, method="pairwise", adjust="tukey")	#only 13N-13Y show sig diff across months.
		  
		  
#### My argument is that we use this model to show that all interactions are significant, and THEN we run models at specific time points to show what is going on at that time. I feel like that should make reviewers happy.   
		  

```

subsets for mcap stats
```{r}
Mcappig03All<-subset(b0153, Species=="Montipora capitata")
Mcappig06All<-subset(b0153456, Species=="Montipora capitata")
Mcappig36All<-subset(b3456, Species=="Montipora capitata")

#03
Mcappig03All_mod<-aov(score~month*Site*B2015, Mcappig03All)
summary(Mcappig03All_mod)
TukeyHSD(Mcappig03All_mod)
#06
Mcappig06All_mod<-aov(score~month*Site*B2015, Mcappig06All)
summary(Mcappig06All_mod)
TukeyHSD(Mcappig06All_mod)

#36
Mcappig06All_mod<-aov(score~month*Site*B2015, Mcappig06All)
summary(Mcappig06All_mod)
TukeyHSD(Mcappig06All_mod)
```


# ARTools TEST

ARTools: full model  AH START HERE
https://cran.r-project.org/web/packages/ARTool/vignettes/art-contrasts.html
```{r}

#03 
Mcappig03All<-Mcappig03All[,c(1,2,4:7)] #get data in order and only data needed for model (unique ID on left col, variable on far right)
hist(Mcappig03All$score) #not normal,use ARTools for non-parametric

model_mcap03_1<-art(score~month*Site*B2015+(1|Pair)+(1|TagID), Mcappig03All)
anova(model_mcap03_1)
  #results
  #Table Type: Analysis of Deviance Table (Type III Wald F tests with Kenward-Roger df) 
  #Model: Mixed Effects (lmer)
  #Response: art(score)

      #                           F Df  Df.res     Pr(>F)    
      #1 month            120.46674  2 149.303 < 2.22e-16 ***
      #2 Site               0.93441  1  39.611    0.33958    
      #3 B2015            918.05363  1  38.942 < 2.22e-16 ***
      #4 month:Site        19.76773  2 149.737 2.4072e-08 ***
      #5 month:B2015      115.88767  2 149.945 < 2.22e-16 ***
      #6 Site:B2015         1.16330  1  39.366    0.28735    
      #7 month:Site:B2015  23.32563  2 150.006 1.5120e-09 ***



#posthoc - so this keeps giving me warnings due to involvement with interactions, ok to use? just warning. These are remodeling! not post hoc on above model
 
  #month only		 - over time (regardless of reef or Bleaching history, months differ)
    model.lm_1<-artlm(model_mcap03_1, "month")  #post hoc, get model ready to be passed into emmeans
		marginal<-emmeans(model.lm_1, ~month) # only modeling by month
		pairs(marginal, adjust="tukey") #tukey, all significant
		cld(marginal, alpha=0.05, Letters=letters, adjust="tukey") # this shows they are all diff. 
		#OR (this might be the same)
		contrast(emmeans(model.lm_1, ~month), method = "pairwise") # only modeling by month

  #Bleaching History (B2015)	- (regarless of time, Bleached corals and NB are diff still)	
		model.lm_2<-artlm(model_mcap03_1, "B2015")  #post hoc, get model ready to be passed into emmeans
		marginal2<-emmeans(model.lm_2, ~B2015) # only modeling by month
		pairs(marginal2, adjust="tukey") #tukey, significant
		cld(marginal2, alpha=0.05, Letters=letters, adjust="tukey") 
		
  #post-hoc for 2-way model (From what I read, I think we can do this for things with 2 levels (ie     
		  #Bleached or not bleached, Reef 4 and Reef13, but not more, so month doesn't work)
 	#Reef by Bleaching history (not by month! moths are averaged)
		model.int_1=artlm(model_mcap03_1, "Site:B2015")
		marginal1=emmeans(model.int_1, ~Site:B2015)
      contrast(marginal1, method="pairwise", adjust="none")	 #<Thoughts on "adjust"? none vs Tukey?
		  contrast(marginal, method="pairwise", adjust="tukey")		

#I think we use this for things with 3+ levels)
		  #Bleaching history by Reef
testInteractions(artlm(model_mcap03_1, "month:B2015"), pairwise=c("month","B2015"), adjustment="holm")		  #Chisq Test: 
    #P-value adjustment method: holm
      #                      Value Df   Chisq Pr(>Chisq)    
    #Month0-Month1.5 : N-Y 148.760  1 103.645  < 2.2e-16 ***
    #  Month0-Month3 : N-Y 217.877  1 217.685  < 2.2e-16 ***
    #Month1.5-Month3 : N-Y  69.117  1  20.488  6.002e-06 ***
# Interpret above as diff between Month0-Month1.5 of NB, and the diff between Month0-Month1.5 of B. FOr some reason, i can't change the order of "month:B2015" to "B2015:month".

#contrast tests...not sure if ok
testInteractions(artlm(model_mcap03_1, "month:Site"), pairwise=c("month","Site"), adjustment="holm")		   #month by Reef
    #Chisq Test: 
   #		                            Value Df   Chisq Pr(>Chisq)    
    #Month0-Month1.5 : 4-13   28.507  1  2.9161     0.0877 .  
    #  Month0-Month3 : 4-13  -77.427  1 21.0712  8.850e-06 ***
    #Month1.5-Month3 : 4-13 -105.933  1 36.9487  3.638e-09 ***
		  

#what about nonparametric Mann-Whitney U tests (bewteen subjects) or wilcoxon (within subjects)
#http://depts.washington.edu/acelab/proj/art/index.html
# cross-factor pairwise comparisons using Mann-Whitney U tests. Look ok? Not by Month, across them. 
		N4_vs_Y4 = wilcox.test(Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "N",]$score, Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "Y",]$score)$p.value
	 	N13_vs_Y13 = wilcox.test(Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "N",]$score, Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "Y",]$score)$p.value
	  Y4_vs_Y13 = wilcox.test(Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "Y",]$score, Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "Y",]$score)$p.value
	  N4_vs_N13 = wilcox.test(Mcappig03All[Mcappig03All$Site == "4" & Mcappig03All$B2015 == "N",]$score, Mcappig03All[Mcappig03All$Site == "13" & Mcappig03All$B2015 == "N",]$score)$p.value

# did below bc doc suggested, not sure how I feel. Now all are significant. http://depts.washington.edu/acelab/proj/art/index.html
# correct for multiple comparisons using Holm's sequential Bonferroni procedure (Holm 1979)
		p.adjust(c(N4_vs_Y4, N13_vs_Y13, Y4_vs_Y13,N4_vs_N13), method="holm")

		
		
#new idea #### AH do you think we can present the 'full model' and then for post hoc run specific models (or do we run into same issues again in review)
  
#new idea 2# 
#### AH #### To get around the issue of not being able to compare reef/bleachinghist/time, could we instead make a code for reef/bleachinghist/time and run that way? https://cran.r-project.org/web/packages/ARTool/vignettes/art-contrasts.html
   
   #OR Tests of differences of differences in interactions <- not running
contrast(emmeans(model_mcap03_1, B2015:Site), method="pairwise", interaction=TRUE)


Mcappig03All_code<-Mcappig03All #copy df to mess with
Mcappig03All_code$code<-paste0(Mcappig03All_code$B2015, Mcappig03All_code$Site) #make code col with b2015, site and month
Mcappig03All_code$code<-as.factor(Mcappig03All_code$code)
   Mcappig03All_code<-Mcappig03All_code[,c(1,2,5,7,6)] #get rid of col and reorder
   
   model_Mcappig03All_code<-artlm(score ~ month*code+(1|TagID/Pair), Mcappig03All_code)
anova(model_mcap03_1)
   
```






*P. compressa*  
Table 2
```{r}
Pcomppig03All<-subset(b0153, Species=="Porites compressa")
Pcomppig06All<-subset(b0153456, Species=="Porites compressa")
Pcomppig36All<-subset(b3456, Species=="Porites compressa")

#0-3
Pcomppig03All_mod<-aov(score~month*Site*B2015, Pcomppig03All)
summary(Pcomppig03All_mod)
TukeyHSD(Pcomppig03All_mod)
#0-6
Pcomppig06All_mod<-aov(score~month*Site*B2015, Pcomppig06All)
summary(Pcomppig06All_mod)
TukeyHSD(Pcomppig06All_mod)

#36
Pcomppig36All_mod<-aov(score~month*Site*B2015, Pcomppig36All)
summary(Pcomppig36All_mod)
TukeyHSD(Pcomppig36All_mod)
```

# Bleaching Recovery Models
SI Table 11  
`Bleaching Score ~ Reef Site * Bleaching History   
Model: `score~Site*B2015`
*Montipora capiatata*  
```{r}
Mcap_bALLData<-subset(bALLData, Species=="Montipora capitata")   # pull out Mcap

#Month0
Mcap_b0<-subset(Mcap_bALLData, month=="Month0")
Mcap_b0_mod1<-aov(score~Site*B2015, Mcap_b0)
summary(Mcap_b0_mod1)
TukeyHSD(Mcap_b0_mod1)

#Month1.5
Mcap_b1_5<-subset(Mcap_bALLData, month=="Month1.5")
Mcap_b1_5_mod1<-aov(score~Site*B2015, Mcap_b1_5)
summary(Mcap_b1_5_mod1)
TukeyHSD(Mcap_b1_5_mod1)

#Month3
Mcap_b3<-subset(Mcap_bALLData, month=="Month3")
Mcap_b3_mod1<-aov(score~Site*B2015, Mcap_b3)
summary(Mcap_b3_mod1)
TukeyHSD(Mcap_b3_mod1)

#Month4.5
Mcap_b4_5<-subset(Mcap_bALLData, month=="Month4.5")
Mcap_b4_5_mod1<-aov(score~Site*B2015, Mcap_b4_5)
summary(Mcap_b4_5_mod1)
TukeyHSD(Mcap_b4_5_mod1)

#Month6
Mcap_b6<-subset(Mcap_bALLData, month=="Month6")
Mcap_b6_mod1<-aov(score~Site*B2015, Mcap_b6)
summary(Mcap_b6_mod1)
TukeyHSD(Mcap_b6_mod1)
```

*Porites compressa*  
SI Table 11
```{r}
Pcomp_bALLData<-subset(bALLData, Species=="Porites compressa")   # pull out Pcomp

#Month0
Pcomp_b0<-subset(Pcomp_bALLData, month=="Month0")
Pcomp_b0_mod1<-aov(score~Site*B2015, Pcomp_b0)
summary(Pcomp_b0_mod1)
TukeyHSD(Pcomp_b0_mod1)

#Month1.5
Pcomp_b1_5<-subset(Pcomp_bALLData, month=="Month1.5")
Pcomp_b1_5_mod1<-aov(score~Site*B2015, Pcomp_b1_5)
summary(Pcomp_b1_5_mod1)
TukeyHSD(Pcomp_b1_5_mod1)

#Month3
Pcomp_b3<-subset(Pcomp_bALLData, month=="Month3")
Pcomp_b3_mod1<-aov(score~Site*B2015, Pcomp_b3)
summary(Pcomp_b3_mod1)

#Month4.5
Pcomp_b4_5<-subset(Pcomp_bALLData, month=="Month4.5")
Pcomp_b4_5_mod1<-aov(score~Site*B2015, Pcomp_b4_5)
summary(Pcomp_b4_5_mod1)

#Month6
Pcomp_b6<-subset(Pcomp_bALLData, month=="Month6")
Pcomp_b6_mod1<-aov(score~Site*B2015, Pcomp_b6)
summary(Pcomp_b6_mod1)
```

## Rare of Bleaching Recovery: Percent change from month 0 to month 3  
We calculated the mean bleaching recovery at 0-3 months and 3-6 months.      
0-3 months RATE
```{r}
## 1. 0-3 months
#Calc average rate per colony 0-3 months
# Recovery in %
b03$Rate03<-(((b03$Month3-b03$Month0)))

#Average the means of the rates. 

# calc average  bleaching rate/month in %  per grouping
b03means <- ddply(b03, c("Site", "Species", "B2015"), summarise, 
                   N    = length(Rate03[!is.na(Rate03)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Rate03, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Rate03, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Rate03, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#b03means #display table

b03means$code<-paste0(b03means$Site, b03means$B2015, b03means$Species)
b03means #display table
```

0-6 months RATE
```{r}
#recovery in %
b06$Rate06<-(((b06$Month6-b06$Month0)))

#Average the means of the rates. 

# calc average  bleaching rate/month in %  per grouping
b06means <- ddply(b06, c("Site", "Species", "B2015"), summarise, 
                   N    = length(Rate06[!is.na(Rate06)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Rate06, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Rate06, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Rate06, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
b06means #display table

b06means$code<-paste0(b06means$Site, b06means$B2015, b06means$Species)
```

Rate change  
`Rate ~ Lagoon*B2015`  
*Montipora capitata only*    
 Months 0-3
```{r}
#subset Mcap
Mcap_b03<-subset(b03, Species=="Montipora capitata") #subset Mcap
#0-3
Mcap_b03r<-aov(Rate03~Site* B2015, Mcap_b03)
summary(Mcap_b03r)
TukeyHSD(Mcap_b03r)

#0-6
Mcap_b06<-subset(b06, Species=="Montipora capitata") #subset Mcap
Mcap_b06r<-aov(Rate06~Site * B2015, Mcap_b06)
summary(Mcap_b06r)
TukeyHSD(Mcap_b06r)
```

##*Porites compressa only*
`Bleaching score ~ Reef Site * Bleaching History`  
###Months 0-3, Rate change
Model: `score~Site*B2015`  
```{r}
#subset Pcomp
Pcomp_b03<-subset(b03, Species=="Porites compressa") #subset Pcomp
Pcomp_b03r<-aov(Rate03~Site  * B2015, Pcomp_b03)
summary(Pcomp_b03r)
TukeyHSD(Pcomp_b03r)

#0-6
Pcomp_b06<-subset(b06, Species=="Porites compressa") #subset Pcomp
Pcomp_b06r<-aov(Rate06~Site * B2015, Pcomp_b06)
summary(Pcomp_b06r)
TukeyHSD(Pcomp_b06r)

```


# Pairs 
(only include colonies when both partners of a pair are found)
```{r,results="hide"}
#Figure out which of the pairs we lose 
#subset only pairs
naData<-bALLData1
#dataset to work from (pull unwanted from)
pData<-bALLData1

# Identify which pairs have a partner not found
# Subset Month0 (two colonies missing values, checked datasheet)
M0<-subset(naData, month=="Month0") # subset month 0
M0_NA <- M0 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#keep it all

  # subset month 0
  M0a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M0<-pData[M0a]
  M0<-subset(M0, month=="Month0")
  M0<-na.omit(M0) # remove the NA colums
  #DIFFS
  M0diff<-M0  
  M0diff<-dcast(M0diff, Species+Site+Pair+month~B2015) # cast()
  M0diff$diff<-M0diff$N-M0diff$Y

# Subset Month1.5 
M1.5<-subset(naData, month=="Month1.5") # subset month 1.5 
M1.5_NA <- M1.5 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 113_114 remove
  # subset month 1.5
  M1.5a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M1.5<-pData[M1.5a]
  M1.5<-subset(M1.5, month=="Month1.5")
  M1.5<-na.omit(M1.5) # remove the NA colums
  M1.5<-M1.5[!(M1.5$Pair=="113_114"),]
 #DIFFS
  M1.5diff<-M1.5  
  M1.5diff<-dcast(M1.5diff, Species+Site+Pair+month~B2015) # cast()
  M1.5diff$diff<-M1.5diff$N-M1.5diff$Y
  
## Subset Month3
M3<-subset(naData, month=="Month3") # subset month 3
M3_NA <- M3 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 107_108, 13_14, 213_214, 215_216, 217_218, 285_286, 31_32 <- both not found
#7_8, 113_114,255_256, 259_260 <- Pairs where only one was not found :/

  # subset month 3
  M3a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M3<-pData[M3a]
  M3<-subset(M3, month=="Month3")
  M3<-na.omit(M3) # remove the NA colums
  # remove 7_8, 113_114,255_256, 259_260 <- Pairs where only one was not found :/
  M3<-M3[!(M3$Pair=="7_8"),]
  M3<-M3[!(M3$Pair=="113_114"),]
  M3<-M3[!(M3$Pair=="255_256"),]
  M3<-M3[!(M3$Pair=="259_260"),]
  
  #DIFFS
  M3diff<-M3  
  M3diff<-dcast(M3diff, Species+Site+Pair+month~B2015) # cast()
  M3diff$diff<-M3diff$N-M3diff$Y

## Subset Month 4.5
M4.5<-subset(naData, month=="Month4.5") # subset month 3
M4.5_NA <- M4.5 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
# 285_286,9_10, 209_210, 37)38, 255_256 <- both not found
# 7_8, 113_114,255_256, 259_260, 263_264, 267_268, 82_83 <- Pairs where only one was not found not 

  # subset month 4.5
  M4.5a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M4.5<-pData[M4.5a]
  M4.5<-subset(M4.5, month=="Month4.5")
  M4.5<-na.omit(M4.5) # remove the NA colums
  # remove # 7_8, 113_114,255_256, 259_260, 263_264, 267_268, 82_83
  M4.5<-M4.5[!(M4.5$Pair=="7_8"),]
  M4.5<-M4.5[!(M4.5$Pair=="113_114"),]
  M4.5<-M4.5[!(M4.5$Pair=="255_256"),]
  M4.5<-M4.5[!(M4.5$Pair=="259_260"),]
  M4.5<-M4.5[!(M4.5$Pair=="263_264"),]
  M4.5<-M4.5[!(M4.5$Pair=="267_268"),]
  M4.5<-M4.5[!(M4.5$Pair=="82_83"),]
 
  #DIFFS
  M4.5diff<-M4.5  
  M4.5diff<-dcast(M4.5diff, Species+Site+Pair+month~B2015) # cast()
  M4.5diff$diff<-M4.5diff$N-M4.5diff$Y
  

# Subset month 6
M6<-subset(naData, month=="Month6")
M6_NA <- M6 %>% filter_all(any_vars(is.na(.))) #pull out NAs to see what rows you need to get rid of 
#  113_114,13_14, 17_18, 213_214, 215_216, 217_218, 223_224,82_83
#<- 111_112, 115_116, 119_120, 207_208, 255_256, 259_260, 263_264,267_269 <-Pairs where only one was not found

  # Subset month 6
  M6a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M6<-pData[M6a]
  M6<-subset(M6, month=="Month6")
  M6<-na.omit(M6) # remove the NA colums
  # remove 111_112, 115_116, 119_120, 207_208, 255_256, 259_260, 263_264, 267_269 <-Pairs where only one was not found :/
  M6<-M6[!(M6$Pair=="111_112"),] # remove the other partner
  M6<-M6[!(M6$Pair=="115_116"),]
  M6<-M6[!(M6$Pair=="119_120"),]
  M6<-M6[!(M6$Pair=="207_208"),]
  M6<-M6[!(M6$Pair=="255_256"),]
  M6<-M6[!(M6$Pair=="259_260"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  M6<-M6[!(M6$Pair=="267_268"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  
  #DIFFS
  M6diff<-M6  
  M6diff<-dcast(M6diff, Species+Site+Pair+month~B2015) # cast()
  M6diff$diff<-M6diff$N-M6diff$Y

#make dfs  
pigPairs<-rbind(M0,M1.5,M3,M4.5,M6)
PairDat<-pigPairs
pigPairs06<-rbind(M0,M1.5,M3,M4.5,M6)
pigPairs03<-rbind(M0,M1.5,M3)
pigPairs36<-rbind(M3,M4.5,M6)

#Mean bleaching scores across the entire project

#calc average bleaching score per grouping by time point
pairmeans <- ddply(PairDat, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

pairmeans$code<-paste0(pairmeans$Site, pairmeans$B2015, pairmeans$Species)
pairmeans #display table
```

*Montipora capitata* 
```{r}
#subset for Mcap only
McapPairplot<-subset(pairmeans, Species=="Montipora capitata")


PairplotMcapB<-ggplot(data=McapPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.2) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Montipora capitata Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));PairplotMcapB
```

*Porites compressa* 
```{r}
#subset for Pcomp only
PcompPairplot<-subset(pairmeans, Species=="Porites compressa")

PairplotPcompB<-ggplot(data=PcompPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Porites compressa Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic"));PairplotPcompB
```

#Mcap only full model Pairs 0-6 months
SI Table 11
Model: `score~Site*B2015*Species`  
```{r}
#set up Pairs data
Mcappig<-subset(PairDat, Species =="Montipora capitata")            

#Month0
MpigPair0<-subset(Mcappig, month =="Month0")           
MpigPair0_mod1<-aov(score~Site*B2015, MpigPair0)
summary(MpigPair0_mod1)
TukeyHSD(MpigPair0_mod1) 

#Month1.5
MpigPair1.5<-subset(Mcappig, month =="Month1.5")
MpigPair1.5_mod1<-aov(score~Site*B2015, MpigPair1.5)
summary(MpigPair1.5_mod1)
TukeyHSD(MpigPair1.5_mod1)

#Month3
MpigPair3<-subset(Mcappig, month =="Month3")
MpigPair3_mod1<-aov(score~Site*B2015, MpigPair3)
summary(MpigPair3_mod1)
TukeyHSD(MpigPair3_mod1)

#Month4.5
MpigPair4.5<-subset(Mcappig, month =="Month4.5")
MpigPair4.5_mod1<-aov(score~Site*B2015, MpigPair4.5)
summary(MpigPair4.5_mod1)
TukeyHSD(MpigPair4.5_mod1)

#Month6
MpigPair6<-subset(Mcappig, month =="Month6")
# anova
MpigPair6_mod1<-aov(score~Site*B2015, MpigPair6)
summary(MpigPair6_mod1)
TukeyHSD(MpigPair6_mod1)
```

###Pcomp only full model Pairs 0-6 months
Model: `score~Site*B2015`  
SI Table 11
```{r}
Pcomppig<- subset(PairDat, Species == "Porites compressa") 

#Month 0
PpigPair0<-subset(Pcomppig, month =="Month0")            
PpigPair0_mod1<-aov(score~Site*B2015, PpigPair0)
summary(PpigPair0_mod1)
TukeyHSD(PpigPair0_mod1) #post hoc

#Month1.5
PpigPair1.5<-subset(Pcomppig, month =="Month1.5")
PpigPair1.5_mod1<-aov(score~Site*B2015, PpigPair1.5)
summary(PpigPair1.5_mod1)
TukeyHSD(PpigPair1.5_mod1)

#Month3
PpigPair3<-subset(Pcomppig, month =="Month3")
PpigPair3_mod1<-aov(score~Site*B2015, PpigPair3)
summary(PpigPair3_mod1)

#Month4.5
PpigPair4.5<-subset(Pcomppig, month =="Month4.5")
PpigPair4.5_mod1<-aov(score~Site*B2015, PpigPair4.5)
summary(PpigPair4.5_mod1)
TukeyHSD(PpigPair4.5_mod1)

#Month6
PpigPair6<-subset(Pcomppig, month =="Month6")
PpigPair6_mod1<-aov(score~Site*B2015, PpigPair6)
summary(PpigPair6_mod1)
```

0-3 and 0-6 month differences
Pairs data
score~month*B2015*lagoon
*M. capitata* Pairs
```{r}
Mcappig03Pairs<-subset(pigPairs03, Species=="Montipora capitata")
Mcappig06Pairs<-subset(pigPairs06, Species=="Montipora capitata")

#0-3
Mcappig03Pairs_mod<-aov(score~month*Site*B2015, Mcappig03Pairs)
summary(Mcappig03Pairs_mod)

#0-6
Mcappig06Pairs_mod<-aov(score~month*Site*B2015, Mcappig06Pairs)
summary(Mcappig06Pairs_mod)
```

Pcomp Pairs
```{r}
Pcomppig03Pairs<-subset(pigPairs03, Species=="Porites compressa")
Pcomppig06Pairs<-subset(pigPairs06, Species=="Porites compressa")
Pcomppig36Pairs<-subset(pigPairs36, Species=="Porites compressa")

#0-3
Pcomppig03Pairs_mod<-aov(score~month*Site*B2015, Pcomppig03Pairs)
summary(Pcomppig03Pairs_mod)
#06
Pcomppig06Pairs_mod<-aov(score~month*Site*B2015, Pcomppig06Pairs)
summary(Pcomppig06Pairs_mod)
```


# DIFF DATA
Differences (nonbleached-bleached) for pairs
```{r}
# data = combine all the M#diffs into a single dataframe 
DiffData06<-rbind(M0diff, M1.5diff, M3diff, M4.5diff, M6diff)
DiffData03<-rbind(M0diff, M1.5diff, M3diff)
DiffData36<-rbind(M3diff, M4.5diff, M6diff)
```

Mean bleaching scores across the entire project
```{r, results=TRUE}
#calc average bleaching score per grouping by time point
diffmeans <- ddply(DiffData06, c("Site", "Species", "month"), summarise, #
                   N    = length(diff[!is.na(diff)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(diff, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(diff, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(diff, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

diffmeans$code<-paste0(diffmeans$Site, diffmeans$B2015, diffmeans$Species)
diffmeans #display table
```

SI Figure 2
*Montipora capitata* diff 
```{r}
#subset for Mcap only
Mcapdiffplot<-subset(diffmeans, Species=="Montipora capitata")

Mcapdiffplota<-ggplot(data=Mcapdiffplot, aes(x=month, y=mean, group = code, color=Site)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=Site, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Difference score") +
    xlab("") + #Label the X Axis
    ylim(-.2, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Montipora capitata")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));Mcapdiffplota
Mcapdiffplota<-Mcapdiffplota + theme(legend.position = "none") #remove legend
Mcapdiffplota
```

###Porites compressa diff 
```{r}
#subset for Mcap only
Pcomdiffplot<-subset(diffmeans, Species=="Porites compressa")

Pcompdiffplota<-ggplot(data=Pcomdiffplot, aes(x=month, y=mean, group = code, color=Site)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=Site, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Difference score") +
    xlab("") + #Label the X Axis
    ylim(-.2, 2.5) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Porites compressa")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));Pcompdiffplota
Pcompdiffplota<-Pcompdiffplota + theme(legend.position = "none") #remove legend

```

SI Figure 2
```{r}
grid.arrange(Mcapdiffplota, Pcompdiffplota, nrow = 1) 
```

SI Table 1
Mcap diff~Lagoon
```{r}
DiffData03_Mcap<-subset(DiffData03, Species=="Montipora capitata")
DiffData06_Mcap<-subset(DiffData06, Species=="Montipora capitata")
DiffData36_Mcap<-subset(DiffData36, Species=="Montipora capitata")

#Month0
dMpigPair03<-subset(DiffData06_Mcap, month =="Month0")          
dMpigPair03_mod1<-aov(diff~Site, dMpigPair03)
summary(dMpigPair03_mod1)

#Month1.5
dMpigPair1.5<-subset(DiffData06_Mcap, month =="Month1.5")
dMpigPair1.5_mod1<-aov(diff~Site, dMpigPair1.5)
summary(dMpigPair1.5_mod1)

#Month3
dMpigPair3<-subset(DiffData06_Mcap, month =="Month3")
dMpigPair3_mod1<-aov(diff~Site, dMpigPair3)
summary(dMpigPair3_mod1)
TukeyHSD(dMpigPair3_mod1)

#Month4.5
dMpigPair4.5<-subset(DiffData06_Mcap, month =="Month4.5")
dMpigPair4.5_mod1<-aov(diff~Site, dMpigPair4.5)
summary(dMpigPair4.5_mod1)

#Month6
dMpigPair6<-subset(DiffData06_Mcap, month =="Month6")
dMpigPair6_mod1<-aov(diff~Site, dMpigPair6)
summary(dMpigPair6_mod1)
TukeyHSD(dMpigPair6_mod1)
```

diffs 0-3 months and 0-6 months
*M. capitata*
```{r}
# 0-3
d_Mcap03_mod1<-aov(diff~Site*month, DiffData03_Mcap)
summary(d_Mcap03_mod1)
TukeyHSD(d_Mcap03_mod1)

# 0-6
d_Mcap06_mod1<-aov(diff~Site*month, DiffData06_Mcap)
summary(d_Mcap06_mod1)
TukeyHSD(d_Mcap06_mod1)
```

*P. compressa*
```{r}
DiffData03_Pcomp<-subset(DiffData03, Species=="Porites compressa")
DiffData06_Pcomp<-subset(DiffData06, Species=="Porites compressa")

#Month0
dPpigPair03<-subset(DiffData03_Pcomp, month =="Month0")          
dPpigPair03_mod1<-aov(diff~Site, dPpigPair03)
summary(dPpigPair03_mod1)
TukeyHSD(dPpigPair03_mod1)

#Month1.5
dPpigPair1.5<-subset(DiffData06_Pcomp, month =="Month1.5")
dPpigPair1.5_mod1<-aov(diff~Site, dPpigPair1.5)
summary(dPpigPair1.5_mod1)

#Month3
dPpigPair3<-subset(DiffData06_Pcomp, month =="Month3")
dPpigPair3_mod1<-aov(diff~Site, dPpigPair3)
summary(dPpigPair3_mod1)

#Month4.5
dPpigPair4.5<-subset(DiffData06_Pcomp, month =="Month4.5")
dPpigPair4.5_mod1<-aov(diff~Site, dPpigPair4.5)
summary(dPpigPair4.5_mod1)

#Month6
dPpigPair6<-subset(DiffData06_Pcomp, month =="Month6")
dPpigPair6_mod1<-aov(diff~Site, dPpigPair6)
summary(dPpigPair6_mod1)
TukeyHSD(dPpigPair6_mod1)
```

diffs 0-3 months and 0-6 months
*P. compressa*
```{r}
# 0-3
d_Pcomp03_mod1<-aov(diff~Site*month, DiffData03_Pcomp)
summary(d_Pcomp03_mod1)
TukeyHSD(d_Pcomp03_mod1)

# 0-6
d_Pcomp06_mod1<-aov(diff~Site*month, DiffData06_Pcomp)
summary(d_Pcomp06_mod1)
TukeyHSD(d_Pcomp06_mod1)

# 3-6
d_Pcomp36_mod1<-aov(diff~Site*month, DiffData36_Pcomp)
summary(d_Pcomp36_mod1)
TukeyHSD(d_Pcomp36_mod1)
```
