---
title: "Pigmentation Score"
author: "Shayle Matsuda"
date: "08/08/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
update:"04/30/2020"

Matsuda, S, Huffmyer, A, Lenz, E, Davidson, J, Hancock, J, Pryzybylowski, A, Innis, T, Gates, RD, and KL Barott. Coral bleaching susceptibility is predictive of subsequent mortality within but not between coral species. 

Pigmentation score for *M. capitata* and *P. compressa* tagged colonies at the Inner Lagoon and Outer Lagoon after the 2015 coral bleaching event in Kaneohe Bay.  Tagged individuals were given a pigment score based on their visual color: 1) white (>80% of colony white with no visible pigmentation); 2) pale (>10% colony affected by pigment loss); or 3) fully pigmented (<10% colony with any pale coloration). 

```{r setup, include=FALSE,results="hide"}
knitr::opts_chunk$set(echo = TRUE)
```

Load in and set up dfs: 
```{r, results="hide", message=FALSE}
#Set up: load libraries

#rm(list=ls(all=TRUE)) 
library("reshape2") #reshape data
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library("multcomp") 
library("multcompView")
library(glmmTMB)
library(lme4)
library(lmerTest)
library(tidyverse)

```

# Bleaching data
```{r, results="hide", message=FALSE}
bData<-read.csv("Bleaching_FINAL.csv") #read in bleaching df
bData$Site<-as.factor(bData$Site) #site as factor
bData$Pair<-as.factor(bData$Pair) #pair ID as factor

#organize data
bALL<-bData
  #melt
  bALLData1<-melt(data = bALL, measure.vars = c("Month0","Month1.5","Month3","Month4.5", "Month6","Month10","Month18","Month24"))   
  colnames(bALLData1)[colnames(bALLData1)=="variable"] <- "month"   # change col name from 'variable' to 'month'
  colnames(bALLData1)[colnames(bALLData1)=="value"] <- "score"   # change col name from 'value' to 'score'
  bALLData1$group<-paste(bALLData1$Species, bALLData1$Site, bALLData1$B2015, bALLData1$month) # add column for groups for figures only
#write.csv(bALL, "PigmentationBNB_20191122.csv")

#Remove all NON PAIRS and all NAs
  bALLData<-na.omit(bALLData1) # remove the NA rows (this removes individuals and NAs from pairs with no score at particular timepoints (due to not finding the coral))
  
  #Change month into TIME numeric for plots
  bALLData$time[bALLData$month=="Month0"]<-0
  bALLData$time[bALLData$month=="Month1.5"]<-1.5
  bALLData$time[bALLData$month=="Month3"]<-3
  bALLData$time[bALLData$month=="Month4.5"]<-4.5
  bALLData$time[bALLData$month=="Month6"]<-6
  bALLData$time[bALLData$month=="Month10"]<-10
  bALLData$time[bALLData$month=="Month18"]<-18
  bALLData$time[bALLData$month=="Month19"]<-19
  bALLData$time[bALLData$month=="Month24"]<-24

#Subsets of df for Mcap and Pcomp
 
Mcap_df<-subset(bALLData, Species=="Montipora capitata")   #Mcap only
Pcomp_df<-subset(bALLData, Species=="Porites compressa")   #Pcomp only

#Mean bleaching scores full df
#calc average bleaching  score per grouping by time point

ballmeans <- ddply(bALLData, c("Site", "Species", "B2015", "time"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
ballmeans #display table

ballmeans$code<-paste0(ballmeans$Site, ballmeans$B2015, ballmeans$Species)
ballmeans #display table
#write.csv(ballmeans,"ballmeans.csv"
```

#Figure 5
Bleaching score means 0-24 months by species. 
*Montipora capitata* df
```{r, results="hide", message=FALSE} 
#subset for Mcap only 
Mcap024Allplot<-subset(ballmeans, Species=="Montipora capitata")     #subset Mcap
  
#subset by reef
#PR13 
Mcap024AllplotPR13<-subset(Mcap024Allplot, Site==13)
#PR4
Mcap024AllplotPR4<-subset(Mcap024Allplot, Site==4)
```

*M. capitata* Outer Lagoon Plot
```{r, results="hide", message=FALSE}
ballplotMcapPR13<-ggplot(data=Mcap024AllplotPR13, aes(x=time, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="black", size=5)+
  scale_shape_manual(values=c(16,1))+
  geom_line(aes(color=B2015,linetype=B2015))+        #adding size=x here only blew it up
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold",hjust = 0.5));ballplotMcapPR13
McapOuter<-ballplotMcapPR13 + theme(legend.position = "none") #remove legend
```

*M. capitata* Inner Lagoon Plot
```{r, results="hide", message=FALSE}
ballplotMcapPR4<-ggplot(data=Mcap024AllplotPR4, aes(x=time, y=mean, group = code, color=B2015),show.legend = FALSE) + #color by bleaching hist. 
 geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Montipora capitata\nColor score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold", hjust = 0.5));ballplotMcapPR4
McapInner<-ballplotMcapPR4 + theme(legend.position = "none") #remove legend
```

*Porites compressa*  df setup
```{r, results="hide", message=FALSE}
Pcomp024Allplot<-subset(ballmeans, Species=="Porites compressa")

#subset by reef 
Pcomp024AllplotPR13<-subset(Pcomp024Allplot, Site==13)
Pcomp024AllplotPR4<-subset(Pcomp024Allplot, Site==4)
```

*P. compress* Outer Lagoon Plot
```{r, results="hide", message=FALSE}
b024plotPcompPR13<-ggplot(data=Pcomp024AllplotPR13, aes(x=time, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Color score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Porites compressa Bleaching Scores times 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR13

PcompOuter<-b024plotPcompPR13+ theme(legend.position = "none") #remove legend
```

*P. compressa* Inner Lagoon
```{r, results="hide", message=FALSE}
b024plotPcompPR4<-ggplot(data=Pcomp024AllplotPR4, aes(x=time, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  
  scale_shape_manual(values=c( 16,1))+
  
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
  ylab("Porites compressa\nColor score") +
    xlab("Recovery Time (times)") + #Label the X Axis
    ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #Sggtitle("Porites compressa Bleaching Scores times 0-24")+
          theme(plot.title = element_text(size=20,face = "italic"));b024plotPcompPR4


PcompInner<-b024plotPcompPR4+ theme(legend.position = "none") #remove legend
```

Figure 5:
```{r}
grid.arrange(McapInner, McapOuter, PcompInner,PcompOuter, nrow = 2)
```

# Bleaching Recovery Models
##Table 2: Color score recovery  
*M. capitata* lmer(score~month*B2015*lagoon+(1|TagID) 
```{r}
#full model
Mcap_df_mod<-lmer(score~B2015*Site*month+(1|TagID), data=Mcap_df)
summary(Mcap_df_mod)
anova(Mcap_df_mod, type=2) #pvals

 #posthoc
  emm<-emmeans(Mcap_df_mod,  ~ month)
  pairs(emm)
  #cld(emm)
  # Month4.5 - Month10  p=0.7754 
  
 emm<-emmeans(Mcap_df_mod,  ~ B2015*month)
  pairs(emm)
  cld(emm)
  #Y,Month0 - Y,Month1.5   -0.74657 0.0719 412 -10.377 <.0001 
  #Y,Month1.5 - Y,Month3   -0.37874 0.0756 411  -5.007 0.0001
  #Y,Month3 - Y,Month4.5   -0.01643 0.0765 414  -0.215 1.0000
  #Y,Month4.5 - Y,Month6   -0.08619 0.0804 412  -1.073 0.9995
  #Y,Month6 - Y,Month10     0.17169 0.0847 409   2.026 0.8094
  #Y,Month10 - Y,Month18   -0.12441 0.0949 417  -1.310 0.9954
  #Y,Month18 - Y,Month24   -0.33618 0.1108 427  -3.035 0.1591 
```

*P. compressa* lmer(score~month*B2015*lagoon+(1|TagID) 
```{r}
#pcomp full model
Pcomp_df_mod2<-lmer(score~B2015*Site*month+(1|TagID), data=Pcomp_df)
summary(Pcomp_df_mod2)
anova(Pcomp_df_mod2, type=2) #pvals

  emm<-emmeans(Pcomp_df_mod2,  ~ B2015*Site*month)
  pairs(emm)

#Tukey post hoc: ~Lagoon*Month

   emm<-emmeans(Pcomp_df_mod2,  ~ Site*month)
  pairs(emm)
      # 4,Month0 - 13,Month0      -0.190476 0.0583 545  -3.268 0.0844 
      #4,Month1.5 - 13,Month1.5  -0.049478 0.0590 545  -0.839 1.0000
  #4,Month3 - 13,Month3      -0.022300 0.0598 545  -0.373 1.0000 
  #4,Month4.5 - 13,Month4.5  -0.151734 0.0610 545  -2.488 0.4822
  #4,Month6 - 13,Month6       0.000289 0.0590 545   0.005 1.0000 
  #4,Month10 - 13,Month10    -0.024925 0.0602 545  -0.414 1.0000 
  #4,Month18 - 13,Month18    -0.033845 0.0694 547  -0.488 1.0000
  #4,Month24 - 13,Month24    -0.074803 0.0779 548  -0.960 0.9999 

     emm<-emmeans(Pcomp_df_mod2,  ~ B2015*month)
  pairs(emm)
  #Y,Month0 - Y,Month1.5   -1.475103 0.0577 473 -25.562 <.0001
  #Y,Month1.5 - Y,Month3   -0.447667 0.0584 473  -7.661 <.0001
  #Y,Month3 - Y,Month4.5   -0.001545 0.0593 476  -0.026 1.0000 
  #Y,Month4.5 - Y,Month6   -0.075574 0.0589 477  -1.283 0.9963
  #Y,Month6 - Y,Month10    -0.000723 0.0584 473  -0.012 1.0000 
  #Y,Month10 - Y,Month18    0.073924 0.0642 490   1.151 0.9989
  #Y,Month18 - Y,Month24   -0.071078 0.0734 513  -0.969 0.9999 
  
    emm<-emmeans(Pcomp_df_mod2,  ~ B2015*month*Site)
  pairs(emm)
  #Y,Month0,4 - Y,Month1.5,4     -1.350207 0.0811 474 -16.642 <.0001 
  #
  
```

#Supplemental Table 7
Models at each timepoint
*M. capitata* lm()
```{r, results="hide", message=FALSE}  
Mcap_0<-subset(Mcap_df, month=="Month0")
Mcap_1.5<-subset(Mcap_df, month=="Month1.5")
Mcap_3<-subset(Mcap_df, month=="Month3")
Mcap_4.5<-subset(Mcap_df, month=="Month4.5")
Mcap_6<-subset(Mcap_df, month=="Month6")
Mcap_10<-subset(Mcap_df, month=="Month10")
Mcap_18<-subset(Mcap_df, month=="Month18")
Mcap_24<-subset(Mcap_df, month=="Month24")

#Month0
Month0_mcap0<-lm(score~Site*B2015, Mcap_0)
summary(Month0_mcap0)
anova(Month0_mcap0)

#Month1.5
Month1.5_mcap1.5<-lm(score~Site*B2015, Mcap_1.5)
summary(Month1.5_mcap1.5)
anova(Month1.5_mcap1.5)


#Month3
Month3_mcap3<-lm(score~Site*B2015, Mcap_3)
summary(Month3_mcap3)
anova(Month3_mcap3)
 emm<-emmeans(Month3_mcap3,  ~ B2015*Site)
  pairs(emm)
  cld(emm, Letter="abcde")
  #N,4 - Y,4     0.6147 0.102 68  6.004  <.0001
  #N,13 - Y,13   0.9412 0.105 68  8.967  <.0001
  
#Month4.5
Month4.5_mcap4.5<-lm(score~Site*B2015, Mcap_4.5)
summary(Month4.5_mcap4.5)
anova(Month4.5_mcap4.5)
 emm<-emmeans(Month4.5_mcap4.5,  ~ B2015*Site)
  pairs(emm)
  #N,4 - Y,4     0.4494 0.134 72  3.349  0.0069
  #N,13 - Y,13   0.8421 0.130 72  6.500  <.0001

  #Month6
Month6_mcap6<-lm(score~Site*B2015, Mcap_6)
summary(Month6_mcap6)
anova(Month6_mcap6)
emm<-emmeans(Month6_mcap6,  ~ Site)
  pairs(emm)
emm<-emmeans(Month6_mcap6,  ~ B2015*Site)
  pairs(emm)
  # N,4 - Y,4     0.3111 0.140 58  2.229  0.1275
  #N,13 - Y,13   0.8000 0.148 58  5.392  <.0001
  
#Month10
Month10_mcap10<-lm(score~Site*B2015, Mcap_10)
summary(Month10_mcap10)
anova(Month10_mcap10)

#Month18
Month18_mcap18<-lm(score~Site*B2015, Mcap_18)
summary(Month18_mcap18)
anova(Month18_mcap18)

#Month24
Month24_mcap24t<-lm(score~Site*B2015, Mcap_24)
summary(Month24_mcap24t)
anova(Month24_mcap24t)

Mcap24_Inner<-subset(Mcap_24, Site=="4")
```

*P. compressa* lm()
```{r, results="hide", message=FALSE} 
Pcomp_0<-subset(Pcomp_df, month=="Month0")
Pcomp_1.5<-subset(Pcomp_df, month=="Month1.5")
Pcomp_3<-subset(Pcomp_df, month=="Month3")
Pcomp_4.5<-subset(Pcomp_df, month=="Month4.5")
Pcomp_6<-subset(Pcomp_df, month=="Month6")
Pcomp_10<-subset(Pcomp_df, month=="Month10")
Pcomp_18<-subset(Pcomp_df, month=="Month18")
Pcomp_24<-subset(Pcomp_df, month=="Month24")


#Month0
Month0_Pcomp0<-lm(score~Site*B2015, Pcomp_0)
summary(Month0_Pcomp0)
anova(Month0_Pcomp0)

#Month1.5
Month1.5_Pcomp1.5<-lm(score~Site*B2015, Pcomp_1.5)
summary(Month1.5_Pcomp1.5)
anova(Month1.5_Pcomp1.5)

#Month3
Month3_Pcomp3<-lm(score~Site*B2015, Pcomp_3)
summary(Month3_Pcomp3)
anova(Month3_Pcomp3)

#Month4.5
Month4.5_Pcomp4.5<-lm(score~Site*B2015, Pcomp_4.5)
summary(Month4.5_Pcomp4.5)
anova(Month4.5_Pcomp4.5)

#Month6
Month6_Pcomp6<-lm(score~Site*B2015,Pcomp_6)
summary(Month6_Pcomp6)
anova(Month6_Pcomp6)

#Month10
Month10_Pcomp10<-lm(score~Site*B2015, Pcomp_10)
summary(Month10_Pcomp10)
anova(Month10_Pcomp10)

#Month18
Month18_Pcomp18<-lm(score~Site*B2015,Pcomp_18)
summary(Month18_Pcomp18)
anova(Month18_Pcomp18)

#Month24
Month24_Pcomp24<-lm(score~Site*B2015,Pcomp_24)
summary(Month24_Pcomp24)
anova(Month24_Pcomp24)
```

# Pairs subset - bleaching recovery models
Pair subset of full dataset with only colonies when both partners of a pair are found at each time point)    
```{r, results="hide", message=FALSE}
naData<-bALLData   #subset only pairs
pData<-bALLData   #dataset to work from (pull unwanted from)

# Identify which pairs have a partner not found by month:

# Subset Month0 (two colonies missing values, checked datasheet)
M0<-subset(naData, month=="Month0") # subset month 0

  # subset month 0
  M0a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M0<-pData[M0a]
  M0<-subset(M0, month=="Month0")

    #For DIFFS df
  M0diff<-M0  
  M0diff<-dcast(M0diff, Species+Site+Pair+month~B2015) # cast()
  M0diff$diff<-M0diff$N-M0diff$Y

# Subset Month1.5 
M1.5<-subset(naData, month=="Month1.5") # subset month 1.5 
  # subset month 1.5
  M1.5a<-c("TagID","Pair","Species", "Site","B2015","month","score") # which columns to pull
  M1.5<-pData[M1.5a]
  M1.5<-subset(M1.5, month=="Month1.5")
  M1.5<-M1.5[!(M1.5$Pair=="113_114"),]
 #DIFFS 
  M1.5diff<-M1.5  
  M1.5diff<-dcast(M1.5diff, Species+Site+Pair+month~B2015) # cast()
  M1.5diff$diff<-M1.5diff$N-M1.5diff$Y
  
## Subset Month3
M3<-subset(naData, month=="Month3") # subset month 3
  M3a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M3<-pData[M3a]
  M3<-subset(M3, month=="Month3")
  M3<-M3[!(M3$Pair=="7_8"),]
  M3<-M3[!(M3$Pair=="113_114"),]
  M3<-M3[!(M3$Pair=="255_256"),]
  M3<-M3[!(M3$Pair=="259_260"),]
  
  #DIFFS
  M3diff<-M3  
  M3diff<-dcast(M3diff, Species+Site+Pair+month~B2015) # cast()
  M3diff$diff<-M3diff$N-M3diff$Y

## Subset Month 4.5
M4.5<-subset(naData, month=="Month4.5") # subset month 3
  M4.5a<-c("TagID","Pair","Species", "Site","B2015","month", "score")
  M4.5<-pData[M4.5a]
  M4.5<-subset(M4.5, month=="Month4.5")
  M4.5<-na.omit(M4.5) # remove the NA colums
  M4.5<-M4.5[!(M4.5$Pair=="7_8"),]
  M4.5<-M4.5[!(M4.5$Pair=="113_114"),]
  M4.5<-M4.5[!(M4.5$Pair=="255_256"),]
  M4.5<-M4.5[!(M4.5$Pair=="259_260"),]
  M4.5<-M4.5[!(M4.5$Pair=="263_264"),]
  M4.5<-M4.5[!(M4.5$Pair=="267_268"),]
  M4.5<-M4.5[!(M4.5$Pair=="82_83"),]
 
  #DIFFS
  M4.5diff<-M4.5  
  M4.5diff<-dcast(M4.5diff, Species+Site+Pair+month~B2015) # cast()
  M4.5diff$diff<-M4.5diff$N-M4.5diff$Y
  

# Subset month 6
M6<-subset(naData, month=="Month6")
  M6a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M6<-pData[M6a]
  M6<-subset(M6, month=="Month6")
  M6<-M6[!(M6$Pair=="111_112"),] # remove the other partner
  M6<-M6[!(M6$Pair=="115_116"),]
  M6<-M6[!(M6$Pair=="119_120"),]
  M6<-M6[!(M6$Pair=="207_208"),]
  M6<-M6[!(M6$Pair=="255_256"),]
  M6<-M6[!(M6$Pair=="259_260"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  M6<-M6[!(M6$Pair=="267_268"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  
  #DIFFS
  M6diff<-M6  
  M6diff<-dcast(M6diff, Species+Site+Pair+month~B2015) # cast()
  M6diff$diff<-M6diff$N-M6diff$Y
  
# Subset month 10
M10<-subset(naData, month=="Month10")
  M10a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M10<-pData[M10a]
  M10<-subset(M10, month=="Month10")
  M10<-M10[!(M10$Pair=="253_254"),] # remove the other partner
  M10<-M10[!(M10$Pair=="259_260"),]
  M10<-M10[!(M10$Pair=="113_114"),]
  M10<-M10[!(M10$Pair=="161_262"),]
  M10<-M10[!(M10$Pair=="263_264"),]
  M10<-M10[!(M10$Pair=="267_268"),]
  M10<-M10[!(M10$Pair=="271_272"),]
  M10<-M10[!(M10$Pair=="273_274"),]
  M10<-M10[!(M10$Pair=="275_276"),]
  M10<-M10[!(M10$Pair=="277_278"),]
  M10<-M10[!(M10$Pair=="31_32"),]
  M10<-M10[!(M10$Pair=="261_262"),]
  #DIFFS
  M10diff<-M10  
  M10diff<-dcast(M10diff, Species+Site+Pair+month~B2015) # cast()
  M10diff$diff<-M10diff$N-M10diff$Y
  
  # Subset month 18
M18<-subset(naData, month=="Month18")
  M18a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M18<-pData[M18a]
  M18<-subset(M18, month=="Month18")
  M18<-M18[!(M18$Pair=="253_254"),] # remove the other partner
  M18<-M18[!(M18$Pair=="259_260"),]
  M18<-M18[!(M18$Pair=="161_262"),]
  M18<-M18[!(M18$Pair=="263_264"),]
  M18<-M18[!(M18$Pair=="267_268"),]
  M18<-M18[!(M18$Pair=="271_272"),]
  M18<-M18[!(M18$Pair=="273_274"),]
  M18<-M18[!(M18$Pair=="275_276"),]
  M18<-M18[!(M18$Pair=="277_278"),]
  M18<-M18[!(M18$Pair=="31_2"),]
  M18<-M18[!(M18$Pair=="115_116"),]
  M18<-M18[!(M18$Pair=="119_120"),]
  M18<-M18[!(M18$Pair=="261_262"),]
  M18<-M18[!(M18$Pair=="269_270"),]
  M18<-M18[!(M18$Pair=="101_102"),]
  M18<-M18[!(M18$Pair=="117_118"),]
  M18<-M18[!(M18$Pair=="265_266"),]
  M18<-M18[!(M18$Pair=="19_20"),]
  M18<-M18[!(M18$Pair=="205_206"),]
  M18<-M18[!(M18$Pair=="223_224"),]
  M18<-M18[!(M18$Pair=="39_40"),]
  
  #DIFFS
  M18diff<-M18  
  M18diff<-dcast(M18diff, Species+Site+Pair+month~B2015) # cast()
  M18diff$diff<-M18diff$N-M18diff$Y
  
  # Subset month 24
M24<-subset(naData, month=="Month24")
  M24a<-c("TagID","Pair","Species", "Site","B2015","month","score")
  M24<-pData[M24a]
  M24<-subset(M24, month=="Month24")
  M24<-M24[!(M24$Pair=="107_108"),] # remove the other partner
  M24<-M24[!(M24$Pair=="13_14"),]
  M24<-M24[!(M24$Pair=="223_224"),]
  M24<-M24[!(M24$Pair=="231_232"),]
  M24<-M24[!(M24$Pair=="261_262"),]
  M24<-M24[!(M24$Pair=="263_264"),]
  M24<-M24[!(M24$Pair=="265_266"),]
  M24<-M24[!(M24$Pair=="283_284"),]
  M24<-M24[!(M24$Pair=="289_290"),]
  M24<-M24[!(M24$Pair=="31_32"),]
  M24<-M24[!(M24$Pair=="33_34"),]
  M24<-M24[!(M24$Pair=="39_40"),]
  M24<-M24[!(M24$Pair=="76_77"),]
  M24<-M24[!(M24$Pair=="78_79"),]

  #DIFFS
  M24diff<-M24  
  M24diff<-dcast(M24diff, Species+Site+Pair+month~B2015) # cast()
  M24diff$diff<-M24diff$N-M24diff$Y
  
#make dfs   
PairDat<-rbind(M0,M1.5,M3,M4.5,M6,M10,M18,M24)
 
#Mean bleaching scores across the entire project

#calc average bleaching score per grouping by time point
pairmeans <- ddply(PairDat, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#ballmeans #display table

pairmeans$code<-paste0(pairmeans$Site, pairmeans$B2015, pairmeans$Species)
pairmeans #display table
```

*Montipora capitata*  plot pairs
```{r}
#subset for Mcap only
McapPairplot<-subset(pairmeans, Species=="Montipora capitata")

PairplotMcapB<-ggplot(data=McapPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.2) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Montipora capitata Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));PairplotMcapB
```

*Porites compressa* plot pairs
```{r}
#subset for Pcomp only 
PcompPairplot<-subset(pairmeans, Species=="Porites compressa")

PairplotPcompB<-ggplot(data=PcompPairplot, aes(x=month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Site), size=3)+
  geom_line(aes(color=B2015, linetype=Site))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
  ylab("Color score") +
    xlab("") + #Label the X Axis
    ylim(1, 3.3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Pairs Porites compressa Bleaching Scores months 0-10")+
          theme(plot.title = element_text(size=20,face = "italic"));PairplotPcompB
```
#Mcap: Pairs 
SI Table 1 
Model: lmer(score~Site*B2015*Species+(1|Pair)+(1|TagID))
```{r}
#set up Pairs data
Mcappig<-subset(PairDat, Species =="Montipora capitata")            

#All lmer
MpigPair_all_mod1<-lmer(score~Site*B2015*month+(1|TagID)+(1|Pair), Mcappig)
summary(MpigPair_all_mod1)
anova(MpigPair_all_mod1, type=2)
```
By month:
```{r}
#Month0
MpigPair0<-subset(Mcappig, month =="Month0")           
MpigPair0_mod1<-lm(score~Site*B2015, MpigPair0)
summary(MpigPair0_mod1)
anova(MpigPair0_mod1)

#Month1.5
MpigPair1.5<-subset(Mcappig, month =="Month1.5")
MpigPair1.5_mod1<-lm(score~Site*B2015, MpigPair1.5)
summary(MpigPair1.5_mod1)
anova(MpigPair1.5_mod1)

#Month3
MpigPair3<-subset(Mcappig, month =="Month3")
MpigPair3_mod1<-lm(score~Site*B2015, MpigPair3)
summary(MpigPair3_mod1)
anova(MpigPair3_mod1)

#Month4.5
MpigPair4.5<-subset(Mcappig, month =="Month4.5")
MpigPair4.5_mod1<-lm(score~Site*B2015, MpigPair4.5)
summary(MpigPair4.5_mod1)
anova(MpigPair4.5_mod1)

#Month6
MpigPair6<-subset(Mcappig, month =="Month6")
MpigPair6_mod1<-lm(score~Site*B2015, MpigPair6)
summary(MpigPair6_mod1)
anova(MpigPair6_mod1)

#Month10
MpigPair10<-subset(Mcappig, month =="Month10")
MpigPair10_mod1<-lm(score~Site*B2015, MpigPair10)
summary(MpigPair10_mod1)
anova(MpigPair10_mod1)

#Month18
MpigPair18<-subset(Mcappig, month =="Month18")
MpigPair18_mod1<-lm(score~Site*B2015, MpigPair18)
summary(MpigPair18_mod1)
anova(MpigPair18_mod1)

#Month24
MpigPair24<-subset(Mcappig, month =="Month24")
MpigPair24_mod1<-lm(score~Site*B2015, MpigPair24)
summary(MpigPair24_mod1)
anova(MpigPair24_mod1)
```

#Pcomp: Pairs
lmer/lm()
```{r}
Pcomppig<- subset(PairDat, Species == "Porites compressa") 

#All timepoints lmer
PpigPair_all_mod1<-lmer(score~Site*B2015*month+(1|TagID)+(1|Pair), Pcomppig)
summary(PpigPair_all_mod1)
anova(PpigPair_all_mod1, type=2)
```
By month:
```{r}
#Month 0
PpigPair0<-subset(Pcomppig, month =="Month0")            
PpigPair0_mod1<-lm(score~Site*B2015, PpigPair0)
summary(PpigPair0_mod1)
anova(PpigPair0_mod1) 

#Month1.5
PpigPair1.5<-subset(Pcomppig, month =="Month1.5")
PpigPair1.5_mod1<-lm(score~Site*B2015, PpigPair1.5)
summary(PpigPair1.5_mod1)
anova(PpigPair1.5_mod1)

#Month3
PpigPair3<-subset(Pcomppig, month =="Month3")
PpigPair3_mod1<-lm(score~Site*B2015, PpigPair3)
anova(PpigPair3_mod1)

#Month4.5
PpigPair4.5<-subset(Pcomppig, month =="Month4.5")
PpigPair4.5_mod1<-lm(score~Site*B2015, PpigPair4.5)
summary(PpigPair4.5_mod1)
anova(PpigPair4.5_mod1)

#Month6
PpigPair6<-subset(Pcomppig, month =="Month6")
PpigPair6_mod1<-lm(score~Site*B2015, PpigPair6)
anova(PpigPair6_mod1)

#Month10
PpigPair10<-subset(Pcomppig, month =="Month10")
PpigPair10_mod1<-lm(score~Site*B2015, PpigPair10)
anova(PpigPair10_mod1)

#Month18
PpigPair18<-subset(Pcomppig, month =="Month18")
PpigPair18_mod1<-lm(score~Site*B2015, PpigPair18)
anova(PpigPair18_mod1)

#Month24
PpigPair24<-subset(Pcomppig, month =="Month24")
PpigPair24_mod1<-lm(score~Site*B2015, PpigPair24)
anova(PpigPair24_mod1)
```