---
title: "ColorScore revision"
author: "Shayle Matsuda"
date: "9/25/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
Last updater: 5/20/2020

Matsuda, S, Huffmyer, A, Lenz, E, Davidson, J, Hancock, J, Pryzybylowski, A, Innis, T, Gates, RD, and KL Barott. Coral bleaching susceptibility is predictive of subsequent mortality within but not between coral species.  

Tissue partial mortality analysis for *M. capitata* and *P. compressa* colonies at the Inner Lagoon (PR4) and Outer Lagoon (PR13) after the 2015 coral bleaching event in Kaneohe Bay.  Mortality scores are in % to the nearest 20%

```{r setup, include=FALSE,results="hide"}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results="hide", message=FALSE}
rm(list=ls(all=TRUE)) 
library("reshape2") #reshape data
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library("multcomp") 
library("multcompView")
library(glmmTMB)
library(lme4)
library(lmerTest)
library(tidyverse)

#load data
mData<-read.csv("Mortality_scores_percent.csv")
mData$Site<-as.factor(mData$Site)

#We excluded these pairs as part of the pair analysis because the Bleached coral was only a 2: Porites Outer Bay 233_234, Porites Inner Bay 297_298, Porites Outer Bay 227_228, Montipora capitata Inner Bay 251_252.
```

# Partial mortality   
set up df:
```{r, results="hide", message=FALSE}
#all Months
  mALL<-mData
  #melt
  mALLData1<-melt(data = mALL, measure.vars = c("Month0","Month1.5","Month3","Month4.5", "Month6","Month10","Month18","Month24"))   
  colnames(mALLData1)[colnames(mALLData1)=="variable"] <- "month"   # change col name from 'variable' to 'month'
  colnames(mALLData1)[colnames(mALLData1)=="value"] <- "score"   # change col name from 'value' to 'score'

#Remove all NON PAIRS and all NAs
  mALLData<-na.omit(mALLData1) # remove the NA colums

#Change scores into %
mALLDataPER<-mALLData
mALLDataPER$Score<-mALLDataPER$score*100

#recast for rate cals
mALLData_cast<-dcast(mALLData, TagID+Pair+Species+Site+B2015~month)
```

Mean partial mortality scores:
```{r, results="hide", message=FALSE}
#calc average tissue loss score per grouping by time point
mALLmeans <- ddply(mALLData, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
mALLmeans$code<-paste0(mALLmeans$Site, mALLmeans$B2015, mALLmeans$Species)
```

Partial mortality score means 0-24 months by species/site/bleaching history
```{r, results=TRUE}
#calc average tissue loss score per grouping by time point
mALLmeansPER <- ddply(mALLDataPER, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(Score[!is.na(Score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Score, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
mALLmeansPER$code<-paste0(mALLmeansPER$Site, mALLmeansPER$B2015, mALLmeansPER$Species)
```

#Figure 5, Partial Mortality  
Figure 5, setup df
```{r, results="hide", message=FALSE}
#subset for Mcap only
Mcap024AllplotM<-subset(mALLmeansPER, Species=="Montipora capitata")     #subset Mcap
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month0"]<-0     # new col with month #s only
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month1.5"]<-1.5  
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month3"]<-3 
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month4.5"]<-4.5
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month6"]<-6  
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month10"]<-10  
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month18"]<-18
Mcap024AllplotM$Month[Mcap024AllplotM$month=="Month24"]<-24 

#PR13
Mcap024AllplotPR13M<-subset(Mcap024AllplotM, Site=="13")
#PR4
Mcap024AllplotPR4M<-subset(Mcap024AllplotM, Site==4)
```

*Montipora capitata*  Figure
```{r, results="hide", message=FALSE}
#Mcap Outer
mallplotMcapPR13<-ggplot(data=Mcap024AllplotPR13M, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
    ylim(0, 80) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
          theme(plot.title = element_text(size=20,face = "bold",hjust = 0.5))
McapOuter<-mallplotMcapPR13 + theme(legend.position = "none") #remove legend

#Mcap Inner
ballplotMcapPR4M<-ggplot(data=Mcap024AllplotPR4M, aes(x=Month, y=mean, group = code, color=B2015),show.legend = FALSE) + #color by bleaching hist. 
 geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
    ylim(0, 80) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
 # ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "bold", hjust = 0.5))
McapInnerMORT<-ballplotMcapPR4M + theme(legend.position = "none") #remove legend
```

*Porites compressa* Figures  
```{r, results="hide", message=FALSE}
Pcomp024Allplot<-subset(mALLmeansPER, Species=="Porites compressa")

# change Month# to just number
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month0"]<-0
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month1.5"]<-1.5
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month3"]<-3
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month4.5"]<-4.5
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month6"]<-6
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month10"]<-10
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month18"]<-18
Pcomp024Allplot$Month[Pcomp024Allplot$month=="Month24"]<-24

Pcomp024AllplotPR13<-subset(Pcomp024Allplot, Site==13)
Pcomp024AllplotPR4<-subset(Pcomp024Allplot, Site==4)

#Pcomp PR13 plot
b024plotPcompPR13<-ggplot(data=Pcomp024AllplotPR13, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
    ylim(0,80) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
          theme(plot.title = element_text(size=20,face = "italic"))
PcompOuter<-b024plotPcompPR13+ theme(legend.position = "none") #remove legend

#Pcomp PR4 plot
b024plotPcompPR4<-ggplot(data=Pcomp024AllplotPR4, aes(x=Month, y=mean, group = code, color=B2015)) + #color by bleaching hist. 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1,show.legend = FALSE)+
  geom_point(aes(shape=B2015),fill="mediumblue", size=5)+
  scale_shape_manual(values=c( 16,1))+
  geom_line(aes(color=B2015, linetype=B2015))+
  scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black","black"))+
    scale_x_continuous(breaks = seq(0, 24, by = 6))+
    ylim(0,80) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
          theme(plot.title = element_text(size=20,face = "italic"))
PcompInner<-b024plotPcompPR4+ theme(legend.position = "none") #remove legend
```

Plot all four together: Figure 5 (See color score Rmd for other Fig 5 A-D)
```{r}
grid.arrange(McapInnerMORT, McapOuter, PcompInner,PcompOuter, nrow = 2)
```

#Table 3: Pigmentation score~month*B2015*lagoon  
```{r, results="hide", message=FALSE}
#subset by species
Mcap_mort<-subset(mALLDataPER, Species=="Montipora capitata")
Pcomp_mort<-subset(mALLDataPER, Species=="Porites compressa")
```

Mortality across all timepoints, Table 3
```{r, results="hide", message=FALSE}
#Mcap
Mcap_mort_mod<-lmer(Score~Site*B2015*month+(1|TagID), Mcap_mort)
anova(Mcap_mort_mod, type=2)

#conduct Tukey post hoc test 
emm = emmeans(Mcap_mort_mod, ~ Site*month, adjust="tukey")
#letter display
pairs(emm)

#Pcomp
Pcomp_mort_mod<-lmer(Score~Site*B2015*month+(1|TagID), Pcomp_mort)
anova(Pcomp_mort_mod, type=2)

#conduct Tukey post hoc test 
emm = emmeans(Pcomp_mort_mod, ~ Site*month, adjust="tukey")
#letter display
pairs(emm)
```

#SI Tables 7 and 8   
Tissue Loss  Models  by month, linear mixed effects models
*Montipora capiatata*
```{r} 
#Month0 
Mcap_m0<-subset(Mcap_mort, month=="Month0")
Mcap_m0_mod1<-lm(score~Site*B2015, Mcap_m0)
summary(Mcap_m0_mod1) 
anova(Mcap_m0_mod1) 

# Month1.5
Mcap_m1_5<-subset(Mcap_mort, month=="Month1.5")
Mcap_m1_5_mod1<-lm(score~Site*B2015, Mcap_m1_5)
summary(Mcap_m1_5_mod1)
anova(Mcap_m1_5_mod1)
    #conduct Tukey post hoc test 
    emm = emmeans(Mcap_m1_5_mod1, ~ B2015, adjust="tukey")
    #letter display
    cld(emm)

#Month3
Mcap_m3<-subset(Mcap_mort, month=="Month3")
Mcap_m3_mod1<-lm(score~Site*B2015, Mcap_m3)
summary(Mcap_m3_mod1)
anova(Mcap_m3_mod1)
  #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m3_mod1, ~ B2015, adjust="tukey") #B2015
      #letter display
      cld(emm)
      emm = emmeans(Mcap_m3_mod1, ~ Site, adjust="tukey") #B2015
      #letter display
      cld(emm)
      #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m3_mod1, ~ Site*B2015, adjust="tukey")  #B2015*Site
      #letter display
      pairs(emm)

#Month4.5
Mcap_m4_5<-subset(Mcap_mort, month=="Month4.5")
Mcap_m4_5_mod1<-lm(score~Site*B2015, Mcap_m4_5)
summary(Mcap_m4_5_mod1)
anova(Mcap_m4_5_mod1)
 #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m4_5_mod1, ~ B2015, adjust="tukey") #B2015
      #letter display
      cld(emm)
      #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m4_5_mod1, ~ Site, adjust="tukey")  #Site
      #letter display
      cld(emm)
      
#Month6
Mcap_m6<-subset(Mcap_mort, month=="Month6")
Mcap_m6_mod1<-lm(score~Site*B2015, Mcap_m6)
summary(Mcap_m6_mod1)
anova(Mcap_m6_mod1)
  #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m6_mod1, ~ B2015, adjust="tukey") #B2015
      #letter display
      cld(emm)
      #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m6_mod1, ~ Site, adjust="tukey")  #Site
      #letter display
      cld(emm)
      
#Month10
Mcap_m10<-subset(Mcap_mort, month=="Month10")
Mcap_m10_mod1<-lm(score~Site*B2015, Mcap_m10)
summary(Mcap_m10_mod1)
anova(Mcap_m10_mod1)
  #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m10_mod1, ~ B2015, adjust="tukey") #B2015
      #letter display
      cld(emm)
      #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m10_mod1, ~ Site, adjust="tukey")  #Site
      #letter display
      cld(emm)
      
#Month18
Mcap_m18<-subset(Mcap_mort, month=="Month18")
Mcap_m18_mod1<-lm(score~Site*B2015, Mcap_m18)
summary(Mcap_m18_mod1)
anova(Mcap_m18_mod1)
      #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m18_mod1, ~ Site, adjust="tukey")  #Site
      #letter display
      cld(emm)
      
#Month24
Mcap_m24<-subset(Mcap_mort, month=="Month24")
Mcap_m24_mod1<-lm(score~Site*B2015, Mcap_m24)
summary(Mcap_m24_mod1)
anova(Mcap_m24_mod1)
      #conduct Tukey post hoc test 
      emm = emmeans(Mcap_m24_mod1, ~ Site, adjust="tukey")  #Site
      #letter display
      cld(emm)
```

*Porites compressa* 
```{r}
#Month0
Pcomp_m0<-subset(Pcomp_mort, month=="Month0")
Pcomp_m0_mod1<-lm(score~Site*B2015, Pcomp_m0)
summary(Pcomp_m0_mod1) 
anova(Pcomp_m0_mod1) 
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m0_mod1, ~ Site*B2015, adjust="tukey")
  #letter display
  cld(emm)

# Month1.5
Pcomp_m1_5<-subset(Pcomp_mort, month=="Month1.5")
Pcomp_m1_5_mod1<-lm(score~Site*B2015, Pcomp_m1_5)
summary(Pcomp_m1_5_mod1)
anova(Pcomp_m1_5_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m1_5_mod1, ~ Site*B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month3
Pcomp_m3<-subset(Pcomp_mort, month=="Month3")
Pcomp_m3_mod1<-lm(score~Site*B2015, Pcomp_m3)
summary(Pcomp_m3_mod1)
anova(Pcomp_m3_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m3_mod1, ~ Site*B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month4.5
Pcomp_m4_5<-subset(Pcomp_mort, month=="Month4.5")
Pcomp_m4_5_mod1<-lm(score~Site*B2015, Pcomp_m4_5)
summary(Pcomp_m4_5_mod1)
anova(Pcomp_m4_5_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m4_5_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month6
Pcomp_m6<-subset(Pcomp_mort, month=="Month6")
Pcomp_m6_mod1<-lm(score~Site*B2015, Pcomp_m6)
summary(Pcomp_m6_mod1)
anova(Pcomp_m6_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m6_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month10
Pcomp_m10<-subset(Pcomp_mort, month=="Month10")
Pcomp_m10_mod1<-lm(score~Site*B2015, Pcomp_m10)
summary(Pcomp_m10_mod1)
anova(Pcomp_m10_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m10_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m10_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
  
#Month18
Pcomp_m18<-subset(Pcomp_mort, month=="Month18")
Pcomp_m18_mod1<-lm(score~Site*B2015, Pcomp_m18)
#ummary(Pcomp_m18_mod1)
anova(Pcomp_m18_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m18_mod1, ~ Site, adjust="tukey")
  #letter display
  pairs(emm)

#Month24
Pcomp_m24<-subset(Pcomp_mort, month=="Month24")
Pcomp_m24_mod1<-lm(score~Site*B2015, Pcomp_m24)
summary(Pcomp_m24_mod1)
anova(Pcomp_m24_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(Pcomp_m24_mod1, ~ Site, adjust="tukey")
  #letter display
  pairs(emm)
```

# Pairs models comparison   
(only include colonies when both partners of a pair are found at a time point)
```{r, results="hide", message=FALSE}
#Figure out which of the pairs we lose 
#dataset to work from (pull unwanted from)
pData<-mALLDataPER

#subset by month
M0<-subset(pData,month=="Month0")
M1.5<-subset(pData,month=="Month1.5")
M3<-subset(pData,month=="Month3")
M4.5<-subset(pData,month=="Month4.5")
M6<-subset(pData,month=="Month6")
M10<-subset(pData,month=="Month10")
M18<-subset(pData,month=="Month18")
M24<-subset(pData,month=="Month24")

# Identify which pairs have a partner not found
# month 0
 M0diff<-M0 
  M0diff<-dcast(M0diff, Species+Site+Pair+month~B2015) # cast()
  M0diff$diff<-M0diff$N-M0diff$Y

# Subset Month1.5 
  M1.5<-M1.5[!(M1.5$Pair=="113_114"),]
 #DIFFS
  M1.5diff<-M1.5  
  M1.5diff<-dcast(M1.5diff, Species+Site+Pair+month~B2015) # cast()
  M1.5diff$diff<-M1.5diff$N-M1.5diff$Y
  
## Subset Month3
  # remove 
  M3<-M3[!(M3$Pair=="7_8"),]
  M3<-M3[!(M3$Pair=="113_114"),]
  M3<-M3[!(M3$Pair=="255_256"),]
  M3<-M3[!(M3$Pair=="259_260"),]
  #DIFFS
  M3diff<-M3  
  M3diff<-dcast(M3diff, Species+Site+Pair+month~B2015) # cast()
  M3diff$diff<-M3diff$N-M3diff$Y

## Subset Month 4.5
  # remove 
  M4.5<-M4.5[!(M4.5$Pair=="7_8"),]
  M4.5<-M4.5[!(M4.5$Pair=="113_114"),]
  M4.5<-M4.5[!(M4.5$Pair=="255_256"),]
  M4.5<-M4.5[!(M4.5$Pair=="259_260"),]
  M4.5<-M4.5[!(M4.5$Pair=="263_264"),]
  M4.5<-M4.5[!(M4.5$Pair=="267_268"),]
  M4.5<-M4.5[!(M4.5$Pair=="82_83"),]
 #DIFFS
  M4.5diff<-M4.5  
  M4.5diff<-dcast(M4.5diff, Species+Site+Pair+month~B2015) # cast()
  M4.5diff$diff<-M4.5diff$N-M4.5diff$Y

# Subset month 6
  # remove 
  M6<-M6[!(M6$Pair=="111_112"),] # remove the other partner
  M6<-M6[!(M6$Pair=="115_116"),]
  M6<-M6[!(M6$Pair=="119_120"),]
  M6<-M6[!(M6$Pair=="207_208"),]
  M6<-M6[!(M6$Pair=="255_256"),]
  M6<-M6[!(M6$Pair=="259_260"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  M6<-M6[!(M6$Pair=="267_268"),]
  M6<-M6[!(M6$Pair=="263_264"),]
  #DIFFS
  M6diff<-M6  
  M6diff<-dcast(M6diff, Species+Site+Pair+month~B2015) # cast()
  M6diff$diff<-M6diff$N-M6diff$Y

  # Subset month 10
  #M10<-na.omit(M10) # remove the NA colums
  # remove 253_254, 259_260, 161_262, 263_264, 267_268, 271_272, 273_274, 275_276,277_278 , 31_2
  M10<-M10[!(M10$Pair=="253_254"),] # remove the other partner
  M10<-M10[!(M10$Pair=="259_260"),]
  M10<-M10[!(M10$Pair=="113_114"),]
  M10<-M10[!(M10$Pair=="161_262"),]
  M10<-M10[!(M10$Pair=="263_264"),]
  M10<-M10[!(M10$Pair=="267_268"),]
  M10<-M10[!(M10$Pair=="271_272"),]
  M10<-M10[!(M10$Pair=="273_274"),]
  M10<-M10[!(M10$Pair=="275_276"),]
  M10<-M10[!(M10$Pair=="277_278"),]
  M10<-M10[!(M10$Pair=="31_32"),]
  M10<-M10[!(M10$Pair=="261_262"),]

  #DIFFS
  M10diff<-M10  
  M10diff<-dcast(M10diff, Species+Site+Pair+month~B2015) # cast()
  M10diff$diff<-M10diff$N-M10diff$Y
  
# Subset month 18
  M18<-M18[!(M18$Pair=="253_254"),] # remove the other partner
  M18<-M18[!(M18$Pair=="259_260"),]
  M18<-M18[!(M18$Pair=="161_262"),]
  M18<-M18[!(M18$Pair=="263_264"),]
  M18<-M18[!(M18$Pair=="267_268"),]
  M18<-M18[!(M18$Pair=="271_272"),]
  M18<-M18[!(M18$Pair=="273_274"),]
  M18<-M18[!(M18$Pair=="275_276"),]
  M18<-M18[!(M18$Pair=="277_278"),]
  M18<-M18[!(M18$Pair=="31_2"),]
  M18<-M18[!(M18$Pair=="115_116"),]
  M18<-M18[!(M18$Pair=="119_120"),]
  M18<-M18[!(M18$Pair=="261_262"),]
  M18<-M18[!(M18$Pair=="269_270"),]
  M18<-M18[!(M18$Pair=="101_102"),]
  M18<-M18[!(M18$Pair=="117_118"),]
  M18<-M18[!(M18$Pair=="265_266"),]
  M18<-M18[!(M18$Pair=="19_20"),]
  M18<-M18[!(M18$Pair=="205_206"),]
  M18<-M18[!(M18$Pair=="223_224"),]
  M18<-M18[!(M18$Pair=="39_40"),]
    M18<-M18[!(M18$Pair=="103_104"),]

  #DIFFS
  M18diff<-M18  
  M18diff<-dcast(M18diff, Species+Site+Pair+month~B2015) # cast()
  M18diff$diff<-M18diff$N-M18diff$Y
  
  # Subset month 24
  M24<-M24[!(M24$Pair=="107_108"),] # remove the other partner
  M24<-M24[!(M24$Pair=="13_14"),]
  M24<-M24[!(M24$Pair=="223_224"),]
  M24<-M24[!(M24$Pair=="231_232"),]
  M24<-M24[!(M24$Pair=="261_262"),]
  M24<-M24[!(M24$Pair=="263_264"),]
  M24<-M24[!(M24$Pair=="265_266"),]
  M24<-M24[!(M24$Pair=="283_284"),]
  M24<-M24[!(M24$Pair=="289_290"),]
  M24<-M24[!(M24$Pair=="31_32"),]
  M24<-M24[!(M24$Pair=="33_34"),]
  M24<-M24[!(M24$Pair=="39_40"),]
  M24<-M24[!(M24$Pair=="76_77"),]
  M24<-M24[!(M24$Pair=="78_79"),]
  #DIFFS
  M24diff<-M24  
  M24diff<-dcast(M24diff, Species+Site+Pair+month~B2015) # cast()
  M24diff$diff<-M24diff$N-M24diff$Y
  
#make dfs  
PairDat<-rbind(M0,M1.5,M3,M4.5,M6,M10,M18,M24)
diffDat<-rbind(M0diff,M1.5diff,M3diff,M4.5diff,M6diff,M10diff,M18diff,M24diff)

#Mean bleaching scores across the entire project

#calc average bleaching score per grouping by time point
pairmeans <- ddply(PairDat, c("Site", "Species", "B2015", "month"), summarise, #
                   N    = length(score[!is.na(score)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(score, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(score, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(score, na.rm=TRUE) #calculate max
)

pairmeans$code<-paste0(pairmeans$Site, pairmeans$B2015, pairmeans$Species)
```

#SI Table 8: Pairs
*M. capitata* partial mortality
```{r} 
#set up Pairs data
Mcappig<-subset(PairDat, Species =="Montipora capitata")            
  
#All
MpigPair_all_mod1<-lmer(score~Site*B2015*month+(1|TagID)+(1|Pair), Mcappig)
summary(MpigPair_all_mod1)
anova(MpigPair_all_mod1, type=2)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair_all_mod1, ~ Site*B2015*month, adjust="tukey")
  #letter display
  cld(emm)
```
By month:
```{r}
#Month0
MpigPair0<-subset(Mcappig, month =="Month0")           
MpigPair0_mod1<-lm(score~Site*B2015, MpigPair0)
summary(MpigPair0_mod1)
anova(MpigPair0_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair0_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month1.5
MpigPair1.5<-subset(Mcappig, month =="Month1.5")
MpigPair1.5_mod1<-lm(score~Site*B2015, MpigPair1.5)
summary(MpigPair1.5_mod1)
anova(MpigPair1.5_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair1.5_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair1.5_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)

#Month3
MpigPair3<-subset(Mcappig, month =="Month3")
MpigPair3_mod1<-lm(score~Site*B2015, MpigPair3)
summary(MpigPair3_mod1)
anova(MpigPair3_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair3_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
   emm = emmeans(MpigPair3_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)
  emm = emmeans(MpigPair3_mod1, ~ B2015*Site, adjust="tukey")
  #letter display
  cld(emm)

#Month4.5
MpigPair4.5<-subset(Mcappig, month =="Month4.5")
MpigPair4.5_mod1<-lm(score~Site*B2015, MpigPair4.5)
summary(MpigPair4.5_mod1)
anova(MpigPair4.5_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair4.5_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
   emm = emmeans(MpigPair4.5_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)

#Month6
MpigPair6<-subset(Mcappig, month =="Month6")
MpigPair6_mod1<-lm(score~Site*B2015, MpigPair6)
summary(MpigPair6_mod1)
anova(MpigPair6_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair6_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
   emm = emmeans(MpigPair6_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)
  
#Month10
MpigPair10<-subset(Mcappig, month =="Month10")
MpigPair10_mod1<-lm(score~Site*B2015, MpigPair10)
summary(MpigPair10_mod1)
anova(MpigPair10_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair10_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
   emm = emmeans(MpigPair10_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)
  
#Month18
MpigPair18<-subset(Mcappig, month =="Month18")
MpigPair18_mod1<-lm(score~Site*B2015, MpigPair18)
summary(MpigPair18_mod1)
anova(MpigPair18_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair18_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
   emm = emmeans(MpigPair18_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)

#Month24
MpigPair24<-subset(Mcappig, month =="Month24")
MpigPair24_mod1<-lm(score~Site*B2015, MpigPair24)
summary(MpigPair24_mod1)
anova(MpigPair24_mod1)
  #conduct Tukey post hoc test 
  emm = emmeans(MpigPair24_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)
```

*P. compressa* partial mortality pairs  
```{r}
#set up Pairs data
Pcomppig<-subset(PairDat, Species =="Porites compressa")            

#All
PpigPair_all_mod1<-lmer(score~Site*B2015*month+(1|TagID)+(1|Pair), Pcomppig) #if we add in (1|Pair) we get failure to converge
summary(PpigPair_all_mod1)
anova(PpigPair_all_mod1, type=2)
  #conduct Tukey post hoc test 
  emm = emmeans(PpigPair_all_mod1, ~ Site*B2015*month, adjust="tukey")
  #letter display
  cld(emm)
```
By month:
```{r}
#Month0
PpigPair0<-subset(Pcomppig, month =="Month0")           
PpigPair0_mod1<-lm(score~Site*B2015, PpigPair0)
summary(PpigPair0_mod1)
anova(PpigPair0_mod1)
 #conduct Tukey post hoc test 
  emm = emmeans(PpigPair0_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)  
  emm = emmeans(PpigPair0_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)  
  emm = emmeans(PpigPair0_mod1, ~ Site*B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month1.5
PpigPair1.5<-subset(Pcomppig, month =="Month1.5")
PpigPair1.5_mod1<-lm(score~Site*B2015, PpigPair1.5)
summary(PpigPair1.5_mod1)
anova(PpigPair1.5_mod1)
 #conduct Tukey post hoc test 
  emm = emmeans(PpigPair1.5_mod1, ~ Site, adjust="tukey")
  #letter display
  cld(emm)  
  emm = emmeans(PpigPair1.5_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)  
  emm = emmeans(PpigPair1.5_mod1, ~ Site*B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month3
PpigPair3<-subset(Pcomppig, month =="Month3")
PpigPair3_mod1<-lm(score~Site*B2015, PpigPair3)
summary(PpigPair3_mod1)
anova(PpigPair3_mod1)
  emm = emmeans(PpigPair3_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)  
  emm = emmeans(PpigPair3_mod1, ~ Site*B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month4.5
PpigPair4.5<-subset(Pcomppig, month =="Month4.5")
PpigPair4.5_mod1<-lm(score~Site*B2015, PpigPair4.5)
summary(PpigPair4.5_mod1)
anova(PpigPair4.5_mod1)
  emm = emmeans(PpigPair4.5_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)  

#Month6
PpigPair6<-subset(Pcomppig, month =="Month6")
PpigPair6_mod1<-lm(score~Site*B2015, PpigPair6)
summary(PpigPair6_mod1)
anova(PpigPair6_mod1)
  emm = emmeans(PpigPair6_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)
  
#Month10
PpigPair10<-subset(Pcomppig, month =="Month10")
PpigPair10_mod1<-lm(score~Site*B2015, PpigPair10)
summary(PpigPair10_mod1)
anova(PpigPair10_mod1)
  emm = emmeans(PpigPair10_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)

#Month18
PpigPair18<-subset(Pcomppig, month =="Month18")
PpigPair18_mod1<-lm(score~Site*B2015, PpigPair18)
summary(PpigPair18_mod1)
anova(PpigPair18_mod1)
  emm = emmeans(PpigPair18_mod1, ~ B2015, adjust="tukey")
  #letter display
  cld(emm)  
  emm = emmeans(PpigPair18_mod1, ~ Site*B2015, adjust="tukey")
  cld(emm)  

#Month24
PpigPair24<-subset(Pcomppig, month =="Month24")
PpigPair24_mod1<-lm(score~Site*B2015, PpigPair24)
summary(PpigPair24_mod1)
anova(PpigPair24_mod1)
  emm = emmeans(PpigPair24_mod1, ~ Site*B2015, adjust="tukey")
  cld(emm)  
```

