---
title: "transect_reps"
author: "Shayle Matsuda"
date: "10/23/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

This script is the analysis of the benthic transect data for the following manuscript:

Matsuda, S.B., Huffmyer, A., Lenz, E., Davidson, J., Hancock, J., Przybylowski, A., Innis, T., Gates, R., Barott, K. (2020). Coral bleaching susceptibility is predictive of subsequent mortality within but not between coral species, Front. Ecol. Evol., doi: 10.3389/fevo.2020.00178

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results="hide", message=FALSE}
rm(list=ls(all=TRUE)) 

library("reshape2") #reshape data
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library("multcomp") 
library("multcompView")
library(glmmTMB)
library(lme4)
library(lmerTest)
library(tidyverse)
```

Setup df:
```{r,results="hide", message=FALSE}
#Setup workspace:
rData1<- read.csv("Benthic_Transect_df.csv", header=T) # load in transect data

# set levels of factor for Bleaching_status
levels(rData1$Bleaching_score)
rData1$BNB<-as.factor(rData1$BNB)
rData1$MonthFactor<-rData1$Month

#combine month 18 and 19 for end of experiment time point comparisons (<3 weeks apart)
rData1$MonthFactor[rData1$MonthFactor=="18"] <-"1819"
rData1$MonthFactor[rData1$MonthFactor=="19"] <-"1819"
rData1$MonthFactor <- factor(rData1$MonthFactor, levels=c("0", "1.5", "4.5","10","1819")) #specify order
```

# Transect color score prevalence plots
SI Figure 2, Mean Table
```{r, results="hide", message=FALSE}
# Calc. mean, sd,se   
repcol<-c("Species", "Reef", "Month","MonthFactor", "Bleaching_status","Percent")
  rDataPlot<-rData1[repcol]

# calc average  bleaching rate/month in %  per grouping
transectPlotmean <- ddply(rDataPlot, c("Species", "Reef", "Month","MonthFactor", "Bleaching_status"), summarise, 
                   N    = length(Percent[!is.na(Percent)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Percent, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Percent, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Percent, na.rm=TRUE) #calculate max 
)

transectPlotmean$code<-paste0(transectPlotmean$Species, transectPlotmean$Reef, transectPlotmean$Bleaching_status) #code for plotting
```

Subsets by Coral and Lagoon for plots:   
*Montipora capitata* Inner Lagoon  
```{r, results="hide", message=FALSE} 
McapInner1<-subset(transectPlotmean, Species=="Mcap")
McapInner<-subset(McapInner1, Reef=="Inner")

col_McapInner <- McapInner %>%   # height for error labels in the bars
group_by(Month) %>%
arrange(Month, desc(Bleaching_status)) %>%
mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_McapInner

# set to 2 decimal places
col_McapInner[,'mean']=round(col_McapInner[,'mean'],2)

#plot
col_McapInner_plot<-ggplot(data=col_McapInner, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Montipora capitata\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    ylim(0, 100) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5))
McapInner<-col_McapInner_plot+ theme(legend.position = "none") #remove legend
```

*Montipora capitata* Outer Lagoon
```{r, results="hide", message=FALSE}
McapOuter<-subset(McapInner1, Reef=="Outer")

col_McapOuter <- McapOuter %>%   
group_by(Month) %>%
arrange(Month, desc(Bleaching_status)) %>%
mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_McapOuter

# set to 2 decimal places
col_McapOuter[,'mean']=round(col_McapOuter[,'mean'],2)

col_McapOuter_plot<-ggplot(data=col_McapOuter, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Montipora capitata\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    ylim(0, 100) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5))
McapOuter<-col_McapOuter_plot+ theme(legend.position = "none") #remove legend
```

*Porites compressa* Inner Lagoon
```{r, results="hide", message=FALSE}
PcompInner1<-subset(transectPlotmean, Species=="Pcomp")
PcompInner<-subset(PcompInner1, Reef=="Inner")

col_PcompInner <- PcompInner %>%   
group_by(Month) %>%
arrange(Month, desc(Bleaching_status)) %>%
mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_PcompInner

# set to 2 decimal places
col_PcompInner[,'mean']=round(col_PcompInner[,'mean'],2)

col_PcompInner_plot<-ggplot(data=col_PcompInner, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Porites compressa\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    ylim(0,100) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
          theme(plot.title = element_text(size=20,face = "italic"))
PcompInner<-col_PcompInner_plot+ theme(legend.position = "none") #remove legend
```

*Porites compressa* Outer Lagoon
```{r, results="hide", message=FALSE}
PcompOuter<-subset(PcompInner1, Reef=="Outer")

col_PcompOuter <- PcompOuter %>% 
group_by(Month) %>%
arrange(Month, desc(Bleaching_status)) %>%
mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_PcompOuter

# set to 2 decimal places
col_PcompOuter[,'mean']=round(col_PcompOuter[,'mean'],2)

col_PcompOuter_plot<-ggplot(data=col_PcompOuter, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Porites compressa\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    ylim(0,100) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5))
PcompOuter<-col_PcompOuter_plot+ theme(legend.position = "none") #remove legend
```

Plot all four together:  Manuscript SI Figure 1
```{r}
grid.arrange(McapInner, McapOuter, PcompInner,PcompOuter, nrow = 2)
```

#Coral bleaching prevalence  
Sum: White (W) + Pale (P) 
Set up df:  
```{r, results="hide", message=FALSE}
rData2<-rData1[!(rData1$Bleaching_status=="Healthy"),]    # Remove Healthy 
rData2$code<-paste0(rData2$Species, rData2$Month, rData2$MonthFactor, rData2$Depth, rData2$Reef, rData2$Rep)#create a code for groups by Species, Month, Depth, Reef, Rep 
rData3<-rData2 %>% group_by(code) %>%summarise(sum=sum(Percent))# Group Data: combine P and B %s

# Match this back with data cols you want
rData3$Species<-rData2$Species[match(rData3$code, rData2$code)]
rData3$Month<-rData2$Month[match(rData3$code, rData2$code)]
rData3$Depth<-rData2$Depth[match(rData3$code, rData2$code)]
rData3$Reef<-rData2$Reef[match(rData3$code, rData2$code)]
rData3$Rep<-rData2$Rep[match(rData3$code, rData2$code)]
rData3$MonthFactor<-rData2$MonthFactor[match(rData3$code, rData2$code)]
rData3<-rData3[,c(1,3,5:8,2)]
rData3$Depth<-as.factor(rData3$Depth)
rData3$Rep<-as.factor(rData3$Rep)
```

#Benthic section: %Coral cover
Benthic transect df setup:   
```{r, results="hide", message=FALSE} 
ccData<-read.csv("Benthic_Transect_Coral_Cover.csv", header=T)
ccData$MonthFactor<-ccData$Month
ccData$MonthFactor[ccData$MonthFactor=="18"] <-"1819" #combine months 18 and 19
ccData$MonthFactor[ccData$MonthFactor=="19"] <-"1819"
ccData$MonthFactor <- factor(ccData$MonthFactor, levels=c("0", "1.5", "4.5","10","1819")) 
#specify order
ccData$Reef<-as.factor(ccData$Reef)
ccData$Depth<-as.factor(ccData$Depth)

# calc average  bleaching rate/month in %  per grouping
ccMeans <- ddply(ccData, c("Reef",  "Month"), summarise, 
                   N    = length(CoralCover[!is.na(CoralCover)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(CoralCover, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(CoralCover, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(CoralCover, na.rm=TRUE) #calculate max
)
```

#Supplementary Table 1  
%Coral cover full model lmer 
```{r} 
CoralCoverMod<-lmer(CoralCover~Reef*MonthFactor+(1|Depth), ccData)
anova(CoralCoverMod, type=2)

#conduct Tukey post hoc test 
    emm = emmeans(CoralCoverMod, ~ Reef, adjust="tukey")
    #letter display
    pairs(emm)
```

#Supplementary Table 2
Bleaching prevalence, all timepoints and by month: W+P  
Full model:lmer(sum~Reef*MonthFactor*Species+(1|Depth))
```{r}  
#fullmodel, df=rData3
Prev_transect_full<-lmer(sum~Reef*MonthFactor*Species+(1|Depth), data=rData3)
anova(Prev_transect_full, type=2)
    #conduct Tukey post hoc test 
    emm = emmeans(Prev_transect_full, ~ Reef*MonthFactor, adjust="tukey")
    #letter display
      pairs(emm) 
    emm = emmeans(Prev_transect_full, ~ Species*MonthFactor, adjust="tukey")
    #letter display
    pairs(emm) 
        cld(emm) 
```


#Supplementary Table 3  
Bleaching prevalence: White only 
```{r}
#bleached
prevWhite<-subset(rData1, Bleaching_status=="Bleached")

WhiteMod<-lmer(Percent~Species*Reef*MonthFactor+(1|Depth),prevWhite)
anova(WhiteMod, type=2)
  #conduct Tukey post hoc test 
    emm = emmeans(WhiteMod, ~ Reef, adjust="tukey")
    #letter display
    cld(emm)
  #conduct Tukey post hoc test 
    emm = emmeans(WhiteMod, ~ Reef*Species*MonthFactor, adjust="tukey")
    #letter display
    cld(emm)
    pairs(emm)
```

#Supplementary Table 4
Bleaching severity (white/white+pale)  
lmer(newPer~Species*Reef*MonthFactor+(1|Depth)+(1|Rep))
```{r}
rData4<-rData2 #make a copy of rData2
rData4$Depth<-as.factor(rData4$Depth)
rData4$Rep<-as.factor(rData4$Rep)

rData4$sum<-rData3$sum[match(rData4$code, rData3$code)] # add the sum row by code to get sum of W+P
rData4$newPer<-rData4$Percent/rData4$sum # Make new row of new percents which is W%/(W%+P%)
rData4<-subset(rData4, Bleaching_status=="Bleached") # get rid of Pale

#full model
SeverityMod<-lmer(newPer~Species*Reef*MonthFactor+(1|Depth),rData4)
anova(SeverityMod, type="2")
 #conduct Tukey post hoc test 
    emm = emmeans(SeverityMod, ~ Species*MonthFactor*Reef, adjust="tukey")
    #letter display
    cld(emm)
    pairs(emm) 
 #conduct Tukey post hoc test 
    emm = emmeans(SeverityMod, ~ Reef*Species*MonthFactor, adjust="tukey")
    #letter display
    pairs(emm)   
```
