---
title: "transect_reps"
author: "Shayle Matsuda"
date: "10/23/2019"
output:
  pdf_document:
    toc: yes
    toc_depth: '6'
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    code_folding: hide
    keep_tex: yes
    pdf_document: null
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script is for the 2015 bleaching recovery paper's transect data and the Inner and Outer Lagoons. Here, we look at std error in the transect replicates. 

This script includes a look at depth, it's not significant at anything but 4.5. We are dropping it.   


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results="hide", message=FALSE}
rm(list=ls(all=TRUE)) 
#setwd("~/Desktop/B_NB")

library("reshape2") #reshape data
library("plotrix") #functions in tapply
library("ggplot2") #plotting
library("reshape") #reshape data
library("plyr")  #splitting, applying, and combining data
library("dplyr")  #splitting, applying, and combining data
library("pscl")
library("plotrix") #plotting
library("gridExtra") #arrange plots for output
library("car") #levenes test
library("lsmeans")
library(effects)
library("multcomp") 
library("multcompView")
library("coefplot") #coefficient plotting
library("seacarb") #seawater carbonate chemistry
library("vegan") #calculating distance matrices
library(betareg)
library(coin)
library(exactRankTests)
library(mvtnorm)
library(DTK)
library(PMCMR)
library(MuMIn)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(ggsignif)
library(latticeExtra)
library(HH)
library(PMCMR)
library(tidyverse)
library(arm)
library(jtools)
library(broom)
library(ggstance)
library(officer)
library(huxtable)
library(flextable)
library(ARTool)
```

```{r,results="hide", message=FALSE}
#Setup workspace:
rData1<- read.csv("transect_data_20191022.csv", header=T) # load in transect data for corals by W, P, H

# set levels of factor for Bleaching_status
levels(rData1$Bleaching_score)
rData1$MonthFactor<-as.factor(rData1$Month)  #create this to order axis in plots
rData1$BNB<-as.factor(rData1$BNB)

rData1$MonthFactor <- factor(rData1$MonthFactor, levels=c("0", "1.5", "4.5","10","18","19")) #specify order
```

# Transect color score prevalence plots
SI Figure 1, Mean Table
```{r, results="hide", message=FALSE}
# Calc. mean, sd,se
# remove rep column
repcol<-c("Species", "Reef", "Month","MonthFactor", "Bleaching_status","Percent")
  rDataPlot<-rData1[repcol]

# calc average  bleaching rate/month in %  per grouping
transectPlotmean <- ddply(rDataPlot, c("Species", "Reef", "Month","MonthFactor", "Bleaching_status"), summarise, 
                   N    = length(Percent[!is.na(Percent)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(Percent, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(Percent, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(Percent, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
#Monthmeans #display table

transectPlotmean$code<-paste0(transectPlotmean$Species, transectPlotmean$Reef, transectPlotmean$Bleaching_status)
```

Subsets by Coral and Lagoon:   
*Montipora capitata* Inner Lagoon  
```{r, results="hide", message=FALSE}
McapInner1<-subset(transectPlotmean, Species=="Mcap")
McapInner<-subset(McapInner1, Reef=="Inner")

 col_McapInner <- McapInner %>%   # do this to get height for your labels in the bars
  group_by(Month) %>%
 arrange(Month, desc(Bleaching_status)) %>%
 mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_McapInner

# set to 2 decimal places
col_McapInner[,'mean']=round(col_McapInner[,'mean'],2)

col_McapInner_plot<-ggplot(data=col_McapInner, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  #scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Montipora capitata\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    #ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));col_McapInner_plot
McapInner<-col_McapInner_plot+ theme(legend.position = "none") #remove legend
```

*Montipora capitata* Outer Lagoon
```{r, results="hide", message=FALSE}
McapOuter<-subset(McapInner1, Reef=="Outer")

 col_McapOuter <- McapOuter %>%   # do this to get height for your labels in the bars
  group_by(Month) %>%
 arrange(Month, desc(Bleaching_status)) %>%
 mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_McapOuter

# set to 2 decimal places
col_McapOuter[,'mean']=round(col_McapOuter[,'mean'],2)

col_McapOuter_plot<-ggplot(data=col_McapOuter, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  #scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Montipora capitata\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    #ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));col_McapOuter_plot
McapOuter<-col_McapOuter_plot+ theme(legend.position = "none") #remove legend

```

*Porites compressa* Inner Lagoon
```{r, results="hide", message=FALSE}
PcompInner1<-subset(transectPlotmean, Species=="Pcomp")
PcompInner<-subset(PcompInner1, Reef=="Inner")

 col_PcompInner <- PcompInner %>%   # do this to get height for your labels in the bars
  group_by(Month) %>%
 arrange(Month, desc(Bleaching_status)) %>%
 mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_PcompInner

# set to 2 decimal places
col_PcompInner[,'mean']=round(col_PcompInner[,'mean'],2)

col_PcompInner_plot<-ggplot(data=col_PcompInner, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  #scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Porites compressa\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    #ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Inner Lagoon")+
          theme(plot.title = element_text(size=20,face = "italic"));col_PcompInner_plot
PcompInner<-col_PcompInner_plot+ theme(legend.position = "none") #remove legend

```

*Porites compressa* Outer Lagoon
```{r, results="hide", message=FALSE}
PcompOuter<-subset(PcompInner1, Reef=="Outer")

 col_PcompOuter <- PcompOuter %>%   # do this to get height for your labels in the bars
  group_by(Month) %>%
 arrange(Month, desc(Bleaching_status)) %>%
 mutate(lab_ypos = cumsum(mean) - 0.5 * mean) 
col_PcompOuter

# set to 2 decimal places
col_PcompOuter[,'mean']=round(col_PcompOuter[,'mean'],2)

col_PcompOuter_plot<-ggplot(data=col_PcompOuter, aes(x=Month, y=mean, group = code, color=Bleaching_status)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1)+
  geom_point(aes(shape=Bleaching_status), size=3)+
  geom_line(aes(color=Bleaching_status, linetype=Bleaching_status))+
  #scale_linetype_manual(values=c("solid", "dotted"))+
  scale_color_manual(values=c("black", "black","black"))+
  ylab("Porites compressa\nPercent") +
    xlab("Recovery time (months)") + #Label the X Axis
    #ylim(1, 3) + #set Y limits
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  #ggtitle("Outer Lagoon")+
          theme(plot.title = element_text(size=20,face = "italic",hjust = 0.5));col_PcompOuter_plot
PcompOuter<-col_PcompOuter_plot+ theme(legend.position = "none") #remove legend

```

Plot all four together: for Manuscript SI Figure 1
```{r}
grid.arrange(McapInner, McapOuter, PcompInner,PcompOuter, nrow = 2)
```

#Coral bleaching prevalence
Here we sum White (W) + Pale (P) to get the total % of affected corals.
```{r}

rData2<-rData1[!(rData1$Bleaching_status=="Healthy"),]    # Remove Healthy 
rData2$code<-paste0(rData2$Species, rData2$Month, rData2$MonthFactor, rData2$Depth, rData2$Reef, rData2$Rep)#create a code for groups by Species, Month, Depth, Reef, Rep 
rData3<-rData2 %>% group_by(code) %>%summarise(sum=sum(Percent))# Group Data: combine P and B %s

# Match this back with data cols you want
rData3$Species<-rData2$Species[match(rData3$code, rData2$code)]
rData3$Month<-rData2$Month[match(rData3$code, rData2$code)]
rData3$Depth<-rData2$Depth[match(rData3$code, rData2$code)]
rData3$Reef<-rData2$Reef[match(rData3$code, rData2$code)]
rData3$Rep<-rData2$Rep[match(rData3$code, rData2$code)]
rData3$MonthFactor<-rData2$MonthFactor[match(rData3$code, rData2$code)]
rData3<-rData3[,c(1,3,5:8,2)]
# subset by species
BS_mcap<-subset(rData3, Species=="Mcap")
BS_mcap<-BS_mcap[,c(1,3:7)]
BS_pcomp<-subset(rData3, Species=="Pcomp")
BS_pcomp<-BS_pcomp[,c(1,3:7)]

#subset by reef
#BS_Inner<-subset(rData3, Reef=="Inner")
```

ARTools: full model 
```{r}
#Mcap
model1<-art(sum~Reef*MonthFactor+(1|Depth/Rep), data=subset(BS_mcap))
anova(model1)   # AH getting some rank deficient warnings....thoughts?
#emmeans(artlm(model1, "MonthFactor"), pairwise ~MonthFactor) #this won't work bc of multiple factors
library(phia)
		testInteractions(artlm(model1, "Reef:MonthFactor"), pairwise=c("Reef","MonthFactor"), adjustment="holm")

```










Prevalence by month - ANOVA 
Mcap and Pcomp by Lagoon at each month
```{r}
#mcap 0
BSc_mcap<-aov(sum~Reef, data=subset(BS_mcap, Month=="0"))
#qqPlot(residuals(BSc_mcap))
summary(BSc_mcap)
TukeyHSD(BSc_mcap)

#pcomp 0
BSc_pcomp<-aov(sum~Reef, data=subset(BS_pcomp, Month=="0"))
#qqPlot(residuals(BSc_pcomp)) 
summary(BSc_pcomp)
TukeyHSD(BSc_pcomp)

#mcap 1.5
BSc_mcap1.5<-aov(sum~Reef, data=subset(BS_mcap, Month=="1.5"))
#qqPlot(residuals(BSc_mcap1.5)) 
summary(BSc_mcap1.5)

#pcomp 1.5
BSc_pcomp1.5<-aov(sum~Reef, data=subset(BS_pcomp, Month=="1.5"))
#qqPlot(residuals(BSc_pcomp1.5)) 
summary(BSc_pcomp1.5)
TukeyHSD(BSc_pcomp1.5)

#mcap 4.5
BSc_mcap4.5<-aov(sum~Reef, data=subset(BS_mcap, Month=="4.5"))
#qqPlot(residuals(BSc_mcap4.5)) 
summary(BSc_mcap4.5)

#pcomp 4.5
BSc_pcomp4.5<-aov(sum~Reef, data=subset(BS_pcomp, Month=="4.5"))
#qqPlot(residuals(BSc_pcomp4.5)) 
summary(BSc_pcomp4.5)

#mcap 10
BSc_mcap10<-aov(sum~Reef, data=subset(BS_mcap, Month=="10"))
#qqPlot(residuals(BSc_mcap10)) 
summary(BSc_mcap10)

#pcomp 10
BSc_pcomp10<-aov(sum~Reef, data=subset(BS_pcomp, Month=="10"))
#qqPlot(residuals(BSc_pcomp10)) 
summary(BSc_pcomp10)
```


#Benthic section: coral cover
Benthic transect data here.   
```{r}
ccData<-read.csv("/Users/shayle/Desktop//B_NB/benthicTransectCoralcover20191024.csv", header=T)
ccData$Reef<-as.character(ccData$Reef)
ccData$Month<-as.character(ccData$Month)

# calc average  bleaching rate/month in %  per grouping
ccMeans <- ddply(ccData, c("Reef",  "Month"), summarise, 
                   N    = length(CoralCover[!is.na(CoralCover)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(CoralCover, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(CoralCover, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(CoralCover, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
ccMeans
#write.csv(ccMeans, "ccMeans_20191024.csv")
```

SI Table 3: CoralCover~Lagoon at peak bleaching
```{r}
ccMonth0<-subset(ccData, Month=="0")
ccMonth0<-aov(CoralCover~Reef, ccMonth0)
#qqPlot(residuals(ccMonth0))
summary(ccMonth0) 
TukeyHSD(ccMonth0) 
```

SI Table 4: CoralCover~Month at each Lagoon
```{r}
#SI2
#coral cover at each reef
ccPR13<-subset(ccData, Reef=="13")
ccPR4<-subset(ccData, Reef=="4")

ccMod13<-aov(CoralCover~Month, ccPR13)
#qqPlot(residuals(ccMod13))
summary(ccMod13) 

ccMod4<-aov(CoralCover~Month, ccPR4)
#qqPlot(residuals(ccMod4)) # looks good
summary(ccMod4) # 
```

SI Table 5: Bleaching prevalence~Lagoon*Species at peak bleaching  
```{r}
# subset Mcap
fmod_mcap<-subset(rData3, Month=="0")
fmod_mcap0<-subset(fmod_mcap, Species=="Mcap")
fmod_mcap0Mod<-aov(sum~Reef, fmod_mcap0)
summary(fmod_mcap0Mod)
TukeyHSD(fmod_mcap0Mod)

# subset Pcomp
fmod_pcomp<-subset(rData3, Month=="0")
fmod_pcomp0<-subset(fmod_pcomp, Species=="Pcomp")
fmod_pcomp0Mod<-aov(sum~Reef, fmod_pcomp0)
summary(fmod_pcomp0Mod)
TukeyHSD(fmod_pcomp0Mod)
```

SI 6: Month 0, by reef, bleaching (white, not pale) prevalence.  
```{r}
#bleached
prevWhite<-subset(rData1, Bleaching_status=="Bleached")
prevWhite0<-subset(prevWhite, Month=="0")

#PR4
PrevWhite0Inner<-subset(prevWhite0, Reef=="Inner")
prev0WhiteMod4<-aov(Percent~Species, PrevWhite0Inner)
summary(prev0WhiteMod4)
TukeyHSD(prev0WhiteMod4)

#PR13
#PrevWhite0Outer<-subset(prevWhite0, Reef=="Outer")
#prev0WhiteMod13<-aov(Percent~Species, PrevWhite0Outer)
#summary(prev0WhiteMod13)
```

##Severity
```{r}
rData4<-rData2 #make a copy of rData2

# add the sum row by code to get sum of W+P
rData4$sum<-rData3$sum[match(rData4$code, rData3$code)]

# Make new row of new percents which is P%/(W+P%)
rData4$newPer<-rData4$Percent/rData4$sum

# calc average  bleaching rate/month in %  per grouping
rMeans <- ddply(rData4, c("Species", "Reef", "Month", "Bleaching_status"), summarise, 
                   N    = length(newPer[!is.na(newPer)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                   mean = mean(newPer, na.rm=TRUE), #calculate mean of response variable, removing NA's
                   sd   = sd(newPer, na.rm=TRUE), #calculate standard deviation
                   se   = sd / sqrt(N), #calculate standard error
                   max = max(newPer, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
rMeans
#write.csv(rMeans, "BleachingSeverityTransects_20191023.csv")
```

SI Table 7: Bleaching severity (white) at PR4 by coral species. 
```{r}
pr4Sev<-subset(rData4, Reef=="Inner") # subset reef
pr4Sev1<-subset(pr4Sev, Month=="0") #subset only month 0
pr4Sev2<-subset(pr4Sev1, Bleaching_status=="Bleached") #subset only bleached

pr4Sev2_aov<-aov(newPer~Species, pr4Sev2) 
summary(pr4Sev2_aov)
TukeyHSD(pr4Sev2_aov)
```

SI Table 8: *M. capitata* bleaching severity (white) at peak bleaching: white~Lagoon
```{r}
mcapSev<-subset(rData4, Species=="Mcap") # subset mcap
mcapSev<-subset(mcapSev, Bleaching_status=="Bleached") #subset only bleached
mcapSev_0<-subset(mcapSev, Month=="0") #subset month 0 only
mcapSev_0_aov<-aov(newPer~Reef, mcapSev_0) 
summary(mcapSev_0_aov)

#pcomp
pcompSev<-subset(rData4, Species=="Pcomp") # subset mcap
pcompSev<-subset(pcompSev, Bleaching_status=="Bleached") #subset only bleached
pcompSev_0<-subset(pcompSev, Month=="0") #subset month 0 only
pcompSev_0_aov<-aov(newPer~Reef, pcompSev_0) 
summary(pcompSev_0_aov)
TukeyHSD(pcompSev_0_aov)
```


SI Table 9: Prevalence of bleaching (white+pale) post peak bleaching during the first year of recovery. 
```{r}
#Month 1.5
M1.5_post<-aov(sum~Species,   data=subset(rData3,Month=="1.5"))
summary(M1.5_post)
TukeyHSD(M1.5_post) #  significant. 

#Month 4.5
M4.5_post<-aov(sum~Species,   data=subset(rData3,Month=="4.5"))
summary(M4.5_post)
TukeyHSD(M4.5_post) #  significant. 

#Month 10
M10_post<-aov(sum~Species,   data=subset(rData3,Month=="10"))
summary(M10_post)
TukeyHSD(M10_post) #  significant. 
```

SI Table 10: Bleaching prevalence (white+pale)~Lagoon at each month of recovery within the first year.     
```{r}
#mcap 0
BSc_mcap<-aov(sum~Reef, data=subset(BS_mcap, Month=="0"))
summary(BSc_mcap)
TukeyHSD(BSc_mcap)

#pcomp 0
BSc_pcomp<-aov(sum~Reef, data=subset(BS_pcomp, Month=="0"))
summary(BSc_pcomp)
TukeyHSD(BSc_pcomp)

#mcap 1.5
BSc_mcap1.5<-aov(sum~Reef, data=subset(BS_mcap, Month=="1.5"))
summary(BSc_mcap1.5)

#pcomp 1.5
BSc_pcomp1.5<-aov(sum~Reef, data=subset(BS_pcomp, Month=="1.5"))
summary(BSc_pcomp1.5)
TukeyHSD(BSc_pcomp1.5)

#mcap 4.5
BSc_mcap4.5<-aov(sum~Reef, data=subset(BS_mcap, Month=="4.5"))
summary(BSc_mcap4.5)

#pcomp 4.5
BSc_pcomp4.5<-aov(sum~Reef, data=subset(BS_pcomp, Month=="4.5"))
summary(BSc_pcomp4.5)

#mcap 10
BSc_mcap10<-aov(sum~Reef, data=subset(BS_mcap, Month=="10"))
summary(BSc_mcap10)

#pcomp 10
BSc_pcomp10<-aov(sum~Reef, data=subset(BS_pcomp, Month=="10"))
summary(BSc_pcomp10)
```
